### 2020-04-20 manuscript.tex
Last version from Alexey Pyrkov.

\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{color}
\usepackage{soul}
% \usepackage{floatflt}
\usepackage{graphicx}
\DeclareMathOperator{\tr}{Tr}
\newcommand*{\eq}[1]{(\ref{#1})}
\newcommand*{\eqs}[2]{(\ref{#1})--(\ref{#2})}
\newcommand{\exi}{\tr{\{e^{-\frac\beta2\sum_k\lambda_k\gamma_k^\dag\gamma_k}\}}}
\renewcommand{\l}{\left(}
\renewcommand{\r}{\right)}

\begin{document}
\title{Quantum assisted unsupervised data clustering on the basis of neural networks}

\author{I.~D.~Lazarev} 
\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
\affiliation{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}

\author{Marek Narozniak}
\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
\affiliation{Department of Physics, New York University, New York, NY, 10003, USA.}

\author{Tim Byrnes}
\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
\affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
\affiliation{Department of Physics, New York University, New York, NY 10003, USA}

\author{A.~N.~Pyrkov} 
\email{Email address:pyrkov@icp.ac.ru}
\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}

\date{\today}

\begin{abstract}
	Unsupervised machine learning is one of the main directions in the area of artificial intelligence. 
	Here, we propose a realization of quantum assisted unsupervised data clustering on the basis of artificial neural network called a self-organizing feature map. 
	We make a proof-of-concept realization of one of the crucial parts of the approach on IBM quantum machine and show that it allows us to reduce number of calculations in a number of clusters. 
	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.  
\end{abstract}

%\pacs{}
\maketitle

\section{Introduction}
The combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}. 
Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}. 
Paradoxically, the exact mechanism of why neural networks are so powerful remain unknown (in many cases it is regarded as a black box).  
It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
In particularly, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) that is trained in an unsupervised manner. 
SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}. 
The SOFMs with a small number of nodes are similar to the K-means algorithm but for larger SOFMs they represent data in a fundamentally topological way that allows one to do dimensionality reduction.
Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.

Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}. 
In the early stage, the main focus in QML was on the use of quantum computers to perform some basic linear algebra subroutines faster (such as Fourier transform and solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}), 
open up the possibilities for exponential speed-ups utilizing quantum phenomena such as quantum entanglement and quantum superposition over classical counterparts \cite{biamonte2017, schuld2014} and,
in the result, quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and so on \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}. 
Recently,  a number of shifts occurred and the hottest trend migrated to developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019}. 
The shift was inspired with the progress in experimental quantum computing when it became possible to play around with parametrized quantum circuits, whose parameters behave much like the weights of a neural network \cite{lewenstein1994}. 
In particular, quantum algorithms for training and evaluating feedforward neural networks, one of the most usable neural network model, was developed \cite{allcock2018, tacchino2019}. 
Recent results connected classical Bayesian approach to deep learning allowed to develop a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}. 
The quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}. Quantum classification, tested on the MNIST dataset, via Slow Feature Analysis based on the use of Quantum Frobenius Distance was proposed \cite{kerenidis2018}. 
Sublinear quantum algorithms for training linear and kernel-based classifiers were designed \cite{li2019}. 
Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, quantum autoencoders \cite{bondarenko2019}, quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, vinci2019, lu2019} were developed. 
Currently, it is believed that implementation of quantum neural networks can be the main test bed to achieve practical quantum supremacy on NISQ devices.

In this paper, we realize a hybrid quantum assisted SOFM (QASOFM) and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.   
We implement the quantum assisted SOFM in such a way that it becomes possible to use the Hamming distance as a distance metric for the training and to reduce the number of distance calculations in the number of clusters. 
We propose an optimized circuit for realizing the Hamming distance on a quantum machine. 
We prepare a toy example of data for clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} and compare it with the classical case.

\begin{figure}
	\label{fig:sofm_fitting}
	\includegraphics[width=0.95\columnwidth]{sofm_fitting.png}
	\caption{
		Schematic illustration of the clustering problem considered in this paper.  
		Blue dots represent clusters and red dots are data points. 
		The training process moves clusters to fit the data points. 
		Note that there are fewer clusters than data points which is the essence of dimensionality reduction and is what permits the model to generalize the data.
	}
\end{figure}

\section{The quantum assisted self-organizing feature map (SOFM)}
\label{sec:qasofm}

SOFM is one of the best-known unsupervised learning methods that is widely used in various areas of modern science. 
It was first proposed by Kohonen as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}. 
The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, as input and maps them onto nodes of a two-dimensional grid.

The input dimensions are associated with features and nodes in the grid, called cluster vectors, are assigned the $N$-dimensional vectors; 
the components of these vectors are usually called weights. 
Initially weight components are chosen randomly. 
We then can train our SOFM adjusting the components through the learning process which can be oversimplified into two basic procedures, 
selecting a winning cluster vector and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
In more detail, they consist of four step process: 
1) select an input vector randomly from the set of all input vectors; 
2) find a cluster vector which is closest to the input vector; 
3) adjust the weights of the winning node in such a way that it becomes even closer to the input vector; 
4) repeat this process many iterations until it converges.

After the winning cluster vector is selected, the weights of the vector are adjusted according to 
%
\begin{equation}
    \label{eq:learning}
    \vec w_{\mathrm{new}} =
		\vec{w}_{\mathrm{old}} 
        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  .
\end{equation}
%
In simple words, this expression can be understood according to the following: 
if a component of the input vector is greater than the corresponding weight, increase the weight by a small amount; 
if the input component is smaller than the weight, decrease the
weight by a small amount. 
The larger the difference between the input component and the weight component, the larger increment (decrement). 
Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects. 
A visualisation of this process is provided on figure (Fig.~\ref{fig:sofm_fitting}).

Usually the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors. 
In our approach, we use the Hamming distance instead of the Euclidean distance to select the winning cluster vector. 
It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.

\subsection{Optimized quantum scheme for Hamming distance calculation}
\label{subsec:qcircuit}

%%%%%
\begin{figure}[t]
	\label{fig:qcircuit}
	\includegraphics[width=0.95\columnwidth]{qcircuit.png}
	\caption{
    	%add the caption with intuitive explanation of the circuit
		A quantum circuit for the Hamming distance calculation algorithm between all samples and clusters at once. 
		First, we apply Hadamard gates in order to get superposition of the sample and cluster registers. 
		Second, we encoded information about pairwise different qubits in a quantum state of the sample register with applying the CNOT gates. 
		Third, Hamming distance values are encoded in the amplitudes of superposition with the control phase rotation and Hadamard gates. 
		Finally, a quantum state of the sample register returned to the initial basis for information retrieval.
	} 
\end{figure}
%%%%%

%%%%%
\begin{figure}[t]
	\label{fig:distance_matrix}
	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
	\caption{
		The Hamming distance matrix between two data sets of binary vectors. 
		Values of distance that less than median distance value was marked black. 
		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure. 
		Result obtained on  IBMQ "ibmq\_16\_melbourne" backend is shown on the right figure. 
		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
	} 
\end{figure}
%%%%%

We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors at once.  
This allows for a simple encoding of the classical information into a quantum register. 
Initially we have two registers each with $n$ qubits in that we encode our set of binary vectors defined as
%
\begin{align}
    \label{eq:encodnig}
    \left| X \right\rangle  & = \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
    \left| Y \right\rangle&  = \sum\limits_{i=1}^{l} \left| y_i \right\rangle,
\end{align}
% 
where $ 1 \le k, l \le n $.  
Here the registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ correspond to the input states and the cluster states respectively. 
The initial state of whole circuit is
%
\begin{equation} 
\left| \psi_0 \right\rangle = 
    \left| X \right\rangle
    \left| Y \right\rangle 
    \left| a \right\rangle
    \label{eq:initial_state}
\end{equation}
%
where $\left| a \right\rangle$ is an auxiliary qubit in the state $\left| 0 \right\rangle$ initially.

We then apply a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$

\begin{equation}
    \left| \psi_1 \right\rangle = 
    \mathrm{CNOT(Y,X)\left| \psi_0 \right\rangle} =  
    \sum\limits_{i, j}^{k} 
    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
    \left| y^1_j, \dots, y^n_j \right\rangle
    \left| 0 \right\rangle 
\end{equation}
where $d^\alpha_{ij} = \mathrm{CNOT}(y^\alpha_i, x^\alpha_j),\, \alpha = 1 \dots n$ and $i,j$ are the qubit indexes in the registers.

In the result, we encoded information about pairwise different qubits in $\{X\}$ and $\{Y\}$ into register $\left| X \right\rangle$.

After that we transfer information about distance between sample state and cluster state into the amplitude of the superposed state. 
In order to make that we apply Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,

\begin{equation}
    \label{eq:control_phase_rotation}
    R(\phi) = 
    \begin{pmatrix}
        1 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 \\
        0 & 0 & 1 & 0 \\
        0 & 0 & 0 & e^{-i\phi}
    \end{pmatrix},
    \quad \phi = \frac{\pi}{n}
\end{equation}
and again Hadamard gate.

After the first Hadamard on the ancillary qubit we have the following state

\begin{equation}
    \left| \psi_2 \right\rangle = H_a\left| \psi_1 \right\rangle = 
    \sum\limits_{i, j}^{k} 
    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
    \left| y^1_j, \dots, y^n_j \right\rangle
    \dfrac{\left| 0 \right\rangle + \left| 1 \right\rangle}{\sqrt{2}}
\end{equation}

Then we apply the control phase gate

\begin{multline}
    \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
    \\ = \dfrac{1}{\sqrt{2}}
				\sum\limits_{i, j}^{k} 
				\left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
        \left| y^1_j, \dots, y^n_j \right\rangle 
        \left| 0 \right\rangle
        \\ + \dfrac{1}{\sqrt{2}}
				\sum\limits_{i, j}^{k}
        \exp\left(\dfrac{-i \pi}{n}\sum\limits_l^n d^l_{ij} \right)
        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
        \left| y^1_j, \dots, y^n_j \right\rangle 
        \left| 1 \right\rangle
\end{multline}

%\\ = \dfrac{1}{\sqrt{2}}
%    \sum\limits_{i, j}^{k}
%    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
%        \left| y^1_j, \dots, y^n_j \right\rangle 
%        \left| 0 \right\rangle
%        \\ + \dfrac{1}{\sqrt{2}}
%				\sum\limits_{i, j}^{k}
%        \left| R(d^1_{ij}, 1)\left(\dfrac{\pi}{n}\right), \dots, R(d^n_{ij}, 1)\left(\dfrac{\pi}{n}\right) \right\rangle 
%        \left| y^1_j, \dots, y^n_j \right\rangle 
%        \left| 1 \right\rangle

Applying another Hadamard on ancillary qubit and doing necessary calculations the state of quantum register reads

\begin{multline}
    \left| \psi_4 \right\rangle = 
    \sum\limits_{i, j}^{k} 
    e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
		\\ \times
        \left\{\cos\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
        \left| y^1_j, \dots, y^n_j \right\rangle 
        \left| 0 \right\rangle\right.
        \\+ 
        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
        \left| y^1_j, \dots, y^n_j \right\rangle 
        \left| 1 \right\rangle\right\}
\end{multline}

Then we return to our initial basis for register $\left| X \right\rangle$ with applying the pairwise CNOT gates:

\begin{multline}
    \left| \psi_5 \right\rangle = 
    \mathrm{CNOT(Y,X)\left| \psi_4 \right\rangle} \\=  
    \sum\limits_{i, j}^{k} 
    e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
				\left\{\cos\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
        \left| X_i \right\rangle 
        \left| Y_j \right\rangle 
        \left| 0 \right\rangle\right.
        \\+
        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
        \left| X_i \right\rangle 
        \left| Y_j \right\rangle 
        \left| 1 \right\rangle\right\}
\end{multline}

In the result we have the Hamming distances encoded into the amplitudes of the final state. 
After getting statistics of the measurement outcomes we can produce the distance matrix between two data sets of binary vectors (Fig.~\ref{fig:distance_matrix}). 
In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 
and we have the inverse relationship when measurement result on the auxiliary qubit is 1. 
In order to reduce noise we average measurement results over different states of the auxiliary qubit as $a_0(x_i,y_j) + (1-a_1(x_i,y_j))$ 
where $a_0(x_i,y_j)$ and $a_1(x_i,y_j)$ are amplitudes for the measurement results $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ and $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ respectively. 
In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, 
depends on actual realization control phase gate in a quantum register on the hardware level), 
reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.

% Please , add comparison of complexities in number of gates for realization between our scheme and previously proposed schemes (one qubit and two qubit separately)  
% ??

%\begin{align}
%\\
%\end{align} 
%

%
%\begin{equation}
%\end{equation}
%(a black region indicates if the word occurs in the abstract)

%%%%%
\begin{figure}[t]
	\label{fig:vectorized_sample}
	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
	\caption{
		(a) Representation of the data set of abstracts with the bag-of-words model is shown. 
		Each abstract is represented by a binary vector with 9 elements. 
		The samples are sorted into groups (QML, MED, BIO), 3 papers for each tag.  
		(b) The Hamming distance between each vectorized abstract is shown with a number in the matrix. 
		The sets of abstracts are well separated.
	} 
\end{figure}
%%%%%

\section{Demonstration of QASOFM on the IBM Q Experience}

We now show results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
We implement unsupervised data clustering for three sets of paper abstracts from three different fields: quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}. 
Each set consists of three random papers that focus on one of following topics: ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO). 
Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}). 
This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms. 
We restricted our bag-of-word size to 9 of the most frequent words from the full ``bags-of-word''  due to limitations of the number of qubits. 
After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.  
We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, showing a successful self-organization (Fig.~\ref{fig:vectorized_sample}).

Classical SOFM \cite{kohonen1990}, as it has been explained earlier, can be oversimplified into two basic procedures: distance calculation between a one sample vector and all cluster vectors and shifting the closest cluster vector to the sample vector. 
In classical SOFM, the complexity of algorithm in the sense of number of distance calculations scales as $O(LMN)$, where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.  
In the QASOFM (Sec. \ref{sec:qasofm}), distance calculations are realized on a quantum device (i.e. the IBM Q Experience) with the use of circuit presented in Fig. \ref{fig:qcircuit}. 
This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates that are possible to realize on currently available quantum computing devices with a limited number of qubits. 
The circuit realized in such a manner that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation. 
The complexity of the quantum assisted SOFM then scales as $O(LN)$.

In order to check that our algorithm gives the correct results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig. \ref{fig:distance_matrix}.  
We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
The theoretical calculations and classical simulations show perfect agreement with each other. 
An example of the QASOFM learning process is given on Fig.~\ref{fig:convergence}. 
Initially, the cluster vectors were randomly chosen (see Fig.~\ref{fig:convergence}) and label of sample distribution is shown for the zeroth epoch in Fig.~\ref{fig:convergence}(a). 
In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix at once we use the standard initialization of QISKIT library. 
Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case) or 27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case). 
In order to realize the circuit for the distance calculation for the quantum case we used the IBM Q machine. 
After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector and shift the closest cluster vectors to the sample one. 
The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
The evolution of the labels presented on Fig. ~\ref{fig:convergence}(c).  
Good convergence is already observed in the fourth epoch.

%%%%%
\begin{figure}[t]
	\label{fig:convergence}
	\includegraphics[width=0.95\columnwidth]{convergence.png}
	\caption{
		(a) Initial random binary vectors of clusters are presented. We randomly sampled 3 binary vectors of size 9 as initial cluster vectors (labeled by 0,1,2) and it is shown how the words from the bag of words model presented in them. 
		(b) The result of applying our QASOFM implemented on the IBMQ "ibmq\_16\_melbourne" backend is shown on the bottom left figure. 
		Vectors mean cluster elements for BIO, MED, QML groups, respectively (up to down). 
		(c) The evolution of label distribution on each learning epoch is presented on the right figure. 
		Good convergence already is observed on the fourth epoch.
	} 
\end{figure}
%%%%%

\section{Discussions}
We have developed a quantum assisted SOFM and showed a proof-of-concept demonstration that it can be used to solve clustering problems in an unsupervised manner with potential to outperform classical counterpart in real life tasks. 
The procedure of solving such clustering problem requires calculating the distance many times in iterative way. 
We introduced an optimized circuit for Hamming distance calculations that can be implemented on currently available quantum computing devices with high fidelity. 
Our circuit could act as a distance-computing component of a classical SOFMs algorithm and in this way improve its performance. 
Due to wide use of classical SOFMs in different areas of modern research this can give opportunities for the use of QASOFM in practical applications in near term.

\section*{Acknowledgments}
The work is supported by the RFBR-NSFC collaborative program (Grant No. 18-57-53007) and the State assignment (N. 0089-2019-0002). I.D.L. acknowledges support from the Russian Foundation for Basic Research, grant No.19-32-80004 and support from the Foundation for the Advancement of Theoretical Physics and Mathematics “BASIS” No.19-1-5-130-1. T.B. is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (61571301, D1210036A); the NSFC Research Fund for International Young Scientists (11650110425, 11850410426); NYU-ECNU Institute of Physics at NYU Shanghai; the Science and Technology Commission of Shanghai Municipality (17ZR1443600); the China Science and Technology Exchange Center (NGA-16-001); and the NSFC-RFBR Collaborative grant (81811530112).

%\section*{Author Contributions}

%I.D.L. and A.N.P. performed the calculations. All authors discussed the results and wrote the paper.

\section*{Conflict of interest}

The authors declare no conflict of interest.

\section*{Keywords}
quantum self-organizing feature map; 
quantum machine learning; 
neural network; 
quantum unsupervised data clustering; 
Hamming distance.

\bibliographystyle{unsrt}
\bibliography{bibliography}

%\begin{thebibliography}{99}
%\end{thebibliography}
\end{document}

### 2020-04-20 manuscript.tex
I defined Hamming distance in terms of projected expectation values

@@
 This allows for a simple encoding of the classical information into a quantum register.
-Initially we have two registers each with $n$ qubits in that we encode our set of binary vectors defined as
+
+The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single auxiliary qubit $\left| a \right\rangle$.
+During entire processing, the $\left| Y \right\rangle$ is used to store the cluster states, then at the beginning and at the end of the processing the $\left| X \right\rangle$ register stores the input vectors, while during the processing it stores the differences between input vectors and cluster states.
+
+Now for this demonstration let us assume we have $k$ input vectors and $l$ cluster states.
+The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$ ($\left| y_j \right\rangle$.
+
+The registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ will properly store input vectors and cluster vectors if initialized as follows.
+
 %
 \begin{align}
+    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
+    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle,
     \label{eq:encodnig}
-    \left| X \right\rangle  & = \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
-    \left| Y \right\rangle&  = \sum\limits_{i=1}^{l} \left| y_i \right\rangle,
 \end{align}
 % 
-where $ 1 \le k, l \le n $.  
-Here the registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ correspond to the input states and the cluster states respectively. 
-The initial state of whole circuit is
+Both registers along with the auxiliary qubit compose into initial state of the entire circuit in a form denoted below. 
 %
@@
 
-We then apply a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
+Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
 
@@
     \mathrm{CNOT(Y,X)\left| \psi_0 \right\rangle} =  
-    \sum\limits_{i, j}^{k} 
+    \frac{1}{\sqrt{kl}} \sum\limits_{i, j}^{k} 
     \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
@@
 
-In the result, we encoded information about pairwise different qubits in $\{X\}$ and $\{Y\}$ into register $\left| X \right\rangle$.
+In the result, at this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors, instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$.
 
-After that we transfer information about distance between sample state and cluster state into the amplitude of the superposed state. 
-In order to make that we apply Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,
+The next stage, for each pair $\{X\}$ and $\{Y\}$, an accumulated information of all the differences is projected into the amplitude of the superposed state.
+this is achieved by applying the Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,
 
@@
 
-Then we return to our initial basis for register $\left| X \right\rangle$ with applying the pairwise CNOT gates:
+This completed the stage of projecting the differences between pairs of $\{X\}$ and $\{Y\}$ onto the amplitude of the auxiliary qubit.
+The process is done in the $x$-basis, which explains the surrounding Hadamard gates.
+There are two possible measurement outcomes of the auxiliary qubit, each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
+
+At this stage, the information of differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, thus we return to our initial basis for register $\left| X \right\rangle$ by applying the pairwise CNOT gates:
 
 \begin{multline}
-    \left| \psi_5 \right\rangle = 
+    \left| \psi_f \right\rangle = 
     \mathrm{CNOT(Y,X)\left| \psi_4 \right\rangle} \\=  
@@
 
+This made $\{X\}$ store input vectors again, just like at the initial stage, however it preserved the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
+
 In the result we have the Hamming distances encoded into the amplitudes of the final state.
 After getting statistics of the measurement outcomes we can produce the distance matrix between two data sets of binary vectors (Fig.~\ref{fig:distance_matrix}).
-In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 
-and we have the inverse relationship when measurement result on the auxiliary qubit is 1. 
-In order to reduce noise we average measurement results over different states of the auxiliary qubit as $a_0(x_i,y_j) + (1-a_1(x_i,y_j))$ 
-where $a_0(x_i,y_j)$ and $a_1(x_i,y_j)$ are amplitudes for the measurement results $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ and $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ respectively.
-In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
-our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, 
-depends on actual realization control phase gate in a quantum register on the hardware level), 
-reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.
+In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 and we have the inverse relationship when measurement result on the auxiliary qubit is 1.
+
+Measuring the Hamming distance of a particular pair of input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form, this can be done using the following projector operator.
+
+\begin{align}
+\Pi_{i,j} = &\left| X_i \rangle\langle X_i \right| \otimes \left| X_i \rangle\langle Y_j \right| \otimes I \nonumber \\ 
+\end{align}
+
+Using the above projection operator, the subspace of the Hilbert space formed by particular pair of input vector and cluster vector can be traced out as follows.
+
+\begin{align}
+    \rho_{i,j} &= \text{Tr}_{1,\dots,2n} (\Pi_{i,j} \left| \psi_f \rangle\langle \psi_f \right| \Pi_{i,j})
+\end{align}
+
+Now the following two relevant amplitudes for the measurement results can be extracted.
+
+\begin{align}
+    a_0(x_i,y_j) & = \frac{\left\langle 0 |\rho_{i,j}| 0 \right\rangle}{\text{tr}(\rho_{i,j})},  \\
+    a_1(x_i,y_j) & = \frac{\left\langle 1 |\rho_{i,j}| 1 \right\rangle}{\text{tr}(\rho_{i,j})},
+\end{align}
+
+In order to reduce noise we average measurement results over different states of the auxiliary qubit, thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ takes the following form.
+
+\begin{align}
+    d_{i,j}^H & \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j)))
+\end{align}
+
+The Hamming distance measured in this way is contained $0 \leq d_{i,j}^H \leq 1$, where intuitively, $0$ would indicate $x_i$ are $y_j$ identical and $1$ would mean they are completely opposite in terms of their pairwise binary coordinates.
+
+In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, depends on actual realization control phase gate in a quantum register on the hardware level), reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.
 

### 2020-04-22 manuscript.tex
fixed bug in last sentence of third paragraph section 2A
Fix style: one line one sentence

@@
 The larger the difference between the input component and the weight component, the larger increment (decrement). 
-Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects.
+Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way 
+that ensures each move is following the current trends inferred from their distances to the input objects. 
 A visualisation of this process is provided on figure (Fig.~\ref{fig:sofm_fitting}).
@@
 In our approach, we use the Hamming distance instead of the Euclidean distance to select the winning cluster vector. 
-It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
+It allows us to use a simpler encoding of the classical information into the quantum state and 
+use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
 such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
@@
 The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single auxiliary qubit $\left| a \right\rangle$. 
-During entire processing, the $\left| Y \right\rangle$ is used to store the cluster states, then at the beginning and at the end of the processing the $\left| X \right\rangle$ register stores the input vectors, while during the processing it stores the differences between input vectors and cluster states.
+During entire processing, the $\left| Y \right\rangle$ is used to store the cluster states, 
+then at the beginning and at the end of the processing the $\left| X \right\rangle$ register stores the input vectors, 
+while during the processing it stores the differences between input vectors and cluster states.
 
 Now for this demonstration let us assume we have $k$ input vectors and $l$ cluster states. 
-The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$ ($\left| y_j \right\rangle$.
+The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$.
 
@@
 
-Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
+Given this initial state we may begin the processing of the problem. 
+We start by applying a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
 
@@
 
-In the result, at this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors, instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$.
+In the result, at this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors, 
+instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$.
 
 The next stage, for each pair $\{X\}$ and $\{Y\}$, an accumulated information of all the differences is projected into the amplitude of the superposed state. 
-this is achieved by applying the Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,
+This is achieved by applying the Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,
 
@@
     \left| \psi_2 \right\rangle = H_a\left| \psi_1 \right\rangle = 
-    \sum\limits_{i, j}^{k} 
+    \frac{1}{\sqrt{kl}} \sum\limits_{i, j}^{k} 
     \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
@@
     \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
-    \\ = \dfrac{1}{\sqrt{2}}
+    \\ = \dfrac{1}{\sqrt{2kl}}
 				\sum\limits_{i, j}^{k} 
@@
         \left| 0 \right\rangle
-        \\ + \dfrac{1}{\sqrt{2}}
+        \\ + \dfrac{1}{\sqrt{2kl}}
 				\sum\limits_{i, j}^{k}
@@
 
-%\\ = \dfrac{1}{\sqrt{2}}
-%    \sum\limits_{i, j}^{k}
-%    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-%        \left| y^1_j, \dots, y^n_j \right\rangle 
-%        \left| 0 \right\rangle
-%        \\ + \dfrac{1}{\sqrt{2}}
-%				\sum\limits_{i, j}^{k}
-%        \left| R(d^1_{ij}, 1)\left(\dfrac{\pi}{n}\right), \dots, R(d^n_{ij}, 1)\left(\dfrac{\pi}{n}\right) \right\rangle 
-%        \left| y^1_j, \dots, y^n_j \right\rangle 
-%        \left| 1 \right\rangle
-
 Applying another Hadamard on ancillary qubit and doing necessary calculations the state of quantum register reads
@@
     \left| \psi_4 \right\rangle = 
-    \sum\limits_{i, j}^{k} 
+    \frac{1}{\sqrt{kl}} \sum\limits_{i, j}^{k} 
     e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
@@
 The process is done in the $x$-basis, which explains the surrounding Hadamard gates. 
-There are two possible measurement outcomes of the auxiliary qubit, each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
+There are two possible measurement outcomes of the auxiliary qubit, each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, 
+the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
 
-At this stage, the information of differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, thus we return to our initial basis for register $\left| X \right\rangle$ by applying the pairwise CNOT gates:
+At this stage, the information of differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, 
+thus we return to our initial basis for register $\left| X \right\rangle$ by applying the pairwise CNOT gates:
 
@@
     \mathrm{CNOT(Y,X)\left| \psi_4 \right\rangle} \\=  
-    \sum\limits_{i, j}^{k} 
+    \frac{1}{\sqrt{kl}}\sum\limits_{i, j}^{k} 
     e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
@@
 
-This made $\{X\}$ store input vectors again, just like at the initial stage, however it preserved the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
+This made $\{X\}$ store input vectors again, just like at the initial stage, 
+however it preserved the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
 
@@
 After getting statistics of the measurement outcomes we can produce the distance matrix between two data sets of binary vectors (Fig.~\ref{fig:distance_matrix}). 
-In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 and we have the inverse relationship when measurement result on the auxiliary qubit is 1.
+In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 and 
+we have the inverse relationship when measurement result on the auxiliary qubit is 1.
 
@@
 
-In order to reduce noise we average measurement results over different states of the auxiliary qubit, thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ takes the following form.
+In order to reduce noise we average measurement results over different states of the auxiliary qubit, 
+thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ takes the following form.
 
@@
 
-The Hamming distance measured in this way is contained $0 \leq d_{i,j}^H \leq 1$, where intuitively, $0$ would indicate $x_i$ are $y_j$ identical and $1$ would mean they are completely opposite in terms of their pairwise binary coordinates.
+The Hamming distance measured in this way is contained $0 \leq d_{i,j}^H \leq 1$, 
+where intuitively, $0$ would indicate $x_i$ are $y_j$ identical and 
+$1$ would mean they are completely opposite in terms of their pairwise binary coordinates.
 
-In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, depends on actual realization control phase gate in a quantum register on the hardware level), reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.
-
-% Please , add comparison of complexities in number of gates for realization between our scheme and previously proposed schemes (one qubit and two qubit separately)  
-% ??
-
-%\begin{align}
-%\\
-%\end{align} 
-%
-
-%
-%\begin{equation}
-%\end{equation}
-%(a black region indicates if the word occurs in the abstract)
+In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
+our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, 
+depends on actual realization control phase gate in a quantum register on the hardware level), 
+reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.
 
@@
 We now show results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
-We implement unsupervised data clustering for three sets of paper abstracts from three different fields: quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}.
+We implement unsupervised data clustering for three sets of paper abstracts from three different fields: 
+quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}. 
 Each set consists of three random papers that focus on one of following topics: ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO). 
@@
 This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
-Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms.
+Preparing the bag-of-words we excluded the words that appear only in one abstract and 
+more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters 
+because it gives instabilities for both classical and quantum algorithms. 
 We restricted our bag-of-word size to 9 of the most frequent words from the full ``bags-of-word''  due to limitations of the number of qubits. 
 After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.  
-We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, showing a successful self-organization (Fig.~\ref{fig:vectorized_sample}).
+We observe that distances between the abstracts inside clusters are smaller 
+than distances between the abstracts from different clusters, showing a successful self-organization (Fig.~\ref{fig:vectorized_sample}).
 
-Classical SOFM \cite{kohonen1990}, as it has been explained earlier, can be oversimplified into two basic procedures: distance calculation between a one sample vector and all cluster vectors and shifting the closest cluster vector to the sample vector.
-In classical SOFM, the complexity of algorithm in the sense of number of distance calculations scales as $O(LMN)$, where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.
+Classical SOFM \cite{kohonen1990}, as it has been explained earlier, can be oversimplified into two basic procedures: 
+distance calculation between a one sample vector and all cluster vectors and shifting the closest cluster vector to the sample vector. 
+In classical SOFM, the complexity of algorithm in the sense of number of distance calculations scales as $O(LMN)$, 
+where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.  
 In the QASOFM (Sec. \ref{sec:qasofm}), distance calculations are realized on a quantum device (i.e. the IBM Q Experience) with the use of circuit presented in Fig. \ref{fig:qcircuit}. 
-This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates that are possible to realize on currently available quantum computing devices with a limited number of qubits.
+This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates 
+that are possible to realize on currently available quantum computing devices with a limited number of qubits. 
 The circuit realized in such a manner that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation. 
@@
 
-In order to check that our algorithm gives the correct results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.
-\ref{fig:distance_matrix}.
+In order to check that our algorithm gives the correct results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, 
+as shown in Fig. \ref{fig:distance_matrix}.  
 We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
@@
 An example of the QASOFM learning process is given on Fig.~\ref{fig:convergence}. 
-Initially, the cluster vectors were randomly chosen (see Fig.~\ref{fig:convergence}) and label of sample distribution is shown for the zeroth epoch in Fig.~\ref{fig:convergence}(a). 
+Initially, the cluster vectors were randomly chosen (see Fig.~\ref{fig:convergence}) and 
+label of sample distribution is shown for the zeroth epoch in Fig.~\ref{fig:convergence}(a). 
 In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix at once we use the standard initialization of QISKIT library. 
-Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case) or 27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case).
+Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case) or 
+27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case). 
 In order to realize the circuit for the distance calculation for the quantum case we used the IBM Q machine. 
-After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector and shift the closest cluster vectors to the sample one.
+After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector 
+and shift the closest cluster vectors to the sample one. 
 The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
@@
 \section{Discussions}
-We have developed a quantum assisted SOFM and showed a proof-of-concept demonstration that it can be used to solve clustering problems in an unsupervised manner with potential to outperform classical counterpart in real life tasks.
+We have developed a quantum assisted SOFM and showed a proof-of-concept demonstration that 
+it can be used to solve clustering problems in an unsupervised manner with potential to outperform classical counterpart in real life tasks. 
 The procedure of solving such clustering problem requires calculating the distance many times in iterative way. 

### 2020-04-23 manuscript.tex
Tims remarks

--- 2020-04-20 manuscript.tex
@@
 \usepackage{soul}
-% \usepackage{floatflt}
+%\usepackage{floatflt}
 \usepackage{graphicx}
@@
 \begin{document}
-\title{Quantum assisted unsupervised data clustering on the basis of neural networks}
+\title{Quantum-assisted unsupervised data clustering}
 
@@
 \begin{abstract}
-	Unsupervised machine learning is one of the main directions in the area of artificial intelligence. 
-	Here, we propose a realization of quantum assisted unsupervised data clustering on the basis of artificial neural network called a self-organizing feature map. 
-	We make a proof-of-concept realization of one of the crucial parts of the approach on IBM quantum machine and show that it allows us to reduce number of calculations in a number of clusters. 
-	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.  
+Unsupervised machine learning is one of the main techniques in artificial intelligence.
+Quantum computers offer opportunities to speed up such machine learning techniques, where the number of input data elements is extremely large.
+Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map artificial neural networks.
+We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience and show that it allows us to reduce number of calculations in a number of clusters.
+[TB: Explain what the main technique that is used is.
+This is basically a QAOA hybrid quantum-classical approach right?
+If so, you should explain this] We compare the results with the classical algorithm on a toy example of unsupervised text clustering.
 \end{abstract}
@@
 It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
-In particularly, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) that is trained in an unsupervised manner. 
+
+In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) that is trained in an unsupervised manner.
 SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
@@
 Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}.
-In the early stage, the main focus in QML was on the use of quantum computers to perform some basic linear algebra subroutines faster (such as Fourier transform and solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}),
-open up the possibilities for exponential speed-ups utilizing quantum phenomena such as quantum entanglement and quantum superposition over classical counterparts \cite{biamonte2017, schuld2014} and,
-in the result, quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and so on \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
-Recently,  a number of shifts occurred and the hottest trend migrated to developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019}. 
-The shift was inspired with the progress in experimental quantum computing when it became possible to play around with parametrized quantum circuits, whose parameters behave much like the weights of a neural network \cite{lewenstein1994}.
-In particular, quantum algorithms for training and evaluating feedforward neural networks, one of the most usable neural network model, was developed \cite{allcock2018, tacchino2019}. 
-Recent results connected classical Bayesian approach to deep learning allowed to develop a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}. 
+The main focus of early works in quantum machine learning (QML) was on the use of quantum computers to perform basic linear algebra subroutines with a quantum speedup (such as Fourier transform and solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}) by utilizing quantum phenomena such as quantum entanglement and quantum superposition \cite{biamonte2017, schuld2014}.
+Quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
+More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019}.
+The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits, where the parameters behave much like the weights of a neural network \cite{lewenstein1994}.
+In particular, quantum algorithms for training and evaluating feedforward neural networks was developed, which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
+Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}.
 The quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}.
-Quantum classification, tested on the MNIST dataset, via Slow Feature Analysis based on the use of Quantum Frobenius Distance was proposed \cite{kerenidis2018}.
+Quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
 Sublinear quantum algorithms for training linear and kernel-based classifiers were designed \cite{li2019}.
 Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, quantum autoencoders \cite{bondarenko2019}, quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, vinci2019, lu2019} were developed.
-Currently, it is believed that implementation of quantum neural networks can be the main test bed to achieve practical quantum supremacy on NISQ devices.
+Currently, it is believed that implementation of quantum neural networks can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.
 
-In this paper, we realize a hybrid quantum assisted SOFM (QASOFM) and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.   
+[TB: I think we definitely need to cite and describe hybrid quantum-classical approaches, like QAOA.  Could you please add several sentences and a solid set of references for these]
+
+In this paper, we analyze  a hybrid quantum assisted SOFM (QASOFM) and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.
 We implement the quantum assisted SOFM in such a way that it becomes possible to use the Hamming distance as a distance metric for the training and to reduce the number of distance calculations in the number of clusters.
 We propose an optimized circuit for realizing the Hamming distance on a quantum machine.
-We prepare a toy example of data for clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} and compare it with the classical case.
+We then apply it to a toy example of data for clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} and compare it to the classical case.
 
 \begin{figure}
-	\label{fig:sofm_fitting}
-	\includegraphics[width=0.95\columnwidth]{sofm_fitting.png}
-	\caption{
-		Schematic illustration of the clustering problem considered in this paper.  
+  \includegraphics[width=0.95\columnwidth]{clusters.png}
+\caption{Schematic illustration of the clustering problem considered in this paper.
 Blue dots represent clusters and red dots are data points.
 The training process moves clusters to fit the data points.
-		Note that there are fewer clusters than data points which is the essence of dimensionality reduction and is what permits the model to generalize the data.
-	}
+Note that there are fewer clusters than data points, which is the essence of dimensionality reduction, and is what permits the model to generalize the data.}
+\label{clusters}
 \end{figure}
@@
 
-SOFM is one of the best-known unsupervised learning methods that is widely used in various areas of modern science. 
+\subsection{The classical algorithm}
+
+The SOFM is one of the most widely-used unsupervised learning methods used in various areas of modern science.
 It was first proposed by Kohonen as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}.
-The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, as input and maps them onto nodes of a two-dimensional grid.
+The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, and maps them onto nodes of a two-dimensional grid.
 
-The input dimensions are associated with features and nodes in the grid, called cluster vectors, are assigned the $N$-dimensional vectors; 
-the components of these vectors are usually called weights. 
-Initially weight components are chosen randomly. 
-We then can train our SOFM adjusting the components through the learning process which can be oversimplified into two basic procedures, 
-selecting a winning cluster vector and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
-In more detail, they consist of four step process: 
-1) select an input vector randomly from the set of all input vectors; 
-2) find a cluster vector which is closest to the input vector; 
-3) adjust the weights of the winning node in such a way that it becomes even closer to the input vector; 
-4) repeat this process many iterations until it converges.
+The input dimensions are associated with the features, and the nodes in the grid (called cluster vectors) are assigned the $N$-dimensional vectors.
+The components of these vectors are usually called weights.
+Initially the weight components are chosen randomly.
+We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of selecting a winning cluster vector and updating its weights (Fig.~\ref{clusters}).
+More specifically, they consist of four step process: 1) selecting an input vector randomly from the set of all input vectors; 2) finding a cluster vector which is closest to the input vector; 3) adjusting the weights of the winning node in such a way that it becomes even closer to the input vector; 4) repeating this process for many iterations until it converges.
 
-After the winning cluster vector is selected, the weights of the vector are adjusted according to 
+After the winning cluster vector is selected, the weights of the vector are adjusted according to [TB: all variables should be defined: what is $ \vec{x} , \alpha $ etc.
+] [TB: this is done iteratively?
+Then should there be an index over  $ \vec{x}_i $?
+]
 %
 \begin{equation}
+    \vec w_{\mathrm{new}} =
+        \vec w_{\mathrm{old}} 
+        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  .
     \label{eq:learning}
-    \vec w_{\mathrm{new}} =
-		\vec{w}_{\mathrm{old}} 
-        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  .
 \end{equation}
 %
-In simple words, this expression can be understood according to the following: 
-if a component of the input vector is greater than the corresponding weight, increase the weight by a small amount; 
-if the input component is smaller than the weight, decrease the
+The above expression can be interpreted according to: if a component of the input vector is greater than the corresponding weight, increase the weight by a small amount; if the input component is smaller than the weight, decrease the
 weight by a small amount.
-The larger the difference between the input component and the weight component, the larger increment (decrement). 
+The larger the difference between the input component and the weight component, the larger the increment (decrement).
 Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects.
-A visualisation of this process is provided on figure (Fig.~\ref{fig:sofm_fitting}).
+A visualisation of this process is shown in Fig.
+\ref{clusters}.
 
@@
 In our approach, we use the Hamming distance instead of the Euclidean distance to select the winning cluster vector.
-It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
-such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
+It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
 
@@
 \begin{figure}[t]
-	\label{fig:qcircuit}
-	\includegraphics[width=0.95\columnwidth]{qcircuit.png}
+\includegraphics[width=0.95\columnwidth]{circuit.png}
 \caption{
-    	%add the caption with intuitive explanation of the circuit
-		A quantum circuit for the Hamming distance calculation algorithm between all samples and clusters at once. 
-		First, we apply Hadamard gates in order to get superposition of the sample and cluster registers. 
+    Quantum circuit for the quantum parallelized Hamming distance calculation between all samples and clusters.
+First,  Hadamard gates are applied in order to obtain a superposition of the sample and cluster registers.
 Second, we encoded information about pairwise different qubits in a quantum state of the sample register with applying the CNOT gates.
@@
 Finally, a quantum state of the sample register returned to the initial basis for information retrieval.
+[TB: The figure is too small to be read, please make this bigger]
 } 
+\label{circuit}
 \end{figure}
@@
 \begin{figure}[t]
+\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
+\caption{Hamming distance matrix between two data sets of binary vectors.
+Values of distance that are less than the median distance value are marked black.
+Classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
+Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend is shown on the right figure.
+We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
+[TB: could you put (a) (b) labels as with the other figures.
+]}
 \label{fig:distance_matrix}
-	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
-	\caption{
-		The Hamming distance matrix between two data sets of binary vectors. 
-		Values of distance that less than median distance value was marked black. 
-		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure. 
-		Result obtained on  IBMQ "ibmq\_16\_melbourne" backend is shown on the right figure. 
-		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
-	} 
 \end{figure}
@@
 
-We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors at once.  
+We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors, making use of quantum parallelism.
 This allows for a simple encoding of the classical information into a quantum register.
-Initially we have two registers each with $n$ qubits in that we encode our set of binary vectors defined as
+[TB: allows?]
+
+The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single auxiliary qubit $\left| a \right\rangle$.
+During the whole process, the $\left| Y \right\rangle$ register is used to store the cluster states.
+At the beginning and end of the procedure the $\left| X \right\rangle$ register stores the input vectors.
+Meanwhile, during the procedure it stores the differences between input vectors and cluster states.
+
+Let us assume we have $k$ input vectors and $l$ cluster states.
+The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$, $\left| y_j \right\rangle$.
+The registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ are initialized to store the input vectors and cluster vectors according to
 %
 \begin{align}
+    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
+    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
     \label{eq:encodnig}
-    \left| X \right\rangle  & = \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
-    \left| Y \right\rangle&  = \sum\limits_{i=1}^{l} \left| y_i \right\rangle,
 \end{align}
 % 
-where $ 1 \le k, l \le n $.  
-Here the registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ correspond to the input states and the cluster states respectively. 
-The initial state of whole circuit is
+The two registers along with the auxiliary qubit comprise the initial state of the quantum computer according to
 %
@@
 
-We then apply a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
+Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
 
-\begin{equation}
-    \left| \psi_1 \right\rangle = 
-    \mathrm{CNOT(Y,X)\left| \psi_0 \right\rangle} =  
-    \sum\limits_{i, j}^{k} 
-    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-    \left| y^1_j, \dots, y^n_j \right\rangle
-    \left| 0 \right\rangle 
-\end{equation}
-where $d^\alpha_{ij} = \mathrm{CNOT}(y^\alpha_i, x^\alpha_j),\, \alpha = 1 \dots n$ and $i,j$ are the qubit indexes in the registers.
-
-In the result, we encoded information about pairwise different qubits in $\{X\}$ and $\{Y\}$ into register $\left| X \right\rangle$.
-
-After that we transfer information about distance between sample state and cluster state into the amplitude of the superposed state. 
-In order to make that we apply Hadamard gate on auxiliary qubit, then control phase gate on $\left| Xa \right\rangle$,
-
+\begin{align}
+    | \psi_1 \rangle & = 
+    \mathrm{CNOT} (Y,X)| \psi_0 \rangle \nonumber \\
+		& =  
+    \frac{1}{\sqrt{kl}} \sum_{i, j=1}^{k} 
+    | d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \rangle 
+    | y^{(1)}_j, \dots, y^{(n)}_j \rangle
+    | 0 \rangle 
+\end{align}
+%
+where $d^\alpha_{ij} = \mathrm{CNOT}(y^\alpha_i, x^\alpha_j)$, $\alpha = 1 \dots n$, and $i,j$ are the qubit indexes in the registers.
+At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors, instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$.
+Next, for each pair $\{X\}$ and $\{Y\}$, the accumulated information of all the differences is projected onto the amplitude of the superposed state.
+This is achieved by applying the Hadamard gate on auxiliary qubit, followed by a controlled phase gate on $\left| Xa \right\rangle$ defined as
+%
 \begin{equation}
@@
         0 & 0 & 0 & e^{-i\phi}
-    \end{pmatrix},
+    \end{pmatrix} .
     \quad \phi = \frac{\pi}{n}
 \end{equation}
-and again Hadamard gate.
+%
+Finally, another Hadamard gate is applied. [TB: on which qubit? I think its better just to describe this using a circuit, rather than describe the steps in words. ]
 
-After the first Hadamard on the ancillary qubit we have the following state
-
+After the first Hadamard on the ancilla qubit the state is
+%
 \begin{equation}
     \left| \psi_2 \right\rangle = H_a\left| \psi_1 \right\rangle = 
-    \sum\limits_{i, j}^{k} 
-    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-    \left| y^1_j, \dots, y^n_j \right\rangle
-    \dfrac{\left| 0 \right\rangle + \left| 1 \right\rangle}{\sqrt{2}}
+    \sum\limits_{i, j=1}^{k} 
+    \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+    \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle
+    \dfrac{(\left| 0 \right\rangle + \left| 1 \right\rangle)}{\sqrt{2}}  .
 \end{equation}
-
-Then we apply the control phase gate
-
+%
+Applying the controlled phase gate the state then becomes [TB: please define all variables like $ R_{(X,a)} $]
+%
 \begin{multline}
@@
     \\ = \dfrac{1}{\sqrt{2}}
-				\sum\limits_{i, j}^{k} 
-				\left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-        \left| y^1_j, \dots, y^n_j \right\rangle 
+				\sum\limits_{i, j=1}^{k} 
+				\left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
         \left| 0 \right\rangle
         \\ + \dfrac{1}{\sqrt{2}}
-				\sum\limits_{i, j}^{k}
-        \exp\left(\dfrac{-i \pi}{n}\sum\limits_l^n d^l_{ij} \right)
-        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-        \left| y^1_j, \dots, y^n_j \right\rangle 
+				\sum\limits_{i, j=1}^{k}
+        \exp\left(\dfrac{-i \pi}{n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+\\ \times        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
         \left| 1 \right\rangle
 \end{multline}
-
+%
+%
 %\\ = \dfrac{1}{\sqrt{2}}
 %    \sum\limits_{i, j}^{k}
-%    \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-%        \left| y^1_j, \dots, y^n_j \right\rangle 
+%    \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+%        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
 %        \left| 0 \right\rangle
@@
 %				\sum\limits_{i, j}^{k}
-%        \left| R(d^1_{ij}, 1)\left(\dfrac{\pi}{n}\right), \dots, R(d^n_{ij}, 1)\left(\dfrac{\pi}{n}\right) \right\rangle 
-%        \left| y^1_j, \dots, y^n_j \right\rangle 
+%        \left| R(d^{(1)}_{ij}, 1)\left(\dfrac{\pi}{n}\right), \dots, R(d^{(n)}_{ij}, 1)\left(\dfrac{\pi}{n}\right) \right\rangle 
+%        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
 %        \left| 1 \right\rangle 
-
-Applying another Hadamard on ancillary qubit and doing necessary calculations the state of quantum register reads
-
+%
+Applying another Hadamard on the ancilla qubit we obtain
+%
 \begin{multline}
     \left| \psi_4 \right\rangle = 
-    \sum\limits_{i, j}^{k} 
-    e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
+    \sum\limits_{i, j=1}^{k} 
+    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
 		\\ \times
-        \left\{\cos\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
-        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-        \left| y^1_j, \dots, y^n_j \right\rangle 
+        \left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
         \left| 0 \right\rangle\right.
         \\+ 
-        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
-        \left| d^1_{ij}, \dots, d^n_{ij} \right\rangle 
-        \left| y^1_j, \dots, y^n_j \right\rangle 
-        \left| 1 \right\rangle\right\}
+        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
+        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
+        \left| 1 \right\rangle\right] .
 \end{multline}
+%
+This completes the step for projecting differences between pairs of $\{X\}$ and $\{Y\}$ onto the amplitude of the auxiliary qubit.
+The process is done in the $x$-basis, achieved by the surrounding Hadamard gates.
+There are two possible measurement outcomes of the auxiliary qubit.
+Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
 
-Then we return to our initial basis for register $\left| X \right\rangle$ with applying the pairwise CNOT gates:
-
+At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, thus we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+%
 \begin{multline}
-    \left| \psi_5 \right\rangle = 
-    \mathrm{CNOT(Y,X)\left| \psi_4 \right\rangle} \\=  
-    \sum\limits_{i, j}^{k} 
-    e^{\left(\dfrac{-i \pi}{2n}\sum\limits_l^n d^l_{ij} \right)}
-				\left\{\cos\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
+    \left| \psi_f \right\rangle = 
+    \mathrm{CNOT} (Y,X)\left| \psi_4 \right\rangle \\=  
+    \sum\limits_{i, j=1}^{k} 
+    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+				\left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| X_i \right\rangle 
@@
         \\+
-        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_l^n d^l_{ij} \right)
+        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| X_i \right\rangle 
         \left| Y_j \right\rangle 
-        \left| 1 \right\rangle\right\}
+        \left| 1 \right\rangle\right] .
 \end{multline}
+%
+This makes $\{X\}$ store the input vectors again, as in the initial step.
+It however preserves the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
 
-In the result we have the Hamming distances encoded into the amplitudes of the final state. 
-After getting statistics of the measurement outcomes we can produce the distance matrix between two data sets of binary vectors (Fig.~\ref{fig:distance_matrix}). 
-In this case, a biggest amplitude in measurement result coincide to a smallest Hamming distance when the measurement result of the auxiliary qubit is 0 
-and we have the inverse relationship when measurement result on the auxiliary qubit is 1. 
-In order to reduce noise we average measurement results over different states of the auxiliary qubit as $a_0(x_i,y_j) + (1-a_1(x_i,y_j))$ 
-where $a_0(x_i,y_j)$ and $a_1(x_i,y_j)$ are amplitudes for the measurement results $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ and $\left| x_i \right\rangle\left| y_j \right\rangle \left| 0 \right\rangle_a$ respectively.
-In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
-our method allows one to reduce number of gates in the circuit (at least in $2n$ one-qubit NOT gates, 
-depends on actual realization control phase gate in a quantum register on the hardware level), 
-reconstruct all distance matrix at once and can be implemented with high fidelity on current quantum devices.
+We thus have the Hamming distances encoded into the amplitudes of the final state.
+From the statistics of the measurement outcomes, the distance matrix between two data sets of binary vectors can be obtained (Fig.~\ref{fig:distance_matrix}).
+In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0.
+If the ancilla qubit is 1, there is an inverse relationship with the measurement result.  [TB: what kind of inverse relationship?]
+
+Measuring the Hamming distance of a particular pair of input vectors $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form, this can be done using the following projection operator
+%
+\begin{align}
+\Pi_{i,j} = &\left| X_i \rangle\langle X_i \right| \otimes \left| X_i \rangle\langle Y_j \right| \otimes I .
+\end{align} 
+%
+Using the above projection operator, the subspace of the Hilbert space formed by a particular pair of input and cluster vectors can be traced out as
+%
+\begin{align}
+    \rho_{i,j} &= \text{Tr}_{1,\dots,2n} (\Pi_{i,j} \left| \psi_f \rangle\langle \psi_f \right| \Pi_{i,j}) .
+\end{align}
+%
+From the state the following two amplitudes for the measurement results can be extracted
+%
+\begin{align}
+    a_0(x_i,y_j) & = \frac{\left\langle 0 |\rho_{i,j}| 0 \right\rangle}{\text{Tr}(\rho_{i,j})}  \\
+    a_1(x_i,y_j) & = \frac{\left\langle 1 |\rho_{i,j}| 1 \right\rangle}{\text{Tr}(\rho_{i,j})} .
+\end{align}
+%
+In order to reduce noise we average the measurement results over different states of the ancilla qubit, thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ is
+%
+\begin{align}
+    d_{i,j}^H & \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))) .
+\end{align}
+%
+The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$, where  $0$ indicates that $x_i$ are $y_j$ identical and $1$ means they are completely opposite in terms of their pairwise binary coordinates.
+
+In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, our method allows one to reduce the number of gates in the circuit and reconstruct all distance matrix in parallel and can be implemented with high fidelity.
+We note that at least for the $2n$ one-qubit NOT gates, this depends on the actual realization of the control phase gate in a quantum register at the hardware level.
+[TB: don't understand this sentence]
 
@@
 \begin{figure}[t]
+\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
+\caption{(a) Representation of the data set of abstracts with the bag-of-words model is shown.
+Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
+The samples are sorted into groups (QML, MED, BIO) with 3 papers for each tag, for a total of 9 paper.
+(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
+} 
 \label{fig:vectorized_sample}
-	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
-	\caption{
-		(a) Representation of the data set of abstracts with the bag-of-words model is shown. 
-		Each abstract is represented by a binary vector with 9 elements. 
-		The samples are sorted into groups (QML, MED, BIO), 3 papers for each tag.  
-		(b) The Hamming distance between each vectorized abstract is shown with a number in the matrix. 
-		The sets of abstracts are well separated.
-	} 
 \end{figure}
@@
 
-\section{Demonstration of QASOFM on the IBM Q Experience}
+\section{Expeirmental demonstration of QASOFM}
 
-We now show results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
-We implement unsupervised data clustering for three sets of paper abstracts from three different fields: quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}.
+We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
+We perform unsupervised data clustering for three sets of paper abstracts from three different fields: quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}.
 Each set consists of three random papers that focus on one of following topics: ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO).
 Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}).
-This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
+[TB: bag of words reference?]  This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words.
 Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms.
-We restricted our bag-of-word size to 9 of the most frequent words from the full ``bags-of-word''  due to limitations of the number of qubits. 
+We restricted our bag-of-word size to 9 of the most frequent words from the full bags-of-word  due to limitations of the number of qubits.
 After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.
-We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, showing a successful self-organization (Fig.~\ref{fig:vectorized_sample}).
+We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, showing successful self-organization (Fig.~\ref{fig:vectorized_sample}).
 
-Classical SOFM \cite{kohonen1990}, as it has been explained earlier, can be oversimplified into two basic procedures: distance calculation between a one sample vector and all cluster vectors and shifting the closest cluster vector to the sample vector.
-In classical SOFM, the complexity of algorithm in the sense of number of distance calculations scales as $O(LMN)$, where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.
-In the QASOFM (Sec. \ref{sec:qasofm}), distance calculations are realized on a quantum device (i.e. the IBM Q Experience) with the use of circuit presented in Fig. \ref{fig:qcircuit}. 
-This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates that are possible to realize on currently available quantum computing devices with a limited number of qubits.
+In classical SOFM \cite{kohonen1990}, first a distance calculation between a a sample vector and all cluster vectors is made, then the closest cluster vector is shifted towards the sample vector.
+The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$, where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.
+In the QASOFM described in the previous section, distance calculations are realized on a quantum device (i.e.
+the IBM Q Experience) with the use of circuit presented in Fig.
+\ref{circuit}.
+This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates that are possible to realize on  quantum computing devices that are currently available.
 The circuit realized in such a manner that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation.
@@
 
-In order to check that our algorithm gives the correct results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.
+In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.
 \ref{fig:distance_matrix}.
@@
 The theoretical calculations and classical simulations show perfect agreement with each other.
-An example of the QASOFM learning process is given on Fig.~\ref{fig:convergence}. 
-Initially, the cluster vectors were randomly chosen (see Fig.~\ref{fig:convergence}) and label of sample distribution is shown for the zeroth epoch in Fig.~\ref{fig:convergence}(a). 
-In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix at once we use the standard initialization of QISKIT library. 
-Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case) or 27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case).
-In order to realize the circuit for the distance calculation for the quantum case we used the IBM Q machine. 
+[TB: could you explain why the figures in Fig.
+3 are not identical]
+
+An example of the QASOFM learning process is shown in Fig.~\ref{convergence}.
+Initially, the cluster vectors were randomly chosen (see Fig.~\ref{convergence}) and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
+In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT library.
+Each epoch [TB: can you define an epoch exactly?] of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case [TB: use the variable names defined above?]) or 27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case [TB: use the variable names defined above?]).
 After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector and shift the closest cluster vectors to the sample one.
 The shift is made by the change of the first binary element in the cluster vectors different from the sample one.
-The evolution of the labels presented on Fig. ~\ref{fig:convergence}(c).  
+The evolution of the labels presented on Fig.
+~\ref{convergence}(c).
 Good convergence is already observed in the fourth epoch.
+
+[TB: Somehow this feels like the paper ends suddenly without making broader statements.
+Do you have any ideas on how to add to this, and develop it a bit more?
+Just some theory is ok but now it feels like a good start but does not develop the ideas much]
 
@@
 \begin{figure}[t]
-	\label{fig:convergence}
 \includegraphics[width=0.95\columnwidth]{convergence.png}
 \caption{
-		(a) Initial random binary vectors of clusters are presented.
-We randomly sampled 3 binary vectors of size 9 as initial cluster vectors (labeled by 0,1,2) and it is shown how the words from the bag of words model presented in them.
-		(b) The result of applying our QASOFM implemented on the IBMQ "ibmq\_16\_melbourne" backend is shown on the bottom left figure. 
-		Vectors mean cluster elements for BIO, MED, QML groups, respectively (up to down). 
-		(c) The evolution of label distribution on each learning epoch is presented on the right figure. 
-		Good convergence already is observed on the fourth epoch.
+    (a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2), for the selected 9 words from the bag of words model.
+(b) The result of applying our QASOFM implemented on the IBM Q Experience ``ibmq\_16\_melbourne'' backend.
+Vectors mean cluster elements for BIO, MED, QML groups, for the C0, C1, C2 cluster vectors, respectively.
+(c) The evolution of label distribution on each learning epoch.
 } 
+\label{convergence}
 \end{figure}
@@
 \section{Discussions}
-We have developed a quantum assisted SOFM and showed a proof-of-concept demonstration that it can be used to solve clustering problems in an unsupervised manner with potential to outperform classical counterpart in real life tasks.
-The procedure of solving such clustering problem requires calculating the distance many times in iterative way. 
+We have developed a quantum assisted SOFM and showed a proof-of-concept experimental demonstration that it can be used to solve clustering problems in an unsupervised manner.
+The procedure of solving such clustering problem requires calculating the distance many times in iterative way, which is calculated using a hybrid quantum-classical procedure.
 We introduced an optimized circuit for Hamming distance calculations that can be implemented on currently available quantum computing devices with high fidelity.
-Our circuit could act as a distance-computing component of a classical SOFMs algorithm and in this way improve its performance. 
-Due to wide use of classical SOFMs in different areas of modern research this can give opportunities for the use of QASOFM in practical applications in near term.
+Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm and in this way improve its performance.
+[TB: Lets include a sentence about the classical and quantum scaling of the problem]
+Due to wide use of classical SOFMs in different areas of modern research and technology, this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
 
@@
 \section*{Keywords}
-quantum self-organizing feature map; 
-quantum machine learning; 
-neural network; 
-quantum unsupervised data clustering; 
-Hamming distance.
+quantum self-organizing feature map; quantum machine learning; neural network; quantum unsupervised data clustering; Hamming distance.
 
-\bibliographystyle{unsrt}
+\bibliographystyle{naturemag}
 \bibliography{bibliography}

### 2020-04-23 manuscript.tex
re-added the normalization
English improvements

@@
 \renewcommand{\r}{\right)}
+
+\usepackage[inline]{enumitem} % inline enumeration
 
@@
 	Unsupervised machine learning is one of the main techniques in artificial intelligence. 
-Quantum computers offer opportunities to speed up such machine learning techniques, where the number of input data elements is extremely large.
+	Quantum computers offer opportunities to speed up such machine learning techniques, 
+	where the number of input data elements is extremely large. 
 	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map artificial neural networks. 
-We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience and show that it allows us to reduce number of calculations in a number of clusters.
-[TB: Explain what the main technique that is used is.
-This is basically a QAOA hybrid quantum-classical approach right?
-If so, you should explain this] We compare the results with the classical algorithm on a toy example of unsupervised text clustering.
+	We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience 
+	and show that it allows us to reduce number of calculations in a number of clusters. 
+	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.  
 \end{abstract}
@@
 Paradoxically, the exact mechanism of why neural networks are so powerful remain unknown (in many cases it is regarded as a black box).  
-It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
+It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, 
+and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
 
-In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) that is trained in an unsupervised manner.
-SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
-The SOFMs with a small number of nodes are similar to the K-means algorithm but for larger SOFMs they represent data in a fundamentally topological way that allows one to do dimensionality reduction.
+In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN) 
+that is trained in an unsupervised manner. 
+SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} 
+and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}. 
+The SOFMs with a small number of nodes are similar to the K-means algorithm 
+but for larger SOFMs they represent data in a fundamentally topological way that allows one to do dimensionality reduction.  
 Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
@@
 More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019}. 
-The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits, where the parameters behave much like the weights of a neural network \cite{lewenstein1994}.
-In particular, quantum algorithms for training and evaluating feedforward neural networks was developed, which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
+The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits, 
+where the parameters behave much like the weights of a neural network \cite{lewenstein1994}. 
+In particular, quantum algorithms for training and evaluating feedforward neural networks was developed, 
+which are one of the most usable neural network models \cite{allcock2018, tacchino2019}. 
 Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}. 
@@
 Sublinear quantum algorithms for training linear and kernel-based classifiers were designed \cite{li2019}. 
-Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, quantum autoencoders \cite{bondarenko2019}, quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, vinci2019, lu2019} were developed.
+Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, 
+quantum autoencoders \cite{bondarenko2019}, 
+quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} 
+and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, vinci2019, lu2019} were developed. 
 Currently, it is believed that implementation of quantum neural networks can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.
 
-[TB: I think we definitely need to cite and describe hybrid quantum-classical approaches, like QAOA.  Could you please add several sentences and a solid set of references for these]
-
-In this paper, we analyze  a hybrid quantum assisted SOFM (QASOFM) and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.
-We implement the quantum assisted SOFM in such a way that it becomes possible to use the Hamming distance as a distance metric for the training and to reduce the number of distance calculations in the number of clusters.
+In this paper, we analyze  a hybrid quantum assisted SOFM (QASOFM) 
+and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.   
+We implement the quantum assisted SOFM in such a way 
+that it becomes possible to use the Hamming distance as a distance metric for the training 
+and to reduce the number of distance calculations in the number of clusters. 
 We propose an optimized circuit for realizing the Hamming distance on a quantum machine. 
-We then apply it to a toy example of data for clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} and compare it to the classical case.
+We then apply it to a toy example of data for clustering paper abstracts 
+and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} 
+and compare it to the classical case.
 
 \begin{figure}
-  \includegraphics[width=0.95\columnwidth]{clusters.png}
-\caption{Schematic illustration of the clustering problem considered in this paper.
+	\label{fig:sofm_fitting}
+  	\includegraphics[width=0.95\columnwidth]{sofm_fitting.png}
+	\caption{
+		Schematic illustration of the clustering problem considered in this paper.  
 		Blue dots represent clusters and red dots are data points. 
 		The training process moves clusters to fit the data points. 
-Note that there are fewer clusters than data points, which is the essence of dimensionality reduction, and is what permits the model to generalize the data.}
-\label{clusters}
+		Note that there are fewer clusters than data points, 
+		which is the essence of dimensionality reduction, 
+		and is what permits the model to generalize the data.
+	}
 \end{figure}
@@
 
-The input dimensions are associated with the features, and the nodes in the grid (called cluster vectors) are assigned the $N$-dimensional vectors.
+The input dimensions are associated with the features, 
+and the nodes in the grid (called cluster vectors) are assigned the $N$-dimensional vectors. 
 The components of these vectors are usually called weights. 
 Initially the weight components are chosen randomly. 
-We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of selecting a winning cluster vector and updating its weights (Fig.~\ref{clusters}).
-More specifically, they consist of four step process: 1) selecting an input vector randomly from the set of all input vectors; 2) finding a cluster vector which is closest to the input vector; 3) adjusting the weights of the winning node in such a way that it becomes even closer to the input vector; 4) repeating this process for many iterations until it converges.
+We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of selecting a winning cluster vector and updating its weights (Fig.~\ref{fig:sofm_fitting}).
+More specifically, they consist of four step process: \begin{enumerate*}
+\item selecting an input vector randomly from the set of all input vectors; 
+\item finding a cluster vector which is closest to the input vector; 
+\item adjusting the weights of the winning node in such a way that it becomes even closer to the input vector; 
+\item repeating this process for many iterations until it converges.
+\end{enumerate*}
 
-After the winning cluster vector is selected, the weights of the vector are adjusted according to [TB: all variables should be defined: what is $ \vec{x} , \alpha $ etc.
-] [TB: this is done iteratively?
-Then should there be an index over  $ \vec{x}_i $?
-]
-%
+After the winning cluster vector is selected, the weights of the vector are adjusted according to
+
 \begin{equation}
+    \label{eq:learning}
     \vec w_{\mathrm{new}} =
     	\vec w_{\mathrm{old}} 
-        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  .
-    \label{eq:learning}
+        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  
 \end{equation}
-%
-The above expression can be interpreted according to: if a component of the input vector is greater than the corresponding weight, increase the weight by a small amount; if the input component is smaller than the weight, decrease the
-weight by a small amount.
+
+The above expression can be interpreted according to: 
+if a component of the input vector is greater than the corresponding weight, 
+increase the weight by a small amount; 
+if the input component is smaller than the weight, decrease the weight by a small amount. 
 The larger the difference between the input component and the weight component, the larger the increment (decrement). 
-Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects.
-A visualisation of this process is shown in Fig.
-\ref{clusters}.
+Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way 
+that ensures each move is following the current trends inferred from their distances to the input objects. 
+A visualisation of this process is shown in Fig. \ref{fig:sofm_fitting}.
 
@@
 In our approach, we use the Hamming distance instead of the Euclidean distance to select the winning cluster vector. 
-It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
+It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
+such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
 
@@
 \begin{figure}[t]
-\includegraphics[width=0.95\columnwidth]{circuit.png}
+	\label{fig:qcircuit}
+	\includegraphics[width=0.95\columnwidth]{qcircuit.png}
 	\caption{
@@
 		Finally, a quantum state of the sample register returned to the initial basis for information retrieval. 	
-[TB: The figure is too small to be read, please make this bigger]
 	} 
-\label{circuit}
 \end{figure}
@@
 \begin{figure}[t]
+	\label{fig:distance_matrix}
 	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
-\caption{Hamming distance matrix between two data sets of binary vectors.
+	\caption{
+		Hamming distance matrix between two data sets of binary vectors. 
 		Values of distance that are less than the median distance value are marked black. 
@@
 		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.  
-[TB: could you put (a) (b) labels as with the other figures.
-]}
-\label{fig:distance_matrix}
+		}
 \end{figure}
@@
 This allows for a simple encoding of the classical information into a quantum register.
-[TB: allows?]
 
@@
 where $d^\alpha_{ij} = \mathrm{CNOT}(y^\alpha_i, x^\alpha_j)$, $\alpha = 1 \dots n$, and $i,j$ are the qubit indexes in the registers. 
-At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors, instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$.
+At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors,
+instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$. 
 Next, for each pair $\{X\}$ and $\{Y\}$, the accumulated information of all the differences is projected onto the amplitude of the superposed state. 
-This is achieved by applying the Hadamard gate on auxiliary qubit, followed by a controlled phase gate on $\left| Xa \right\rangle$ defined as
+This is achieved by applying the Hadamard gate on auxiliary qubit, 
+followed by a controlled phase gate on $\left| Xa \right\rangle$ defined as
 %
@@
 %
-Finally, another Hadamard gate is applied. [TB: on which qubit? I think its better just to describe this using a circuit, rather than describe the steps in words. ]
+Finally, another Hadamard gate is applied.
 
@@
     \left| \psi_2 \right\rangle = H_a\left| \psi_1 \right\rangle = 
-    \sum\limits_{i, j=1}^{k} 
+    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
     \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
@@
 %
-Applying the controlled phase gate the state then becomes [TB: please define all variables like $ R_{(X,a)} $]
+Applying the controlled phase gate the state then becomes: 
 %
@@
     \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
-    \\ = \dfrac{1}{\sqrt{2}}
+    \\ = \dfrac{1}{\sqrt{2kl}}
 				\sum\limits_{i, j=1}^{k} 
@@
         \left| 0 \right\rangle
-        \\ + \dfrac{1}{\sqrt{2}}
+        \\ + \dfrac{1}{\sqrt{2kl}}
 				\sum\limits_{i, j=1}^{k}
@@
 %
-%
-%\\ = \dfrac{1}{\sqrt{2}}
-%    \sum\limits_{i, j}^{k}
-%    \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-%        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
-%        \left| 0 \right\rangle
-%        \\ + \dfrac{1}{\sqrt{2}}
-%				\sum\limits_{i, j}^{k}
-%        \left| R(d^{(1)}_{ij}, 1)\left(\dfrac{\pi}{n}\right), \dots, R(d^{(n)}_{ij}, 1)\left(\dfrac{\pi}{n}\right) \right\rangle 
-%        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
-%        \left| 1 \right\rangle 
-%
 Applying another Hadamard on the ancilla qubit we obtain
@@
     \left| \psi_4 \right\rangle = 
-    \sum\limits_{i, j=1}^{k} 
+    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
     \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
@@
 There are two possible measurement outcomes of the auxiliary qubit.  
-Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
+Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, 
+the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
 
-At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, thus we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, 
+thus we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
 %
@@
     \mathrm{CNOT} (Y,X)\left| \psi_4 \right\rangle \\=  
-    \sum\limits_{i, j=1}^{k} 
+    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
     \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
@@
 In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0. 
-If the ancilla qubit is 1, there is an inverse relationship with the measurement result.  [TB: what kind of inverse relationship?]
+If the ancilla qubit is 1, there is an inverse relationship with the measurement result.
 
-Measuring the Hamming distance of a particular pair of input vectors $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form, this can be done using the following projection operator
+Measuring the Hamming distance of a particular pair of input vectors $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form,
+this can be done using the following projection operator
 %
@@
 %
-In order to reduce noise we average the measurement results over different states of the ancilla qubit, thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ is
+In order to reduce noise we average the measurement results over different states of the ancilla qubit, 
+thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ is
 %
@@
 %
-The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$, where  $0$ indicates that $x_i$ are $y_j$ identical and $1$ means they are completely opposite in terms of their pairwise binary coordinates.
+The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$, 
+where  $0$ indicates that $x_i$ are $y_j$ identical and $1$ means they are completely opposite in terms of their pairwise binary coordinates.
 
-In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, our method allows one to reduce the number of gates in the circuit and reconstruct all distance matrix in parallel and can be implemented with high fidelity.
-We note that at least for the $2n$ one-qubit NOT gates, this depends on the actual realization of the control phase gate in a quantum register at the hardware level.
-[TB: don't understand this sentence]
-
-% Please , add comparison of complexities in number of gates for realization between our scheme and previously proposed schemes (one qubit and two qubit separately)  
-% ??
-
-%\begin{align}
-%\\
-%\end{align} 
-%
-
-%
-%\begin{equation}
-%\end{equation}
-%(a black region indicates if the word occurs in the abstract)
+In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
+our method allows one to reduce the number of gates in the circuit and reconstruct all distance matrix in parallel and can be implemented with high fidelity.  
+We note that at least for the $2n$ one-qubit NOT gates, 
+this depends on the actual realization of the control phase gate in a quantum register at the hardware level.
 
@@
 \begin{figure}[t]
+	\label{fig:vectorized_sample}
 	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
-\caption{(a) Representation of the data set of abstracts with the bag-of-words model is shown.
+	\caption{
+		(a) Representation of the data set of abstracts with the bag-of-words model is shown. 
 		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis. 
@@
 	} 
-\label{fig:vectorized_sample}
 \end{figure}
@@
 We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.  
-We perform unsupervised data clustering for three sets of paper abstracts from three different fields: quantum physics \cite{qml0, qml1, qml2}, medicine \cite{med0, med1, med2} and biology \cite{bio0, bio1, bio2}.
-Each set consists of three random papers that focus on one of following topics: ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO).
+We perform unsupervised data clustering for three sets of paper abstracts from three different fields: 
+quantum physics \cite{qml0, qml1, qml2}, 
+medicine \cite{med0, med1, med2} 
+and biology \cite{bio0, bio1, bio2}. 
+Each set consists of three random papers that focus on one of following topics: 
+``Quantum Machine Learning'' (QML), 
+``Cancer'' (MED) 
+and ``Gene Expression'' (BIO). 
 Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}).  
-[TB: bag of words reference?]  This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words.
+This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
 Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms.
@@
 After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.  
-We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, showing successful self-organization (Fig.~\ref{fig:vectorized_sample}).
+We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, 
+showing successful self-organization (Fig.~\ref{fig:vectorized_sample}).
 
-In classical SOFM \cite{kohonen1990}, first a distance calculation between a a sample vector and all cluster vectors is made, then the closest cluster vector is shifted towards the sample vector.
-The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$, where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.
+In classical SOFM \cite{kohonen1990}, first a distance calculation between a a sample vector and all cluster vectors is made, 
+then the closest cluster vector is shifted towards the sample vector. 
+The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$, 
+where $N$ is number of samples, 
+$M$ is number of randomly sampled cluster vectors, 
+and $L$ is number of the shifts of cluster vectors.  
 In the QASOFM described in the previous section, distance calculations are realized on a quantum device (i.e.
 the IBM Q Experience) with the use of circuit presented in Fig.
-\ref{circuit}.
-This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates that are possible to realize on  quantum computing devices that are currently available.
-The circuit realized in such a manner that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation.
+\ref{fig:qcircuit}.
+This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates
+that are possible to realize on  quantum computing devices that are currently available. 
+The circuit realized in such a manner 
+that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation. 
 The complexity of the quantum assisted SOFM then scales as $O(LN)$.
@@
 The theoretical calculations and classical simulations show perfect agreement with each other.
-[TB: could you explain why the figures in Fig.
-3 are not identical]
 
 An example of the QASOFM learning process is shown in Fig.~\ref{convergence}. 
-Initially, the cluster vectors were randomly chosen (see Fig.~\ref{convergence}) and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
+Initially, the cluster vectors were randomly chosen (see Fig.~\ref{convergence}) 
+and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
 In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT library.
-Each epoch [TB: can you define an epoch exactly?] of the algorithm requires 9 distance calculations in the  quantum implementation (number of samples in general case [TB: use the variable names defined above?]) or 27 distance calculations for classical realization (product of number of samples and number of cluster vectors in general case [TB: use the variable names defined above?]).
-After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector and shift the closest cluster vectors to the sample one.
+Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation 
+(number of samples in general case) 
+or 27 distance calculations for classical realization 
+(product of number of samples and number of cluster vectors in general case). 
+After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector 
+and shift the closest cluster vectors to the sample one. 
 The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
-The evolution of the labels presented on Fig.
-~\ref{convergence}(c).
-Good convergence is already observed in the fourth epoch.
-
-[TB: Somehow this feels like the paper ends suddenly without making broader statements.
-Do you have any ideas on how to add to this, and develop it a bit more?
-Just some theory is ok but now it feels like a good start but does not develop the ideas much]
+The evolution of the labels presented on Fig. ~\ref{convergence}(c).  Good convergence is already observed in the fourth epoch.
 
@@
 	\caption{
-    (a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2), for the selected 9 words from the bag of words model.
+		(a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2), 
+		for the selected 9 words from the bag of words model. 
 		(b) The result of applying our QASOFM implemented on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
-Vectors mean cluster elements for BIO, MED, QML groups, for the C0, C1, C2 cluster vectors, respectively.
+		Vectors mean cluster elements for BIO, MED, QML groups, 
+		for the C0, C1, C2 cluster vectors, respectively. 
 		(c) The evolution of label distribution on each learning epoch.
@@
 \section{Discussions}
-We have developed a quantum assisted SOFM and showed a proof-of-concept experimental demonstration that it can be used to solve clustering problems in an unsupervised manner.
-The procedure of solving such clustering problem requires calculating the distance many times in iterative way, which is calculated using a hybrid quantum-classical procedure.
+We have developed a quantum assisted SOFM and showed a proof-of-concept experimental demonstration 
+that it can be used to solve clustering problems in an unsupervised manner.  
+The procedure of solving such clustering problem requires calculating the distance many times in iterative way, 
+which is calculated using a hybrid quantum-classical procedure.  
 We introduced an optimized circuit for Hamming distance calculations that can be implemented on currently available quantum computing devices with high fidelity. 
-Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm and in this way improve its performance.
-[TB: Lets include a sentence about the classical and quantum scaling of the problem]
-Due to wide use of classical SOFMs in different areas of modern research and technology, this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
+Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm 
+and in this way improve its performance. 
+Due to wide use of classical SOFMs in different areas of modern research and technology, 
+this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
 

### 2020-04-29 manuscript.tex
Alexey's response

@@
 	Unsupervised machine learning is one of the main techniques in artificial intelligence. 
-	Quantum computers offer opportunities to speed up such machine learning techniques, 
-	where the number of input data elements is extremely large. 
+	Quantum computers offer opportunities to speed up such machine learning techniques. 
 	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map artificial neural networks. 
@@
 Quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
-More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019}. 
+More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019, broughton2020}. 
 The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits, 
@@
 quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} 
-and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, vinci2019, lu2019} were developed. 
-Currently, it is believed that implementation of quantum neural networks can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.
+and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019} were developed. 
+At the same time, non-neural-network based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}. 
+Such hybrid algorithms involve as a part some quantum circuits for speed-up and usually trained in a standard classical learning loop. 
+In particular, the Quantum Approximate Optimization Algorithm was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016} 
+and designed to problems such as MAX-CUT, Grover’s search algorithm and so on \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019}. 
+Another example of well-known hybrid quantum classical algorithms is Variational Quantum Eigensolver (VQE) for applications in quantum chemistry 
+\cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
+Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.
 
@@
 \begin{figure}
-	\label{fig:sofm_fitting}
   	\includegraphics[width=0.95\columnwidth]{sofm_fitting.png}
@@
 	}
+	\label{fig:sofm_fitting}
 \end{figure}
 
-\section{The quantum assisted self-organizing feature map (SOFM)}
+\section{The quantum assisted self-organizing feature map (QASOFM)}
 \label{sec:qasofm}
@@
 It was first proposed by Kohonen as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}. 
-The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, and maps them onto nodes of a two-dimensional grid.
+The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, 
+and describes a mapping from a higher-dimensional input space to a lower-dimensional map space.
 
@@
 After the winning cluster vector is selected, the weights of the vector are adjusted according to
-
+%
 \begin{equation}
+	\vec w_{i+1} =
+		\vec w_i
+		+ \alpha\left(\vec{x} - \vec w_i\right).
     \label{eq:learning}
-    \vec w_{\mathrm{new}} =
-    	\vec w_{\mathrm{old}} 
-        + \alpha\left(\vec{x} - \vec w_{\mathrm{old}}\right)  
 \end{equation}
-
+%
 The above expression can be interpreted according to: 
-if a component of the input vector is greater than the corresponding weight, 
-increase the weight by a small amount; 
+if a component of the input vector $\vec{x}$ is greater than the corresponding weight $ \vec{w}_i $, 
+increase the weight by a small amount with the learning rate $\alpha$; 
 if the input component is smaller than the weight, decrease the weight by a small amount. 
@@
 %%%%%
-\begin{figure}[t]
-	\label{fig:qcircuit}
-	\includegraphics[width=0.95\columnwidth]{qcircuit.png}
+\begin{figure*}[t]
+	\includegraphics[width=1.9\columnwidth]{qcircuit.png}
 	\caption{
@@
 	} 
-\end{figure}
+	\label{fig:qcircuit}
+\end{figure*}
 %%%%%
@@
 \begin{figure}[t]
-	\label{fig:distance_matrix}
 	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
 	\caption{
-		Hamming distance matrix between two data sets of binary vectors. 
+		(a) Hamming distance matrix between two data sets of binary vectors. 
 		Values of distance that are less than the median distance value are marked black. 
-		Classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure. 
-		Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend is shown on the right figure. 
+		Classical simulation of the quantum circuit shows perfect agreement to theoretical calculations. 
+		(b) Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
 		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.  
 	}
+	\label{fig:distance_matrix}
 \end{figure}
@@
 We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors, making use of quantum parallelism.
-This allows for a simple encoding of the classical information into a quantum register.
+
+%This allows for a simple encoding of the classical information into a quantum register.
 
@@
 Let us assume we have $k$ input vectors and $l$ cluster states. 
-The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$, $\left| y_j \right\rangle$. 
+The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| X_i \right\rangle$, $\left| Y_j \right\rangle$. 
 The registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ are initialized to store the input vectors and cluster vectors according to
@@
 \begin{align}
-    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
-    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
+    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| X_i \right\rangle,  \\
+    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| Y_j \right\rangle .
     \label{eq:encodnig}
@@
     \label{eq:control_phase_rotation}
-    R(\phi) = 
+    R_{(X,a)}(\phi) = 
     \begin{pmatrix}
@@
 %
-Finally, another Hadamard gate is applied.
+Finally, another Hadamard gate is applied on the ancilla qubit (see Fig. \ref{fig:qcircuit}).
 
@@
 %
-Applying the controlled phase gate the state then becomes: 
+Applying the controlled phase gate the state then becomes
 %
@@
 
-At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ is no longer relevant, 
-thus we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ \hl{encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis} for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
 %
@@
 %
-This makes $\{X\}$ store the input vectors again, as in the initial step. 
-It however preserves the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
+This makes $\{X\}$ store the input vectors again, as in the initial step 
+and preserves the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
 
@@
 In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0. 
-If the ancilla qubit is 1, there is an inverse relationship with the measurement result.
+If the ancilla qubit is 1, the smallest amplitude of the measurement result coincides with the smallest Hamming distance.
 
@@
 \begin{align}
-\Pi_{i,j} = &\left| X_i \rangle\langle X_i \right| \otimes \left| X_i \rangle\langle Y_j \right| \otimes I .
+\Pi_{i,j} = &\left| X_i \rangle\langle X_i \right| \otimes \left| Y_j \rangle\langle Y_j \right| \otimes I .
 \end{align} 
@@
 In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
-our method allows one to reduce the number of gates in the circuit and reconstruct all distance matrix in parallel and can be implemented with high fidelity.  
-We note that at least for the $2n$ one-qubit NOT gates, 
-this depends on the actual realization of the control phase gate in a quantum register at the hardware level.
+our method allows one to reduce the number of gates in the circuit (at least in $2n$ one-qubit NOT gates)
+and reconstruct all distance matrix in parallel and can be implemented with high fidelity.
 
@@
 \begin{figure}[t]
-	\label{fig:vectorized_sample}
 	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
@@
 	} 
+	\label{fig:vectorized_sample}
 \end{figure}
@@
 
-\section{Expeirmental demonstration of QASOFM}
+\section{Experimental demonstration of QASOFM}
 
@@
 and ``Gene Expression'' (BIO). 
-Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}).  
+Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
 This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
@@
 We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
-The theoretical calculations and classical simulations show perfect agreement with each other.
+The theoretical calculations and classical simulations show perfect agreement with each other (Fig. \ref{fig:distance_matrix}(a)). 
+The difference between Figs. \ref{fig:distance_matrix} (a) and (b) appears due to noise in the currently available non fault-tolerant quantum processors.
 
@@
 In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT library.
-Each epoch of the algorithm requires 9 distance calculations in the  quantum implementation 
-(number of samples in general case) 
+Each epoch of the algorithm consists of distance calculation between all data and cluster vectors 
+and requires 9 distance calculations in the  quantum implementation ($N$ in general case) 
 or 27 distance calculations for classical realization 
-(product of number of samples and number of cluster vectors in general case). 
+($MN$ in general case). 
 After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector 
@@
 The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
-The evolution of the labels presented on Fig. ~\ref{convergence}(c).  Good convergence is already observed in the fourth epoch.
+The evolution of the labels presented on Fig. ~\ref{convergence}(c).  
+Good convergence is already observed in the fourth epoch.
+
+This proof of concept example shows that developing of noisy intermediate scale quantum hardware will allow to solve some practical problems in unsupervised manner with very simple encoding of categorical data to quantum register.
+Furthermore, it shows that it is possible to develop distance based hybrid quantum classical algorithms 
+that speed up classical counterparts and can be implemented on near future quantum devices 
+and that further developing distance based hybrid algorithms, 
+that can learn in unsupervised manner, for categorical data such as genomic and so on has significant interest.
 
@@
 		for the selected 9 words from the bag of words model. 
-		(b) The result of applying our QASOFM implemented on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
+		(b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend. 
 		Vectors mean cluster elements for BIO, MED, QML groups, 
@@
 and in this way improve its performance. 
+The complexity of the quantum assisted SOFM scales as $O(LN)$ 
+while the complexity of the fully classical SOFM scales as $O(LMN)$, 
+where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, 
+and $L$ is number of the shifts of cluster vectors.
 Due to wide use of classical SOFMs in different areas of modern research and technology, 

### 2020-05-03 manuscript.tex
Fix qcircuit figure and caption
See #14

@@
 The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, 
-and describes a mapping from a higher-dimensional input space to a lower-dimensional map space.
+and describes a mapping from a higher-dimensional input space to a lower-dimensional map space, commonly a bi-dimensional map.
 
@@
 The components of these vectors are usually called weights. 
-Initially the weight components are chosen randomly. 
-We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of selecting a winning cluster vector and updating its weights (Fig.~\ref{fig:sofm_fitting}).
-More specifically, they consist of four step process: \begin{enumerate*}
+Initially the weight components are chosen randomly 
+and topological distances between neurons given. 
+We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of 
+selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
+More specifically, they consist of four step process: 
+\begin{enumerate*}
 \item selecting an input vector randomly from the set of all input vectors; 
-\item finding a cluster vector which is closest to the input vector; 
-\item adjusting the weights of the winning node in such a way that it becomes even closer to the input vector; 
+\item finding a cluster vector which is closest to the input vector (BMU); 
+\item adjusting the weights of the BMU and neurons close to it on feature map in such a way 
+	that these vectors  becomes even closer to the input vector; 
 \item repeating this process for many iterations until it converges.
@@
 
-After the winning cluster vector is selected, the weights of the vector are adjusted according to
+On a step $t$ when the BMU $w_{c}$ is selected, 
+the weights of the BMU and its neigbours on feature map are adjusted according to
 %
 \begin{equation}
-	\vec w_{i+1} =
-		\vec w_i
-		+ \alpha\left(\vec{x} - \vec w_i\right).
     \label{eq:learning}
+	\vec{w_{i}}(t + 1) 
+	= \vec{w_i}(t)
+	+ \theta(c, i, t) \alpha(t) 
+		\left(\vec{x}(t) - \vec{w_i}(t)\right),
 \end{equation}
 %
-The above expression can be interpreted according to: 
-if a component of the input vector $\vec{x}$ is greater than the corresponding weight $ \vec{w}_i $, 
-increase the weight by a small amount with the learning rate $\alpha$; 
-if the input component is smaller than the weight, decrease the weight by a small amount. 
-The larger the difference between the input component and the weight component, the larger the increment (decrement). 
+where  $\alpha(t)$ is the learning rate and $\theta(c, i, t)$ is the neighbor function,
+which defines the topological neighbor neurons to be updated. 
+Note that neighbor function depend on distance on feature map given initialy, 
+but not distance metrics between vectors.
+
 Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way 
@@
 
-Usually the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors. 
-In our approach, we use the Hamming distance instead of the Euclidean distance to select the winning cluster vector. 
-It allows us to use a simpler encoding of the classical information into the quantum state and use an effective procedure for the calculation of the Hamming distance on the quantum machine, 
-such as to reduce the number of calculations in number of cluster vectors in comparison to the classical case.
+In SOFM original version was designed to cluster real-valued data
+and the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors.
+The paper deals with binary vectors clustering problem 
+and for binary data the Hamming distance is more suitable \cite{appiah2009, santana2017}.
+Using a known technique of encoding classical information into a quantum register, 
+based on probabilistic quantum memories \cite{trugenberger2001},
+we introduce an optimized algorithm for calculating the matrix of Hamming distances 
+between each pair of binary vectors  of two sets taking advantage of quantum parallelism.
 
@@
 %%%%%
-\begin{figure*}[t]
-	\includegraphics[width=1.9\columnwidth]{qcircuit.png}
+\begin{figure}[t]
+	\includegraphics[width=\columnwidth]{qcircuit.eps}
 	\caption{
-		Quantum circuit for the quantum parallelized Hamming distance calculation between all samples and clusters. 
-		First,  Hadamard gates are applied in order to obtain a superposition of the sample and cluster registers. 
-		Second, we encoded information about pairwise different qubits in a quantum state of the sample register with applying the CNOT gates. 
-		Third, Hamming distance values are encoded in the amplitudes of superposition with the control phase rotation and Hadamard gates. 
-		Finally, a quantum state of the sample register returned to the initial basis for information retrieval. 	
+		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded \cite{trugenberger2001} in $X$ and $Y$ quantum registers   respectively.
+		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 
+		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate and Hadamard gates. 
+		Finally, a quantum state of the $X$-register returned to the initial basis for information retrieval. 	
 	} 
 	\label{fig:qcircuit}
-\end{figure*}
+\end{figure}
 %%%%%
@@
 %%%%%
-
-We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors, making use of quantum parallelism.
-
-%This allows for a simple encoding of the classical information into a quantum register.
 
@@
 This is achieved by applying the Hadamard gate on auxiliary qubit, 
-followed by a controlled phase gate on $\left| Xa \right\rangle$ defined as
+followed by a controlled rotation around $z$-axis gate on $\left| Xa \right\rangle$ defined as
 %
@@
         1 & 0 & 0 & 0 \\
-        0 & 1 & 0 & 0 \\
+        0 & e^{-i \frac \phi 2} & 0 & 0 \\
         0 & 0 & 1 & 0 \\
-        0 & 0 & 0 & e^{-i\phi}
+		0 & 0 & 0 & e^{i \frac \phi 2}
     \end{pmatrix} .
@@
 %
-Applying the controlled phase gate the state then becomes
+Applying the controlled rotation around $z$-axis gate the state then becomes
 %
@@
 Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, 
-the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
+the controlled rotation around $z$-axis gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
 
@@
 and ``Gene Expression'' (BIO). 
-Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
+Abstracts were vectorized by the bag-of-words\cite{weikang2016} model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
 This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 

### 2020-05-03 manuscript.tex
Fix this allows for a simple encoding #15

--- 2020-04-29 manuscript.tex
@@
 
-We now introduce an optimized algorithm for calculating the matrix of Hamming distances \cite{trugenberger2001} between a sample vector and all cluster vectors, making use of quantum parallelism.
-
-%This allows for a simple encoding of the classical information into a quantum register.
+Using a known technique of encoding classical information into a quantum register, 
+based on probabilistic quantum memories \cite{trugenberger2001},
+we introduce an optimized algorithm for calculating the matrix of Hamming distances between each pair of binary vectors  of two sets taking advantage of quantum parallelism.
+This allows for a simple encoding of the classical information into a quantum register.
 

### 2020-05-03 manuscript.tex
Imrove Step by step circuit explanation #16

--- 2020-05-03 manuscript.tex
@@
 		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 
-		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate and Hadamard gates. 
+		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate~(\ref{eq:controled_rotation}) and Hadamard gates. 
 		Finally, a quantum state of the $X$-register returned to the initial basis for information retrieval. 	
@@
 Let us assume we have $k$ input vectors and $l$ cluster states. 
-The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| X_i \right\rangle$, $\left| Y_j \right\rangle$. 
+The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$, $\left| y_j \right\rangle$. 
 The registers $\left| X \right\rangle$ and $\left| Y \right\rangle$ are initialized to store the input vectors and cluster vectors according to
@@
 \begin{align}
-    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| X_i \right\rangle,  \\
-    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| Y_j \right\rangle .
+    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
+    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
     \label{eq:encodnig}
@@
 
-Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate between $\left| X \right\rangle$ and $\left| Y \right\rangle$
+Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate on $\left| x^{(\alpha)} y^{(\alpha)} \right\rangle$ and $\alpha = 1 \dots n$
 
-\begin{align}
-    | \psi_1 \rangle & = 
-    \mathrm{CNOT} (Y,X)| \psi_0 \rangle \nonumber \\
-		& =  
+\begin{equation}
+    | \psi_1 \rangle  =  
     \frac{1}{\sqrt{kl}} \sum_{i, j=1}^{k} 
@@
     | 0 \rangle 
-\end{align}
+\end{equation}
 %
-where $d^\alpha_{ij} = \mathrm{CNOT}(y^\alpha_i, x^\alpha_j)$, $\alpha = 1 \dots n$, and $i,j$ are the qubit indexes in the registers. 
+where $d^{(\alpha)}_{ij} = \mathrm{CNOT}(y^{(\alpha)}_i, x^{(\alpha)}_j)$, and $\alpha = 1 \dots n$  is the qubit index in the register. 
 At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors,
@@
 \begin{equation}
-    \label{eq:control_phase_rotation}
-    R_{(X,a)}(\phi) = 
+    \label{eq:controled_rotation}
+    C_{R_z}(\phi) = 
     \begin{pmatrix}
@@
 \begin{equation}
-    \left| \psi_2 \right\rangle = H_a\left| \psi_1 \right\rangle = 
+    \left| \psi_2 \right\rangle = 
     \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
-    \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-    \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle
+    \left| d_{ij} \right\rangle 
+    \left| y_j \right\rangle
     \dfrac{(\left| 0 \right\rangle + \left| 1 \right\rangle)}{\sqrt{2}}  .
@@
 %
-Applying the controlled rotation around $z$-axis gate the state then becomes
+Applying the controlled rotation around $z$-axis gate on $\left| x^{(\alpha)} a \right\rangle$ where $\alpha = 1\dots n$ and the state then becomes
 %
 \begin{multline}
-    \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
-    \\ = \dfrac{1}{\sqrt{2kl}}
-				\sum\limits_{i, j=1}^{k} 
-				\left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
-        \left| 0 \right\rangle
-        \\ + \dfrac{1}{\sqrt{2kl}}
-				\sum\limits_{i, j=1}^{k}
-        \exp\left(\dfrac{-i \pi}{n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-\\ \times        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
+    \left| \psi_3 \right\rangle  
+	= \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k} 
+		\exp\left(
+		    \dfrac{-i \pi}{2n}
+		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
+		\right)
+		\left| d_{ij} \right\rangle 
+		\left| y_j \right\rangle 
+		\left| 0 \right\rangle \\
+    + \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k}
+		\exp\left(
+		    \dfrac{i \pi}{2n}
+		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
+		\right)
+		\left| d_{ij} \right\rangle  
+		\left| y_j \right\rangle 
 		\left| 1 \right\rangle
@@
     \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
-    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-		\\ \times
-        \left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
+        \left[ 
+			\cos\left(
+			    \dfrac{\pi}{2n}
+			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
+			\right)
+			\left| d_{ij} \right\rangle 
+			\left| y_j \right\rangle 
 			\left| 0 \right\rangle\right.
 			\\+ 
-        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-        \left| d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \right\rangle 
-        \left| y^{(1)}_j, \dots, y^{(n)}_j \right\rangle 
-        \left| 1 \right\rangle\right] .
+			\left. i \sin\left(
+			    \dfrac{\pi}{2n} 
+			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
+			\right)
+			\left| d_{ij} \right\rangle 
+			\left| y_j \right\rangle 
+			\left| 1 \right\rangle
+		\right] .
 \end{multline}
@@
     \left| \psi_f \right\rangle =  
-    \mathrm{CNOT} (Y,X)\left| \psi_4 \right\rangle \\=  
     \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
-    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-				\left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-        \left| X_i \right\rangle 
-        \left| Y_j \right\rangle 
+	\left[ 
+	    \cos\left(
+	        \dfrac{\pi}{2n}
+	        \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
+	    \right)
+        \left| x_i \right\rangle 
+        \left| y_j \right\rangle 
         \left| 0 \right\rangle\right.
         \\+
-        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
-        \left| X_i \right\rangle 
-        \left| Y_j \right\rangle 
-        \left| 1 \right\rangle\right] .
+        \left. i \sin\left(
+            \dfrac{\pi}{2n}
+            \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
+        \right)
+        \left| x_i \right\rangle 
+        \left| y_j \right\rangle 
+        \left| 1 \right\rangle
+    \right] .
 \end{multline}
@@
 
-Measuring the Hamming distance of a particular pair of input vectors $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form,
+Measuring the Hamming distance of a particular pair of input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form,
 this can be done using the following projection operator
@@
 \begin{align}
-\Pi_{i,j} = &\left| X_i \rangle\langle X_i \right| \otimes \left| Y_j \rangle\langle Y_j \right| \otimes I .
+\Pi_{i,j} = &\left| x_i \rangle\langle x_i \right| \otimes \left| y_j \rangle\langle y_j \right| \otimes I .
 \end{align} 
@@
 In order to reduce noise we average the measurement results over different states of the ancilla qubit, 
-thus the measured Hamming distance between the input vector $\left| X_i \right\rangle$ and cluster vector $\left| Y_j \right\rangle$ is
+thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
 %

### 2020-05-03 manuscript.tex
Imrove Step by step circuit explanation #16

@@
 	\caption{
-		(a) Representation of the data set of abstracts with the bag-of-words model is shown. 
+		(a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown. 
 		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis. 
@@
 and ``Gene Expression'' (BIO). 
-Abstracts were vectorized by the bag-of-words model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
+Abstracts were vectorized by the bag-of-words \cite{weikang2016} model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
 This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 

### 2020-05-03 manuscript.tex
Improve SOFM description #2 #4 #13 #15 #19

@@
 %%%%%
-
-Using a known technique of encoding classical information into a quantum register, 
-based on probabilistic quantum memories \cite{trugenberger2001},
-we introduce an optimized algorithm for calculating the matrix of Hamming distances between each pair of binary vectors  of two sets taking advantage of quantum parallelism.
-This allows for a simple encoding of the classical information into a quantum register.
 

### 2020-05-04 manuscript.tex
Fix controlled rotation mention

--- 2020-05-03 manuscript.tex
@@
 This is achieved by applying the Hadamard gate on auxiliary qubit, 
-followed by a controlled rotation around $z$-axis gate on $\left| Xa \right\rangle$ defined as
+followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$ where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
 %

### 2020-05-05 manuscript.tex
Change for issue #22

@@
 This is achieved by applying the Hadamard gate on auxiliary qubit,
-followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$ where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
+followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$
+where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
 %
@@
 this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
+\hl{In addition, as our algorithm performs the Hamming distance calculation, it has potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
+In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required, but when knowledge of nearest vectors is sufficient.}
 

### 2020-05-05 manuscript.tex
solving issue #22
removing highlighted changes for issue #22

--- 2020-05-04 manuscript.tex
@@
 This is achieved by applying the Hadamard gate on auxiliary qubit, 
-followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$ where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
+followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$ 
+where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
 %
@@
 this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
+In addition, as our algorithm performs the Hamming distance calculation, it has potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
+In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required, but when knowledge of nearest vectors is sufficient.
 

### 2020-05-06 manuscript.tex
Fix measurement result dependence on ancila qubit #17

@@
 \begin{multline}
+	\label{eq:final_state}
     \left| \psi_f \right\rangle =  
@@
 
-We thus have the Hamming distances encoded into the amplitudes of the final state. 
-From the statistics of the measurement outcomes, the distance matrix between two data sets of binary vectors can be obtained (Fig.~\ref{fig:distance_matrix}). 
-In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0. 
-If the ancilla qubit is 1, the smallest amplitude of the measurement result coincides with the smallest Hamming distance.
+From the statistics of the measurement outcomes of final state~(\ref{eq:final_state}) we recreate the amplitudes of ancilla qubit states. 
+From those amplitudes estimations we are able to plot the distance matrix between two data sets of binary vectors.
+The probability amplitude of the ancilla qubit outcomes captures the exact Hamming distance as the result of the preprocessing function.
+There are two possible outcomes of measurement of the ancilla qubit, each has own probability amplitude and own interpretation of that amplitude. 
+For instance, for the |0> outcome, the larger the amplitude the smaller the Hamming distance, 
+and for the |1> outcome it is the other way around, magnitude of the amplitude of that outcome is proportional to the Hamming distance.
 

### 2020-05-06 manuscript.tex
resolving issue #26

--- 2020-05-05 manuscript.tex
@@
 
-In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
-our method allows one to reduce the number of gates in the circuit (at least in $2n$ one-qubit NOT gates)
-and reconstruct all distance matrix in parallel and can be implemented with high fidelity.
+Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
+However, for most of algorithms, SOFM included, the information of nearest vectors is sufficient.
+Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
+In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}, but in general case for multiple input vectors the ``Decoding`` stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
+The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001}.
 

### 2020-05-06 manuscript.tex
formatting

@@
 Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
-In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}, but in general case for multiple input vectors the ``Decoding`` stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
+In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}, but in general case for multiple input vectors the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
 The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001}.

### 2020-05-06 manuscript.tex
splitting the argument into multiple paragraphs to improve readability

@@
 Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
-However, for most of algorithms, SOFM included, the information of nearest vectors is sufficient.
+Fortunately, for most of algorithms, SOFM included, the information of nearest vectors is sufficient.
+
 Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
-In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}, but in general case for multiple input vectors the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
+In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}.
+
+In the general case when multiple input vectors are present in the register, the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
 The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001}.

### 2020-05-06 manuscript.tex
multiline style
fix latex style

--- 2020-05-06 manuscript.tex
@@
 
-We thus have the Hamming distances encoded into the amplitudes of the final state. 
-From the statistics of the measurement outcomes, the distance matrix between two data sets of binary vectors can be obtained (Fig.~\ref{fig:distance_matrix}). 
-In this case, the biggest amplitude of the measurement result coincides with the smallest Hamming distance when the measurement result of the ancilla qubit is 0. 
-If the ancilla qubit is 1, the smallest amplitude of the measurement result coincides with the smallest Hamming distance.
+From the statistics of the measurement outcomes of final state~(\ref{eq:final_state}) we recreate the amplitudes of ancilla qubit states. 
+From those amplitudes estimations we are able to plot the distance matrix between two data sets of binary vectors.
+The probability amplitude of the ancilla qubit outcomes captures the exact Hamming distance as the result of the preprocessing function.
+There are two possible outcomes of measurement of the ancilla qubit, each has own probability amplitude and own interpretation of that amplitude. 
+For instance, for the \left| 0 \right\rangle outcome, the larger the amplitude the smaller the Hamming distance, 
+and for the \left| 1 \right\rangle outcome it is the other way around, magnitude of the amplitude of that outcome is proportional to the Hamming distance.
 
@@
 
-Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
-However, for most of algorithms, SOFM included, the information of nearest vectors is sufficient.
-Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
-In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}, but in general case for multiple input vectors the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
-The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001}.
+In comparison with the circuits for Hamming distance calculations proposed previously \cite{trugenberger2001}, 
+our method allows one to reduce the number of gates in the circuit (at least in $2n$ one-qubit NOT gates)
+and reconstruct all distance matrix in parallel and can be implemented with high fidelity.
 

### 2020-05-06 manuscript.tex
Fix latex style and pictures

@@
 
-On a step $t$ when the BMU $w_{c}$ is selected, 
-the weights of the BMU and its neigbours on feature map are adjusted according to
+On a step $t$ when the BMU $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected, 
+the weights $\vec{w}_{i}$ of the BMU and its neigbours on feature map are adjusted according to
 %
@@
     \label{eq:learning}
-	\vec{w_{i}}(t + 1) 
-	= \vec{w_i}(t)
+	\vec{w}_{i}(t + 1) 
+	= \vec{w}_{i}(t)
 	+ \theta(c, i, t) \alpha(t) 
-		\left(\vec{x}(t) - \vec{w_i}(t)\right),
+		\left(\vec{x}(t) - \vec{w}_{i}(t)\right) ,
 \end{equation}
@@
 that ensures each move is following the current trends inferred from their distances to the input objects. 
-A visualisation of this process is shown in Fig. \ref{fig:sofm_fitting}.
+A visualisation of this process is shown in Fig.~\ref{fig:sofm_fitting}.
 
@@
 \begin{align}
+    \label{eq:encodnig}
     \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
     \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
-    \label{eq:encodnig}
 \end{align}
@@
 \begin{equation} 
+    \label{eq:initial_state}
 	\left| \psi_0 \right\rangle = 
@@
     \left| Y \right\rangle 
-    \left| a \right\rangle
-    \label{eq:initial_state}
+    \left| a \right\rangle ,
 \end{equation}
@@
     | y^{(1)}_j, \dots, y^{(n)}_j \rangle
-    | 0 \rangle 
+    | 0 \rangle ,
 \end{equation}
@@
 		0 & 0 & 0 & e^{i \frac \phi 2}
-    \end{pmatrix} .
-    \quad \phi = \frac{\pi}{n}
+    \end{pmatrix} ,
+    \quad \phi = \frac{\pi}{n}.
 \end{equation}
 %
-Finally, another Hadamard gate is applied on the ancilla qubit (see Fig. \ref{fig:qcircuit}).
+Finally, another Hadamard gate is applied on the ancilla qubit (see Fig.~\ref{fig:qcircuit}).
 
@@
 		\left| y_j \right\rangle 
-		\left| 1 \right\rangle
+		\left| 1 \right\rangle .
 \end{multline}
@@
 There are two possible outcomes of measurement of the ancilla qubit, each has own probability amplitude and own interpretation of that amplitude. 
-For instance, for the \left| 0 \right\rangle outcome, the larger the amplitude the smaller the Hamming distance, 
-and for the \left| 1 \right\rangle outcome it is the other way around, magnitude of the amplitude of that outcome is proportional to the Hamming distance.
+For instance, for the $\left| 0 \right\rangle$ outcome, the larger the amplitude the smaller the Hamming distance, 
+and for the $\left| 1 \right\rangle$ outcome it is the other way around, magnitude of the amplitude of that outcome is proportional to the Hamming distance.
 
@@
 
-Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
-In this special case scenario the circuit depth complexity is matching with \cite{shuld2014}.
+Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
+In this special case scenario the circuit depth complexity is matching with \cite{schuld2014}.
 
@@
 In the QASOFM described in the previous section, distance calculations are realized on a quantum device (i.e.
-the IBM Q Experience) with the use of circuit presented in Fig.
-\ref{fig:qcircuit}.
+the IBM Q Experience) with the use of circuit presented in Fig.~\ref{fig:qcircuit}.
 This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates
@@
 
-In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.
-\ref{fig:distance_matrix}.
+In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.~\ref{fig:distance_matrix}.
 We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
-The theoretical calculations and classical simulations show perfect agreement with each other (Fig. \ref{fig:distance_matrix}(a)). 
-The difference between Figs. \ref{fig:distance_matrix} (a) and (b) appears due to noise in the currently available non fault-tolerant quantum processors.
+The theoretical calculations and classical simulations show perfect agreement with each other (Fig.~\ref{fig:distance_matrix}(a)). 
+The difference between Figs.~\ref{fig:distance_matrix} (a) and (b) appears due to noise in the currently available non fault-tolerant quantum processors.
 
@@
 and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
-In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT library.
+In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT\cite{qiskit} library.
 Each epoch of the algorithm consists of distance calculation between all data and cluster vectors 
@@
 The shift is made by the change of the first binary element in the cluster vectors different from the sample one. 
-The evolution of the labels presented on Fig. ~\ref{convergence}(c).  
+The evolution of the labels presented on Fig.~\ref{convergence}(c).  
 Good convergence is already observed in the fourth epoch.
@@
 this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms. 
-In addition, as our algorithm performs the Hamming distance calculation, it has potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
-In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required, but when knowledge of nearest vectors is sufficient.
+In addition, as our algorithm performs the Hamming distance calculation, 
+it has potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form. 
+In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure 
+and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required, 
+but when knowledge of nearest vectors is sufficient.
 
@@
 
-%\begin{thebibliography}{99}
-%\end{thebibliography}
 \end{document}

### 2020-05-06 manuscript.tex
Keep single figure with Ham dist See #8.

@@
 \begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
+	\includegraphics[width=0.95\columnwidth]{distance_matrix.eps}
 	\caption{
-		(a) Hamming distance matrix between two data sets of binary vectors. 
-		Values of distance that are less than the median distance value are marked black. 
-		Classical simulation of the quantum circuit shows perfect agreement to theoretical calculations. 
-		(b) Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
-		We can see good agreement between the distance matrices calculated classically and on the IBM Q Experience.  
+		Hamming distance matrix between two data sets of binary vectors calculated on the IBM Q Experience. 
+		Since the information of closest vectors is sufficient values of distance that are less than 1.5 are marked black. 
+		Errors caused by noise in comparison with theoretical calulation are marked red cross.
+		Result obtained on the IBM Q Experience ``ibmq\_16\_meleourne'' backend. 
+		Fidelity of final state~(\ref{eq:final_state}) measurement is 0.8.
 	}
@@
 Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
-Fortunately, for most of algorithms, SOFM included, the information of nearest vectors is sufficient.
+Fortunately, for most of algorithms, SOFM included, the information of nearest vectors is sufficient 
+and can be obtained with high fidelity~(Fig.~\ref{fig:distance_matrix}).
 

### 2020-05-07 manuscript.tex

@@
 We see good agreement between the distance matrices calculated classically and on the IBM Q Experience. 
-The theoretical calculations and classical simulations show perfect agreement with each other (Fig.~\ref{fig:distance_matrix}(a)). 
-The difference between Figs.~\ref{fig:distance_matrix} (a) and (b) appears due to noise in the currently available non fault-tolerant quantum processors.
+The match is close, but not perfect, and we attribute it to noise in the currently available non fault-tolerant quantum processors.
 

### 2020-05-07 manuscript.tex
Dehighlight Alexey comment #37

@@
 
-At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ \hl{encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis} for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
 %

### 2020-05-08 manuscript.tex
resolving issue #47

@@
 	\caption{
-		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded \cite{trugenberger2001} in $X$ and $Y$ quantum registers   respectively.
+		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded in $X$ and $Y$ quantum registers   respectively.
 		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 

### 2020-05-08 manuscript.tex
resolving issue #49

@@
 In the general case when multiple input vectors are present in the register, the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
-The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001}.
+The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001} but the number of remaining gates is reduced compared to [80], leading to less deep circuit, which is significant for NISQ devices.
 

### 2020-05-08 manuscript.tex
resolving issue #44

@@
 We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of 
-selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
+selecting a winning cluster vector, also called the best matching unit, and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
 More specifically, they consist of four step process: 
@@
 \item selecting an input vector randomly from the set of all input vectors; 
-\item finding a cluster vector which is closest to the input vector (BMU); 
-\item adjusting the weights of the BMU and neurons close to it on feature map in such a way 
+\item finding a cluster vector which is closest to the input vector; 
+\item adjusting the weights of the best matching unit and neurons close to it on feature map in such a way 
 	that these vectors  becomes even closer to the input vector; 
@@
 
-On a step $t$ when the BMU $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected, 
-the weights $\vec{w}_{i}$ of the BMU and its neigbours on feature map are adjusted according to
+On a step $t$ when the best matching unit $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected, 
+the weights $\vec{w}_{i}$ of the best matching unit and its neigbours on feature map are adjusted according to
 %

### 2020-09-15 manuscript.tex
Improve by Alexey
Sorry for the delay, I needed to take a break in order to think and take
a look from another point of view but unfortunately, it did not help so
much. Marek and Iliya, thanks for your comments. Actually, I didn't want
you to send us the definitions, references, and other information from
Wikipedia which I know and understand not worse than you but wanted some
explanation why it is better in comparison with the previous version. As
for section 2 A, basically I prefer the simpler adjusting rule from the
previous version but if everybody prefers the variant from Wikipedia I
am o

@@
 \begin{document}
-\title{Quantum-assisted unsupervised data clustering}
+\title{Hybrid quantum-classical unsupervised data clustering on the basis of Self-Organizing Feature Map}
 
@@
 	We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience 
-	and show that it allows us to reduce number of calculations in a number of clusters. 
+	and show that it allows us to reduce the number of calculations in a number of clusters. 
 	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.  
@@
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-Paradoxically, the exact mechanism of why neural networks are so powerful remain unknown (in many cases it is regarded as a black box).  
-It has been speculated that limits to the neural network approach based on the computational power of von Neumann architecture are being approached, 
+Paradoxically, the intrinsic mechanism of how neural networks work and why are so powerful remains unknown (in many cases it is regarded as a black box).  
+It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached, 
 and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
@@
 SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016} 
-and in comparison with many other artificial neural networks they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}. 
-The SOFMs with a small number of nodes are similar to the K-means algorithm 
-but for larger SOFMs they represent data in a fundamentally topological way that allows one to do dimensionality reduction.  
+and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}. 
+The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.  
 Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
@@
 Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}. 
-The main focus of early works in quantum machine learning (QML) was on the use of quantum computers to perform basic linear algebra subroutines with a quantum speedup (such as Fourier transform and solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}) by utilizing quantum phenomena such as quantum entanglement and quantum superposition \cite{biamonte2017, schuld2014}.
-Quantum versions were developed for linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
+The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving linear equations \cite{wiebe2012,harrow2009,childs2017}.
+Quantum algorithms were developed to perform  linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
 More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019, broughton2020}. 
@@
 which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
+Quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}.
 Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}.
-The quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}. 
-Quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}. 
-Sublinear quantum algorithms for training linear and kernel-based classifiers were designed \cite{li2019}. 
-Furthermore, continuous-variable quantum neural networks \cite{killoran2019}, 
-quantum autoencoders \cite{bondarenko2019}, 
-quantum approaches for reinforcement learning \cite{dunjko2017, nautrup2019, foesel2018} 
-and some other \cite{rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019} were developed. 
+For the problem of classification which is closely connected to the problem of clusterization, a protocol of quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
+Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
 At the same time, non-neural-network based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}.
@@
 
-In this paper, we analyze  a hybrid quantum assisted SOFM (QASOFM) 
-and apply it to the data clustering problem in an unsupervised manner on a toy example of clustering paper abstracts.   
-We implement the quantum assisted SOFM in such a way 
-that it becomes possible to use the Hamming distance as a distance metric for the training 
-and to reduce the number of distance calculations in the number of clusters. 
-We propose an optimized circuit for realizing the Hamming distance on a quantum machine. 
-We then apply it to a toy example of data for clustering paper abstracts 
-and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq} 
+In this paper, we develop a hybrid quantum assisted SOFM (QASOFM) 
+and apply it to the data clustering problem in an unsupervised manner.
+The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol with assistance of quantum computational machines.
+In order to make our protocol more appropriate for the currently available NISQ quantum devices, we optimized a circuit for realizing the Hamming distance with reducing number of one-qubit operations.
+We then apply it to a toy example of clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq}
 and compare it to the classical case.
@@
 		Schematic illustration of the clustering problem considered in this paper.  
-		Blue dots represent clusters and red dots are data points. 
+		Blue triangles represent clusters and red dots are data points. 
 		The training process moves clusters to fit the data points. 
@@
 The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector, 
-and describes a mapping from a higher-dimensional input space to a lower-dimensional map space, commonly a bi-dimensional map.
+and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserve the topological properties of the input space, commonly a bi-dimensional map.
 
@@
 The components of these vectors are usually called weights. 
-Initially the weight components are chosen randomly 
-and topological distances between neurons given. 
+Initially the weight components are chosen randomly. 
 We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of 
@@
 On a step $t$ when the best matching unit $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected, 
-the weights $\vec{w}_{i}$ of the best matching unit and its neigbours on feature map are adjusted according to
+the weights $\vec{w}_{i}$ of the best matching unit and its neighbors on feature map are adjusted according to
 %
@@
 %
-where  $\alpha(t)$ is the learning rate and $\theta(c, i, t)$ is the neighbor function,
-which defines the topological neighbor neurons to be updated. 
-Note that neighbor function depend on distance on feature map given initialy, 
-but not distance metrics between vectors.
+where  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually Gaussian, Delta or the mexican-hat function) which defines the vicinity of the BMU noted by index $c$ in which the weights also should be adjusted and depends on distance on feature map.
+This expression can be interpreted according to: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
+The larger the difference between the input component and the weight component, the larger the increment (decrement).
 
-Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors in space one at a time in a way 
-that ensures each move is following the current trends inferred from their distances to the input objects. 
+Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (blue triangles on Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (red dots on Fig.~\ref{fig:sofm_fitting}).
 A visualisation of this process is shown in Fig.~\ref{fig:sofm_fitting}.
 
-In SOFM original version was designed to cluster real-valued data
-and the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors.
-The paper deals with binary vectors clustering problem 
-and for binary data the Hamming distance is more suitable \cite{appiah2009, santana2017}.
-Using a known technique of encoding classical information into a quantum register, 
-based on probabilistic quantum memories \cite{trugenberger2001},
-we introduce an optimized algorithm for calculating the matrix of Hamming distances 
-between each pair of binary vectors  of two sets taking advantage of quantum parallelism.
+In the original version of SOFM, the winning cluster vector was selected based on the Euclidean distance between an input vector and the cluster vectors.
+In this paper we deal with binary vectors clustering problem and for binary data the Hamming distance is more suitable \cite{appiah2009, santana2017}.
+Using a simple technique of encoding classical binary information into a quantum register \cite{trugenberger2001},
+we introduce an optimized algorithm for calculating the matrix of Hamming distances between all input and cluster vectors at once.
 
@@
 	\includegraphics[width=\columnwidth]{qcircuit.eps}
-	\caption{
+	\caption{ \hl{Iliya, could you please change the picture to the case with the controlled phase gate and make correction of the caption in appropriate way}
 		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded in $X$ and $Y$ quantum registers   respectively.
 		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 
-		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate~(\ref{eq:controled_rotation}) and Hadamard gates. 
+		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate(see Eq.~(\ref{eq:controled_rotation})) and Hadamard gates. 
 		Finally, a quantum state of the $X$-register returned to the initial basis for information retrieval. 	
@@
 \begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{distance_matrix.eps}
-	\caption{
-		Hamming distance matrix between two data sets of binary vectors calculated on the IBM Q Experience. 
-		Since the information of closest vectors is sufficient values of distance that are less than 1.5 are marked black. 
-		Errors caused by noise in comparison with theoretical calulation are marked red cross.
-		Result obtained on the IBM Q Experience ``ibmq\_16\_meleourne'' backend. 
-		Fidelity of final state~(\ref{eq:final_state}) measurement is 0.8.
+	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
+	\caption{\hl{[AP: Iliya, could you please recalculate the fidelity for this picture it looks that it is even higher for thee picture because there is basically less red crosses and rewrite the caption in the appropriate way]}
+		Hamming distance matrix between two data sets of binary vectors. 
+		Values of distance that less than median distance value was marked black.
+The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
+Result obtained on  IBMQ "ibmq\_16\_melbourne" backend is shown on the right figure.
+		Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
+		Fidelity of final state~(\ref{eq:final_state}) measurement is 0.8.}
 	}
@@
 During the whole process, the $\left| Y \right\rangle$ register is used to store the cluster states.  
-At the beginning and end of the procedure the $\left| X \right\rangle$ register stores the input vectors.  
-Meanwhile, during the procedure it stores the differences between input vectors and cluster states.
+At the beginning and end of the procedure, the $\left| X \right\rangle$ register stores the input vectors.  
+During the procedure it stores the differences between input vectors and cluster states.
 
-Let us assume we have $k$ input vectors and $l$ cluster states. 
+Let us assume we have $N_x$ input vectors and $N_y$ cluster states. 
 The $i$th input vector and $j$th cluster vector are respectively denoted as $\left| x_i \right\rangle$, $\left| y_j \right\rangle$. 
@@
     \label{eq:encodnig}
-    \left| X \right\rangle  & = \frac{1}{\sqrt{k}} \sum\limits_{i=1}^{k} \left| x_i \right\rangle,  \\
-    \left| Y \right\rangle&  = \frac{1}{\sqrt{l}} \sum\limits_{j=1}^{l} \left| y_j \right\rangle .
+    \left| X \right\rangle  & = \frac{1}{\sqrt{N_x}} \sum\limits_{i=1}^{N_x} \left| x_i \right\rangle,  \\
+    \left| Y \right\rangle&  = \frac{1}{\sqrt{N_y}} \sum\limits_{j=1}^{N_y} \left| y_j \right\rangle .
 \end{align}
@@
 
-Given this initial state we may begin the processing of the problem. We start by applying a CNOT gate on $\left| x^{(\alpha)} y^{(\alpha)} \right\rangle$ and $\alpha = 1 \dots n$
+Given this initial state we may begin the processing of the problem.
+The quantum circuit that we follow is shown in Fig.\ref{fig:qcircuit}.
+We start by elementwise applying CNOT gates between all the qubits of the X and Y registers (the "Encoding" part of Fig.
+\ref{fig:qcircuit}) $\left| x^{(\alpha)} y^{(\alpha)} \right\rangle$
 
@@
     | \psi_1 \rangle  =  
-    \frac{1}{\sqrt{kl}} \sum_{i, j=1}^{k} 
+    \frac{1}{\sqrt{N_x N_y}} \sum_{i, j=1}^{N_x,N_y} 
     | d^{(1)}_{ij}, \dots, d^{(n)}_{ij} \rangle 
@@
 At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors,
-instead it stores the information about pairwise different qubits between the input vector $\{X\}$ and cluster vector $\{Y\}$. 
-Next, for each pair $\{X\}$ and $\{Y\}$, the accumulated information of all the differences is projected onto the amplitude of the superposed state. 
-This is achieved by applying the Hadamard gate on auxiliary qubit, 
-followed by a controlled rotation around z-axis gate~(\ref{eq:controled_rotation}) on $\left|a d_{ij}^{(\alpha)}\right\rangle$ 
-where $d_{ij}^{(\alpha)}$ is the control qubit and $\left| a \right\rangle$ ancilla qubit register is the target.
-%
+instead it stores the information about pairwise different qubits between the input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$.
+
+The next step, depicted on the Fig.\ref{fig:qcircuit} as the "Extraction" stage, is to extract the accumulated information about the differences between each pair $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ by projection onto the amplitude of the superposed state.
+This is achieved by applying the Hadamard gate on auxiliary qubit, followed by a controlled phase gate on $\left|a\right\rangle \left|d_{ij}^{(\alpha)}\right\rangle$
+%~(\ref{eq:control_phase_rotation})
 \begin{equation}
-    \label{eq:controled_rotation}
-    C_{R_z}(\phi) = 
+    \label{eq:control_phase_rotation}
+    R_{(X,a)}(\phi) = 
     \begin{pmatrix}
         1 & 0 & 0 & 0 \\
-        0 & e^{-i \frac \phi 2} & 0 & 0 \\
+        0 & 1 & 0 & 0 \\
         0 & 0 & 1 & 0 \\
-		0 & 0 & 0 & e^{i \frac \phi 2}
+        0 & 0 & 0 & e^{-i\phi}
     \end{pmatrix} ,
-    \quad \phi = \frac{\pi}{n}.
+    \quad \phi = \frac{\pi}{n}
 \end{equation}
 %
+where $d_{ij}^{(\alpha)}$ is the control qubit and the ancilla qubit $\left| a \right\rangle$ is the target.
 Finally, another Hadamard gate is applied on the ancilla qubit (see Fig.~\ref{fig:qcircuit}).
@@
     \left| \psi_2 \right\rangle = 
-    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
+    \frac{1}{\sqrt{N_x N_y}}\sum\limits_{i, j=1}^{N_x,N_y} 
     \left| d_{ij} \right\rangle 
     \left| y_j \right\rangle
-    \dfrac{(\left| 0 \right\rangle + \left| 1 \right\rangle)}{\sqrt{2}}  .
+    \dfrac{(\left| 0 \right\rangle + \left| 1 \right\rangle)}{\sqrt{2}},
 \end{equation}
 %
-Applying the controlled rotation around $z$-axis gate on $\left| x^{(\alpha)} a \right\rangle$ where $\alpha = 1\dots n$ and the state then becomes
+where $\left| d_{ij} \right\rangle = \left| d_{ij}^{(1)},\ldots,d_{ij}^{(n)} \right\rangle$.
+Applying the controlled phase gate on $\left| x^{(\alpha)} a \right\rangle$ where $\alpha = 1\dots n$ and the state then becomes
 %
 \begin{multline}
-    \left| \psi_3 \right\rangle  
-	= \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k} 
-		\exp\left(
-		    \dfrac{-i \pi}{2n}
-		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
-		\right)
+    \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
+    \\ = \dfrac{1}{\sqrt{2 N_x N_y}}
+				\sum\limits_{i, j=1}^{N_x,N_y} 
 				\left| d_{ij} \right\rangle 
         \left| y_j \right\rangle 
-		\left| 0 \right\rangle \\
-    + \dfrac{1}{\sqrt{2kl}} \sum\limits_{i, j=1}^{k}
-		\exp\left(
-		    \dfrac{i \pi}{2n}
-		    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
-		\right)
+        \left| 0 \right\rangle
+        \\ + \dfrac{1}{\sqrt{2 N_x N_y}}
+				\sum\limits_{i, j=1}^{N_x,N_y}
+        \exp\left(\dfrac{-i \pi}{n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| d_{ij} \right\rangle 
-		\left| y_j \right\rangle 
-		\left| 1 \right\rangle .
+\\ \times        \left| y_j \right\rangle 
+        \left| 1 \right\rangle
 \end{multline}
@@
     \left| \psi_4 \right\rangle = 
-    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
-        \left[ 
-			\cos\left(
-			    \dfrac{\pi}{2n}
-			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij}
-			\right)
+    \frac{1}{\sqrt{N_x N_y}}\sum\limits_{i, j=1}^{N_x,N_y} 
+    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+		\\ \times
+        \left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| d_{ij} \right\rangle 
@@
         \\+ 
-			\left. i \sin\left(
-			    \dfrac{\pi}{2n} 
-			    \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
-			\right)
+        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| d_{ij} \right\rangle 
         \left| y_j \right\rangle 
-			\left| 1 \right\rangle
-		\right] .
+        \left| 1 \right\rangle\right] .
 \end{multline}
 %
-This completes the step for projecting differences between pairs of $\{X\}$ and $\{Y\}$ onto the amplitude of the auxiliary qubit. 
+This completes the step for projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit. 
 The process is done in the $x$-basis, achieved by the surrounding Hadamard gates. 
 There are two possible measurement outcomes of the auxiliary qubit.  
-Each pair of $\{X\}$ and $\{Y\}$ forms a subspace of the Hilbert space, 
-the controlled rotation around $z$-axis gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\{X\}$ and $\{Y\}$ are.
+Each pair of $\left|x_i \right\rangle$ and $\left| y_j \right\rangle$ forms a subspace of the Hilbert space, 
+the controlled phase gate ensures to change amplitudes of those outcomes within this subspace depending on how different the spin configurations between $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
 
-At this stage, the information regarding the differences between pairs of $\{X\}$ and $\{Y\}$ encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+At this stage, the information regarding the differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
 %
@@
     \left| \psi_f \right\rangle = 
-    \frac{1}{\sqrt{kl}}\sum\limits_{i, j=1}^{k} 
-	\left[ 
-	    \cos\left(
-	        \dfrac{\pi}{2n}
-	        \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
-	    \right)
+    \mathrm{CNOT} (Y,X)\left| \psi_4 \right\rangle \\=  
+    \sum\limits_{i, j=1}^{N_x,N_y} 
+    \exp \left(\dfrac{-i \pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
+				\left[ \cos\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| x_i \right\rangle 
@@
         \\+
-        \left. i \sin\left(
-            \dfrac{\pi}{2n}
-            \sum\limits_{\alpha=1}^n d^{(\alpha)}_{ij} 
-        \right)
+        \left. i \sin\left(\dfrac{\pi}{2n}\sum\limits_{l=1}^n d^{(l)}_{ij} \right)
         \left| x_i \right\rangle 
         \left| y_j \right\rangle 
-        \left| 1 \right\rangle
-    \right] .
+        \left| 1 \right\rangle\right] .
 \end{multline}
 %
-This makes $\{X\}$ store the input vectors again, as in the initial step 
-and preserves the amplitudes of the auxiliary qubit which are proportional to how different each pairs of $\{X\}$ and $\{Y\}$ are.
+Thus we return int the initial basis and the amplitudes of the auxiliary qubit are proportional to how different each pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
 
@@
 The probability amplitude of the ancilla qubit outcomes captures the exact Hamming distance as the result of the preprocessing function.
-There are two possible outcomes of measurement of the ancilla qubit, each has own probability amplitude and own interpretation of that amplitude. 
+There are two possible outcomes of measurement of the ancilla qubit, each has its probability amplitude and own interpretation of that amplitude. 
 For instance, for the $\left| 0 \right\rangle$ outcome, the larger the amplitude the smaller the Hamming distance, 
-and for the $\left| 1 \right\rangle$ outcome it is the other way around, magnitude of the amplitude of that outcome is proportional to the Hamming distance.
+and for the $\left| 1 \right\rangle$ outcome it is the other way around, the magnitude of the amplitude of that outcome is proportional to the Hamming distance.
 
-Measuring the Hamming distance of a particular pair of input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form,
+Measuring the Hamming distance of a particular pair of input vectors $\left| x_i \right\rangle$ and cluster vectors $\left| y_j \right\rangle$ consists of extracting the relevant amplitude from the subspace that those states form,
 this can be done using the following projection operator
@@
 \begin{align}
-    \rho_{i,j} &= \text{Tr}_{1,\dots,2n} (\Pi_{i,j} \left| \psi_f \rangle\langle \psi_f \right| \Pi_{i,j}) .
+    \rho_{i,j} &= \text{Tr}_{X,Y} (\Pi_{i,j} \left| \psi_f \rangle\langle \psi_f \right| \Pi_{i,j}) .
 \end{align}
 %
-From the state the following two amplitudes for the measurement results can be extracted
+From the reduced density matrix, the following two amplitudes for the measurement results can be extracted
 %
@@
 %
-In order to reduce noise we average the measurement results over different states of the ancilla qubit, 
-thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
+Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
+In order to reduce noise we average the measurement results over different states of the ancillary qubit, thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
 %
@@
 
-Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
-Fortunately, for most of algorithms, SOFM included, the information of nearest vectors is sufficient 
-and can be obtained with high fidelity~(Fig.~\ref{fig:distance_matrix}).
+The number of controlled gate operations to define the full distance matrix matches the number of controlled gate operations in \cite{trugenberger2001}, where the original algorithm for calculation of the Humming distance was introduced, but the number of remaining gates is reduced compared to \cite{trugenberger2001}, leading to less deep circuit, which is significant for NISQ devices.
 
-Because of that property, those algorithms fall into a special case category for which, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
-In this special case scenario the circuit depth complexity is matching with \cite{schuld2014}.
-
-In the general case when multiple input vectors are present in the register, the ``Decoding'' stage still needs to be included leading to larger circuit depth and less attractive complexity in terms of number of controlled gate operations.
-The number of controlled gate operations in this general case of multiple input vectors is matching the number of controlled gate operations in \cite{trugenberger2001} but the number of remaining gates is reduced compared to [80], leading to less deep circuit, which is significant for NISQ devices.
+%In this special case scenario the circuit depth complexity is matching with \cite{schuld2014}.
 
@@
 		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis. 
-		The samples are sorted into groups (QML, MED, BIO) with 3 papers for each tag, for a total of 9 paper.    
+		The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.    
 		(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix. 
@@
 and biology \cite{bio0, bio1, bio2}. 
-Each set consists of three random papers that focus on one of following topics: 
-``Quantum Machine Learning'' (QML), 
-``Cancer'' (MED) 
-and ``Gene Expression'' (BIO). 
-Abstracts were vectorized by the bag-of-words \cite{weikang2016} model in order to choose most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
-This model represents text as a multiset ``bag'' of its words taking into account only multiplicity of words. 
+Each set consists of three papers selected at random that focus on one of following topics: 
+``Quantum Machine Learning'', 
+``Cancer'' 
+and ``Gene Expression''. 
+Abstracts were vectorized by the bag-of-words \cite{weikang2016} model in order to choose the most defining words in each data set (see Fig.~\ref{fig:vectorized_sample}) \cite{mctear2016}.  
+This model represents text as a multiset ``bag'' of its words taking into account only the multiplicity of words. 
 Preparing the bag-of-words we excluded the words that appear only in one abstract and more than in 4 abstracts and we also excluded the word ``level'' from consideration due to the frequent overlap between the clusters because it gives instabilities for both classical and quantum algorithms.
-We restricted our bag-of-word size to 9 of the most frequent words from the full bags-of-word  due to limitations of the number of qubits. 
-After vectorizing and pre-processing  the data, the clusters are well separated with the Hamming distance.  
+We restricted our bag-of-word size to 9 of the most frequent words from the full bags-of-word  due to limitations on the number of qubits. 
+After vectorizing and pre-processing  the data, the clusters are well-separated with the Hamming distance.  
 We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters, 
@@
 
-In classical SOFM \cite{kohonen1990}, first a distance calculation between a a sample vector and all cluster vectors is made, 
+In classical SOFM \cite{kohonen1990}, first a distance calculation between a sample vector and all cluster vectors is made, 
 then the closest cluster vector is shifted towards the sample vector. 
@@
 that are possible to realize on  quantum computing devices that are currently available. 
-The circuit realized in such a manner 
-that calculation of Hamming distance between the sample vector and all cluster vectors is realized in one operation. 
+The circuit realized in this way such 
+that the calculation of Hamming distance implemented step-by-step between each sample vector and all cluster vectors is realized in one operation.
+In this case, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
 The complexity of the quantum assisted SOFM then scales as $O(LN)$.
@@
 An example of the QASOFM learning process is shown in Fig.~\ref{convergence}. 
-Initially, the cluster vectors were randomly chosen (see Fig.~\ref{convergence}) 
-and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
+Initially, the cluster vectors were randomly chosen and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
 In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT\cite{qiskit} library.
@@
 In addition, as our algorithm performs the Hamming distance calculation, 
-it has potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form. 
+it has the potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form. 
 In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure 

### 2020-09-15 manuscript.tex
fix spaces
Apply Tim comments

@@
 \begin{document}
-\title{Hybrid quantum-classical unsupervised data clustering on the basis of Self-Organizing Feature Map}
+\title{Hybrid quantum-classical unsupervised data clustering based on the Self-Organizing Feature Map}
 
@@
 	Quantum computers offer opportunities to speed up such machine learning techniques.
-	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map artificial neural networks. 
-	We make a proof-of-concept realization of one of the central components of the approach on the IBM Q Experience 
+	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural networks.
+	We make a proof-of-concept realization of one of the central components on the IBM Q Experience
 	and show that it allows us to reduce the number of calculations in a number of clusters.
@@
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-Paradoxically, the intrinsic mechanism of how neural networks work and why are so powerful remains unknown (in many cases it is regarded as a black box).  
+Paradoxically, the intrinsic mechanism of how neural networks work and why they are so powerful remains unknown (in many cases it is regarded as a black box).
 It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,
@@
 where the parameters behave much like the weights of a neural network \cite{lewenstein1994}.
-In particular, quantum algorithms for training and evaluating feedforward neural networks was developed, 
+In particular, quantum algorithms for training and evaluating feedforward neural networks were developed,
 which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
@@
 Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
+
 At the same time, non-neural-network based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}.
-Such hybrid algorithms involve as a part some quantum circuits for speed-up and usually trained in a standard classical learning loop. 
+Such hybrid algorithms involve quantum circuits used for part of the execution and is typically trained in a classical learning loop.
 In particular, the Quantum Approximate Optimization Algorithm was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016}
-and designed to problems such as MAX-CUT, Grover’s search algorithm and so on \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019}. 
-Another example of well-known hybrid quantum classical algorithms is Variational Quantum Eigensolver (VQE) for applications in quantum chemistry 
+and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019}.
+Another example of a well-known hybrid quantum classical algorithm is the Variational Quantum Eigensolver (VQE) for applications in quantum chemistry
 \cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
@@
 
-In this paper, we develop a hybrid quantum assisted SOFM (QASOFM) 
+In this paper, we develop a hybrid quantum-assisted SOFM (QASOFM)
 and apply it to the data clustering problem in an unsupervised manner.
-The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol with assistance of quantum computational machines.
-In order to make our protocol more appropriate for the currently available NISQ quantum devices, we optimized a circuit for realizing the Hamming distance with reducing number of one-qubit operations.
+The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
+In order to make our protocol more appropriate for the currently available generation of NISQ quantum devices, we optimized a circuit for realizing the Hamming distance by reducing number of one-qubit operations.
 We then apply it to a toy example of clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq}
@@
 The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector,
-and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserve the topological properties of the input space, commonly a bi-dimensional map.
+and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserves the topological properties of the input space, commonly a bi-dimensional map.
 
@@
 Initially the weight components are chosen randomly.
-We then can train our SOFM adjusting the components through the learning process which occur in the two basic procedures of 
-selecting a winning cluster vector, also called the best matching unit, and updating its weights (Fig.~\ref{fig:sofm_fitting}). 
+We then train our SOFM by adjusting the components through the learning process which occur in the two basic procedures of
+selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}).
 More specifically, they consist of four step process:
@@
 %
-where  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually Gaussian, Delta or the mexican-hat function) which defines the vicinity of the BMU noted by index $c$ in which the weights also should be adjusted and depends on distance on feature map.
+where  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually Gaussian or Delta) which defines the vicinity of the BMU noted by index $c$ in which the weights also should be adjusted and depends on distance on feature map.
 This expression can be interpreted according to: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
@@
 Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (blue triangles on Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (red dots on Fig.~\ref{fig:sofm_fitting}).
-A visualisation of this process is shown in Fig.~\ref{fig:sofm_fitting}.
 
 In the original version of SOFM, the winning cluster vector was selected based on the Euclidean distance between an input vector and the cluster vectors.
-In this paper we deal with binary vectors clustering problem and for binary data the Hamming distance is more suitable \cite{appiah2009, santana2017}.
+In this paper we deal with a binary vector clustering problem, where the Hamming distance is more suitable \cite{appiah2009, santana2017}.
 Using a simple technique of encoding classical binary information into a quantum register \cite{trugenberger2001},
@@
 	\includegraphics[width=\columnwidth]{qcircuit.eps}
-	\caption{ \hl{Iliya, could you please change the picture to the case with the controlled phase gate and make correction of the caption in appropriate way}
-		Quantum circuit for the quantum parallelized Hamming distance calculating  between all pairs of binary vectors from two sets ${X}$ and ${Y}$ encoded in $X$ and $Y$ quantum registers   respectively.
-		First, we encoded information about pairwise different qubits in a quantum state of the $X$-register with applying the CNOT gates. 
-		Second, Hamming distance values are extracted into the amplitudes of superposition with the controled rotation around $z$-axis gate(see Eq.~(\ref{eq:controled_rotation})) and Hadamard gates. 
-		Finally, a quantum state of the $X$-register returned to the initial basis for information retrieval. 	
+    \caption{
+		Quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of binary vectors from two sets ${X}$ and ${Y}$.
+		First, we encode information about pairwise different qubits in a quantum state of the $X$-register by applying the CNOT gates.
+		Second, the Hamming distance values are extracted to the amplitudes of superposition with the control phase rotation (see Eq.~(\ref{eq:controled_phase_rotation})) and Hadamard gates.
+		Finally, a quantum space of the $X$-register is returned to the initial basis for information retrieval.
 	}
@@
 	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
-	\caption{\hl{[AP: Iliya, could you please recalculate the fidelity for this picture it looks that it is even higher for thee picture because there is basically less red crosses and rewrite the caption in the appropriate way]}
+	\caption{
 		Hamming distance matrix between two data sets of binary vectors.
@@
 		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
-Result obtained on  IBMQ "ibmq\_16\_melbourne" backend is shown on the right figure.
-		Result obtained on the IBM Q Experience ``ibmq\_16\_melbourne'' backend. 
-		Fidelity of final state~(\ref{eq:final_state}) measurement is 0.8.}
+		Result obtained on  IBMQ ``ibmq\_16\_melbourne'' backend is shown on the right figure.
 	}
@@
     \left| Y \right\rangle
-    \left| a \right\rangle ,
+    \left| 0 \right\rangle.
 \end{equation}
-%
-where $\left| a \right\rangle$ is an auxiliary qubit in the state $\left| 0 \right\rangle$ initially.
 
@@
 The quantum circuit that we follow is shown in Fig.\ref{fig:qcircuit}.
-We start by elementwise applying CNOT gates between all the qubits of the X and Y registers (the "Encoding" part of Fig.
-\ref{fig:qcircuit}) $\left| x^{(\alpha)} y^{(\alpha)} \right\rangle$
-
+We start by elementwise applying CNOT gates between all the qubits of the $X$ and $Y$ registers (the ``Encoding'' part of Fig.
+\ref{fig:qcircuit})
+%
 \begin{equation}
@@
 
-The next step, depicted on the Fig.\ref{fig:qcircuit} as the "Extraction" stage, is to extract the accumulated information about the differences between each pair $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ by projection onto the amplitude of the superposed state.
-This is achieved by applying the Hadamard gate on auxiliary qubit, followed by a controlled phase gate on $\left|a\right\rangle \left|d_{ij}^{(\alpha)}\right\rangle$
+The next step, depicted on the Fig.\ref{fig:qcircuit} as the ``Extraction'' stage, is to extract the accumulated information about the differences between each pair $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ by projection onto the amplitude of the superposed state.
+This is achieved by applying the Hadamard gate on the auxiliary qubit, followed by a controlled phase gate on $\left|a\right\rangle \left|d_{ij}^{(\alpha)}\right\rangle$, where the phase gate is defined by
 %~(\ref{eq:control_phase_rotation})
 \begin{equation}
-    \label{eq:control_phase_rotation}
+    \label{eq:controled_phase_rotation}
     R_{(X,a)}(\phi) =
@@
 where $\left| d_{ij} \right\rangle = \left| d_{ij}^{(1)},\ldots,d_{ij}^{(n)} \right\rangle$.
-Applying the controlled phase gate on $\left| x^{(\alpha)} a \right\rangle$ where $\alpha = 1\dots n$ and the state then becomes
+Applying the controlled phase gate on $\left| x^{(\alpha)} a \right\rangle$, where $\alpha = 1\dots n$  the state then becomes
 %

### 2020-09-17 manuscript.tex
FIX: english improvements

@@
 \begin{abstract}
-	Unsupervised machine learning is one of the main techniques in artificial intelligence.
+	Unsupervised machine learning is one of the main techniques employed in artificial intelligence.
 	Quantum computers offer opportunities to speed up such machine learning techniques.
-	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural networks.
+	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural network.
 	We make a proof-of-concept realization of one of the central components on the IBM Q Experience
@@
 Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}.
-The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving linear equations \cite{wiebe2012,harrow2009,childs2017}.
-Quantum algorithms were developed to perform  linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
+The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
+Quantum algorithms were developed for  linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
 More recently, there has been attention on developing quantum neural networks \cite{kamruzzaman2019, schuld2014b, jeswal2019, broughton2020}.
@@
 In particular, the Quantum Approximate Optimization Algorithm was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016}
-and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019}.
-Another example of a well-known hybrid quantum classical algorithm is the Variational Quantum Eigensolver (VQE) for applications in quantum chemistry
+and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019,byrnes2018}.
+Another example of a well-known hybrid quantum classical algorithm is the Variational Quantum Eigensolver (VQE) for applications in quantum simulations
 \cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
-Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices.
+Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices \cite{preskill2018}.
 
@@
 		Schematic illustration of the clustering problem considered in this paper.
-		Blue triangles represent clusters and red dots are data points.
+		Blue triangles represent clusters and red circles are data points.
 		The training process moves clusters to fit the data points.
@@
 The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector,
-and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserves the topological properties of the input space, commonly a bi-dimensional map.
+and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserves the topological properties of the input space, commonly referred to as a bi-dimensional map.
 
@@
 Initially the weight components are chosen randomly.
-We then train our SOFM by adjusting the components through the learning process which occur in the two basic procedures of
+The SOFM is then trained by adjusting the components through the learning process which occur in the two basic procedures of
 selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}).
-More specifically, they consist of four step process:
+More specifically, they consist of the four step process:
 \begin{enumerate*}
-\item selecting an input vector randomly from the set of all input vectors;
-\item finding a cluster vector which is closest to the input vector;
-\item adjusting the weights of the best matching unit and neurons close to it on feature map in such a way
-	that these vectors  becomes even closer to the input vector;
-\item repeating this process for many iterations until it converges.
+\item select an input vector randomly from the set of all input vectors;
+\item find a cluster vector which is closest to the input vector;
+\item adjust the weights of the best matching unit and neurons close to it on the feature map in such a way
+	that these vectors  becomes closer to the input vector;
+\item repeat this process for many iterations until it converges.
 \end{enumerate*}
@@
 On a step $t$ when the best matching unit $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected,
-the weights $\vec{w}_{i}$ of the best matching unit and its neighbors on feature map are adjusted according to
+the weights $\vec{w}_{i}$ of the best matching unit and its neighbors on the feature map are adjusted according to
 %
@@
 	+ \theta(c, i, t) \alpha(t)
-		\left(\vec{x}(t) - \vec{w}_{i}(t)\right) ,
+		\left(\vec{x}(t) - \vec{w}_{i}(t)\right) .
 \end{equation}
 %
-where  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually Gaussian or Delta) which defines the vicinity of the BMU noted by index $c$ in which the weights also should be adjusted and depends on distance on feature map.
-This expression can be interpreted according to: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
+Here,  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually taken as a Gaussian or delta function) which defines the vicinity of the BMU labeled by an index $c$, the weights of the neighbors in the vicinity should also be adjusted in the same manner as for the BMU.
+This expression can be understood in the following way: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
 The larger the difference between the input component and the weight component, the larger the increment (decrement).
 
-Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (blue triangles on Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (red dots on Fig.~\ref{fig:sofm_fitting}).
+Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (blue triangles on Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (red circles on Fig.~\ref{fig:sofm_fitting}).
 
-In the original version of SOFM, the winning cluster vector was selected based on the Euclidean distance between an input vector and the cluster vectors.
-In this paper we deal with a binary vector clustering problem, where the Hamming distance is more suitable \cite{appiah2009, santana2017}.
+In the original version of SOFM, the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors.
+In this paper, we deal with a binary vector clustering problem, where the Hamming distance is more suitable \cite{appiah2009, santana2017}.
 Using a simple technique of encoding classical binary information into a quantum register \cite{trugenberger2001},
@@
 		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
-		Result obtained on  IBMQ ``ibmq\_16\_melbourne'' backend is shown on the right figure.
+		Result obtained on  IBMQ ``ibmq\_16\_melbourne'' backend is shown on the right figure and the accuracy in comparison with the classical simulation is more than 92 $\%$.
 	}
@@
     \left| \psi_3 \right\rangle = R_{(X,a)}\left(\dfrac{\pi}{n}\right)\left| \psi_2 \right\rangle
-    \\ = \dfrac{1}{\sqrt{2 N_x N_y}}
+     = \dfrac{1}{\sqrt{2 N_x N_y}}
 				\sum\limits_{i, j=1}^{N_x,N_y}
@@
         \left| d_{ij} \right\rangle
-\\ \times        \left| y_j \right\rangle
-        \left| 1 \right\rangle
+        \left| y_j \right\rangle
+        \left| 1 \right\rangle .
 \end{multline}
@@
 %
-This completes the step for projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit.
+This completes the step of projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit.
 The process is done in the $x$-basis, achieved by the surrounding Hadamard gates.
@@
 
-At this stage, the information regarding the differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ encoded in the amplitudes, in order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
+At this stage, the information regarding the differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ is encoded in the amplitudes.
+In order to extract the Hamming distances between the relevant $\left| x_i \right\rangle$, $\left| y_j \right\rangle$ we return to our initial basis for register $\left| X \right\rangle$ by applying pairwise CNOT gates:
 %
@@
 %
-Thus we return int the initial basis and the amplitudes of the auxiliary qubit are proportional to how different each pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
+Thus we return into the initial basis and the amplitudes of the auxiliary qubit are proportional to how different each pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
 
-From the statistics of the measurement outcomes of final state~(\ref{eq:final_state}) we recreate the amplitudes of ancilla qubit states.
+From the statistics of the measurement outcomes of the final state~(\ref{eq:final_state}) we recreate the amplitudes of ancilla qubit states.
 From those amplitudes estimations we are able to plot the distance matrix between two data sets of binary vectors.
@@
 %
-Currently available quantum platforms are still subject to substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
+Currently available quantum platforms are still subject to a substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
 In order to reduce noise we average the measurement results over different states of the ancillary qubit, thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
@@
 The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$,
-where  $0$ indicates that $x_i$ are $y_j$ identical and $1$ means they are completely opposite in terms of their pairwise binary coordinates.
+where  $0$ indicates that $x_i$ and $y_j$ are identical and $1$ means they are the completely opposite in terms of their pairwise binary coordinates.
 
-The number of controlled gate operations to define the full distance matrix matches the number of controlled gate operations in \cite{trugenberger2001}, where the original algorithm for calculation of the Humming distance was introduced, but the number of remaining gates is reduced compared to \cite{trugenberger2001}, leading to less deep circuit, which is significant for NISQ devices.
+The number of controlled gate operations to define the full distance matrix matches the number of controlled gate operations in \cite{trugenberger2001}, where the original algorithm for calculation of the Humming distance was introduced, but the number of remaining gates is reduced compared to \cite{trugenberger2001}, leading to a less deep circuit, which is significant for NISQ devices.
 
@@
 and biology \cite{bio0, bio1, bio2}.
-Each set consists of three papers selected at random that focus on one of following topics:
+Each set consists of three papers selected at random that focus on one of the following topics:
 ``Quantum Machine Learning'',
@@
 that are possible to realize on  quantum computing devices that are currently available.
-The circuit realized in this way such
-that the calculation of Hamming distance implemented step-by-step between each sample vector and all cluster vectors is realized in one operation.
-In this case, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as measurement no longer needs to indicate for which input vector the distance has been measured.
+The calculation of Hamming distance implemented step-by-step between each sample vector and all cluster vectors is realized in one operation.
+In this case, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as the measurement no longer needs to indicate for which input vector the distance has been measured.
 The complexity of the quantum assisted SOFM then scales as $O(LN)$.
@@
 After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector
-and shift the closest cluster vectors to the sample one.
-The shift is made by the change of the first binary element in the cluster vectors different from the sample one.
+and shift the closest cluster vectors to the sample vectors.
+The shift is made by the change of the first binary element in the cluster vectors which does not match the binary element in the sample vector.
 The evolution of the labels presented on Fig.~\ref{convergence}(c).
@@
 
-This proof of concept example shows that developing of noisy intermediate scale quantum hardware will allow to solve some practical problems in unsupervised manner with very simple encoding of categorical data to quantum register.
+This proof of concept example shows that the development of noisy intermediate scale quantum hardware will allow to solve some practical problems in unsupervised manner with a very simple encoding of categorical data to the quantum register.
 Furthermore, it shows that it is possible to develop distance based hybrid quantum classical algorithms
-that speed up classical counterparts and can be implemented on near future quantum devices
-and that further developing distance based hybrid algorithms,
-that can learn in unsupervised manner, for categorical data such as genomic and so on has significant interest.
+that speed up classical counterparts and can be implemented on near--term quantum devices
+developing distance based hybrid algorithms. Such algorithms,
+that can learn in an unsupervised manner, for categorical data such as genomic data has significant interest.
 
@@
 that it can be used to solve clustering problems in an unsupervised manner.
-The procedure of solving such clustering problem requires calculating the distance many times in iterative way,
+The procedure of solving such a clustering problem requires calculating the distance many times in iterative way,
 which is calculated using a hybrid quantum-classical procedure.
@@
 Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm
-and in this way improve its performance.
+and in this way improves its performance.
 The complexity of the quantum assisted SOFM scales as $O(LN)$
@@
 Due to wide use of classical SOFMs in different areas of modern research and technology,
-this can give opportunities for the use of QASOFM in practical applications in near term, outperforming classical algorithms.
+this can give opportunities for the use of QASOFM in practical applications in the near term, outperforming classical algorithms.
 In addition, as our algorithm performs the Hamming distance calculation,
@@
 18-57-53007) and the State assignment (N.
-0089-2019-0002).
+AAAA-A19-119071190017-7).
 I.D.L.

### 2020-09-19 manuscript.tex
FIX typos from Tim

@@
 	Quantum computers offer opportunities to speed up such machine learning techniques.
-	Here, we propose a realization of quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural network.
+	Here, we introduce an algorithm for  quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural network.
 	We make a proof-of-concept realization of one of the central components on the IBM Q Experience
@@
 Such hybrid algorithms involve quantum circuits used for part of the execution and is typically trained in a classical learning loop.
-In particular, the Quantum Approximate Optimization Algorithm was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016}
+In particular, the Quantum Approximate Optimization Algorithm (QAOA) was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016}
 and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019,byrnes2018}.
@@
 \cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
-Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on noisy intermediate scale quantum (NISQ) devices \cite{preskill2018}.
+Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on Noisy Intermediate Scale Quantum (NISQ) devices \cite{preskill2018}.
 
@@
 This expression can be understood in the following way: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
-The larger the difference between the input component and the weight component, the larger the increment (decrement).
+The larger the difference between the input component and the weight component, the larger the increment or decrement.
 
@@
 
-The number of controlled gate operations to define the full distance matrix matches the number of controlled gate operations in \cite{trugenberger2001}, where the original algorithm for calculation of the Humming distance was introduced, but the number of remaining gates is reduced compared to \cite{trugenberger2001}, leading to a less deep circuit, which is significant for NISQ devices.
+The number of controlled gate operations to define the full distance matrix matches the number of controlled gate operations in Ref.
+\cite{trugenberger2001}, where the original algorithm for calculation of the Hamming distance was introduced, but the number of remaining gates is reduced compared to Ref.
+\cite{trugenberger2001}, leading to a less deep circuit, which is significant for NISQ devices.
 

### 2020-09-22 PNAS-template-main.tex
REFACTOR to PNAS style

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Template for preparing your research report submission to PNAS using Overleaf}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,c,1]{Author One}
\author[b,1,2]{Author Two} 
\author[a]{Author Three}

\affil[a]{Affiliation One}
\affil[b]{Affiliation Two}
\affil[c]{Affiliation Three}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership. The significance statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any competing interests here.}
\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Guide to using this template on Overleaf}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{https://www.pnas.org/page/authors/format}{PNAS Information for Authors}.

If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

\subsection*{Author Affiliations}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  title, author line and affiliations, keywords, abstract, significance statement, introduction, results, discussion, materials and methods, acknowledgments, and references. Other orders and headings are permitted.

\subsection*{Manuscript Length}

A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables). The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references cited in the main text should be included in the main manuscript file.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the \href{https://www.pnas.org/page/authors/journal-policies#xi}{Information for Authors}, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/page/authors/language-editing). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{frog}
\caption{Placeholder image of a frog with a long example legend to show justification setting.}
\label{fig:frog}
\end{figure}

\begin{SCfigure*}[\sidecaptionrelwidth][t]
\centering
\includegraphics[width=11.4cm,height=11.4cm]{frog}
\caption{This legend would be placed at the side of the figure, rather than below it.}\label{fig:side}
\end{SCfigure*}

\subsection*{Digital Figures}

EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Images must be provided at final size, preferably 1 column width (8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.

Figures and tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side legends.

\subsection*{Tables}
Tables should be included in the main manuscript file and should not be uploaded separately.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.

Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.

\begin{figure*}[bt!]
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{figure*}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information Appendix (SI)}

Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references. PNAS will publish SI uncomposed, as the authors have provided it. Additional details can be found here: \href{https://www.pnas.org/page/authors/format#Supporting_Information}{policy on SI}. The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}. Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.

Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

\subsubsection*{SI Datasets}

Supply .xlsx, .csv, .txt, .rtf, or .pdf files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files. Movie legends should be included in the SI Appendix file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10MB in size.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{pnas-sample}

\end{document}

### 2020-09-22 manuscript.tex
REFACTOR to PNAS style

@@
-\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
+\PassOptionsToPackage{inline}{enumitem}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\templatetype{pnasresearcharticle} % Choose template
+
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
+
 \DeclareMathOperator{\tr}{Tr}
@@
 
-\usepackage[inline]{enumitem} % inline enumeration
-
-\begin{document}
 \title{Hybrid quantum-classical unsupervised data clustering based on the Self-Organizing Feature Map}
 
-\author{I.~D.~Lazarev}
-\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
-\affiliation{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
+\author[a,b]{I.~D.~Lazarev}
+\author[c,d]{Marek Narozniak}
+\author[c,d,e,f,g]{Tim Byrnes}
+\author[a,1]{A.~N.~Pyrkov}
+\leadauthor{Pyrkov}
 
-\author{Marek Narozniak}
-\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
-\affiliation{Department of Physics, New York University, New York, NY, 10003, USA.}
+\affil[a]{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
+\affil[b]{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
 
-\author{Tim Byrnes}
-\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
-\affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
-\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
-\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
-\affiliation{Department of Physics, New York University, New York, NY 10003, USA}
+\affil[c]{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
+\affil[d]{Department of Physics, New York University, New York, NY, 10003, USA.}
 
-\author{A.~N.~Pyrkov}
-\email{Email address:pyrkov@icp.ac.ru}
-\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
+\affil[e]{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
+\affil[f]{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
+\affil[g]{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
 
-\date{\today}
+%\authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{The authors declare no conflict of interest.}
+
+%\equalauthors{\textsuperscript{1}A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable).}
+\correspondingauthor{\textsuperscript{1}E-mail: pyrkov\@icp.ac.ru}
+
+% Please add a significance statement to explain the relevance of your work
+\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
+The primary goal of the significance statement is to explain the relevance of the work in broad context to a broad readership.
+The significance statement appears in the paper itself and is required for all research papers.}
+
+\keywords{
+	quantum self-organizing feature map $|$
+	quantum machine learning $|$
+	neural network $|$
+	quantum unsupervised data clustering $|$
+	Hamming distance
+}
 
@@
 
-%\pacs{}
+\date{\today}
+%\doi{}
+
+\begin{document}
 \maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\section{Introduction}
-The combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
+\dropcap{T}he combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
@@
 
-\section{The quantum assisted self-organizing feature map (QASOFM)}
+\section*{The quantum assisted self-organizing feature map (QASOFM)}
 \label{sec:qasofm}
 
-\subsection{The classical algorithm}
+\subsection*{The classical algorithm}
 
@@
 
-\subsection{Optimized quantum scheme for Hamming distance calculation}
+\subsection*{Optimized quantum scheme for Hamming distance calculation}
 \label{subsec:qcircuit}
@@
 
-\section{Experimental demonstration of QASOFM}
+\section*{Experimental demonstration of QASOFM}
 
@@
 
-\section{Discussions}
+\section*{Discussions}
 We have developed a quantum assisted SOFM and showed a proof-of-concept experimental demonstration
@@
 
-\section*{Acknowledgments}
+\acknow{
 The work is supported by the RFBR-NSFC collaborative program (Grant No.
@@
 is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (61571301, D1210036A); the NSFC Research Fund for International Young Scientists (11650110425, 11850410426); NYU-ECNU Institute of Physics at NYU Shanghai; the Science and Technology Commission of Shanghai Municipality (17ZR1443600); the China Science and Technology Exchange Center (NGA-16-001); and the NSFC-RFBR Collaborative grant (81811530112).
+}
+\showacknow{}
 
@@
 
-\section*{Conflict of interest}
-
-The authors declare no conflict of interest.
-
-\section*{Keywords}
-quantum self-organizing feature map; quantum machine learning; neural network; quantum unsupervised data clustering; Hamming distance.
-
-\bibliographystyle{naturemag}
+%\bibliographystyle{naturemag}
 \bibliography{bibliography}

### 2023-07-15 manuscript.tex
Update Marek affiliation

--- 2020-09-19 manuscript.tex
@@
 \affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
-\affiliation{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
+\affiliation{Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
 
 \author{Marek Narozniak}
-\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
+\affiliation{New York University Shanghai, 567 West Yangsi Road, Shanghai, 200122, China.}
 \affiliation{Department of Physics, New York University, New York, NY, 10003, USA.}
+\affiliation{sqrtxx.com, Krakowiak\'ow 4, 20{-}255, Warsaw, Poland}
 
 \author{Tim Byrnes}
-\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
+\affiliation{New York University Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
+\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
+\affiliation{Shanghai Frontiers Science Center of Artificial Intelligence and Deep Learning, 567 West Yangsi Road, Shanghai, 200126, China.}
 \affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
-\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
-\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
+\affiliation{Center for Quantum and Topological Systems (CQTS), NYUAD Research Institute, New York University Abu Dhabi, UAE.}
 \affiliation{Department of Physics, New York University, New York, NY 10003, USA}
@@
 \section*{Acknowledgments}
-The work is supported by the RFBR-NSFC collaborative program (Grant No.
-18-57-53007) and the State assignment (N.
-AAAA-A19-119071190017-7).
-I.D.L.
-acknowledges support from the Russian Foundation for Basic Research, grant No.19-32-80004 and support from the Foundation for the Advancement of Theoretical Physics and Mathematics “BASIS” No.19-1-5-130-1.
-T.B.
-is supported by the Shanghai Research Challenge Fund; New York University Global Seed Grants for Collaborative Research; National Natural Science Foundation of China (61571301, D1210036A); the NSFC Research Fund for International Young Scientists (11650110425, 11850410426); NYU-ECNU Institute of Physics at NYU Shanghai; the Science and Technology Commission of Shanghai Municipality (17ZR1443600); the China Science and Technology Exchange Center (NGA-16-001); and the NSFC-RFBR Collaborative grant (81811530112).
+The work was performed as a part of a state task, State Registration No. AAAA-A19-119071190017-7.
+T.B. is supported by the National Natural Science Foundation of China (62071301); 
+NYU-ECNU Institute of Physics at NYU Shanghai; 
+Shanghai Frontiers Science Center of Artificial Intelligence and Deep Learning; 
+the Joint Physics Research Institute Challenge Grant; 
+the Science and Technology Commission of Shanghai Municipality (19XD1423000,22ZR1444600); 
+the NYU Shanghai Boost Fund; the China Foreign Experts Program (G2021013002L); 
+the NYU Shanghai Major-Grants Seed Fund; Tamkeen under the NYU Abu Dhabi Research Institute grant CG008; 
+and the SMEC Scientific Research Innovation Project (2023ZKZD55).
 

### 2023-07-15 manuscript.tex
Fix whitespace
Update Tim Byrnes affiliation

--- 2020-09-19 manuscript.tex
@@
 \affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
-\affiliation{ Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
+\affiliation{Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
 
@@
 \author{Tim Byrnes}
-\affiliation{New York University Shanghai, 1555 Century Ave, Pudong, Shanghai 200122, China}
+\affiliation{New York University Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
+\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
+\affiliation{Shanghai Frontiers Science Center of Artificial Intelligence and Deep Learning, 567 West Yangsi Road, Shanghai, 200126, China.}
 \affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
-\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 3663 Zhongshan Road North, Shanghai 200062, China}
-\affiliation{National Institute of Informatics, 2-1-2 Hitotsubashi, Chiyoda-ku, Tokyo 101-8430, Japan}
+\affiliation{Center for Quantum and Topological Systems (CQTS), NYUAD Research Institute, New York University Abu Dhabi, UAE.}
 \affiliation{Department of Physics, New York University, New York, NY 10003, USA}

### 2024-11-10 manuscript/main.tex
Move manuscript source to separated folder
Update introduction with some new refs

--- 2023-07-15 manuscript.tex
@@
 \author{I.~D.~Lazarev}
-\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
+\affiliation{Federal Research Center of Problems of Chemical Physics and Medicinal Chemistry RAS, Chernogolovka, Moscow Region, Russia, 142432}
 \affiliation{Faculty of Fundamental Physical-Chemical Engineering, Lomonosov Moscow State University, GSP-1, Moscow, Russia 119991}
@@
 \email{Email address:pyrkov@icp.ac.ru}
-\affiliation{Institute of Problems of Chemical Physics of Russian Academy of Sciences, Acad. Semenov av. 1, Chernogolovka, Moscow Region, Russia, 142432}
+\affiliation{Federal Research Center of Problems of Chemical Physics and Medicinal Chemistry RAS, Chernogolovka, Moscow Region, Russia, 142432}
 
@@
 \section{Introduction}
-The combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
+The combination of big data and artificial intelligence --  dubbed the fourth industrial revolution -- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-Paradoxically, the intrinsic mechanism of how neural networks work and why they are so powerful remains unknown (in many cases it is regarded as a black box).
+In 2022-2023, AI and neural networks experienced remarkable advancements, sparking substantial excitement in the field, particularly with the emergence of ChatGPT-4.
+Nonetheless, the quantity of parameters in such neural networks rapidly increase, leading to a significant impact on their interpretability and, consequently, their trustworthiness.
+Paradoxically, the intrinsic mechanism of how such neural networks work and why they are so powerful remains unknown (in many cases it is regarded as universal black box oracle).
+Additionally, these black-box oracles demand substantial resources for their training and exhibit a propensity to overfit when data is scarce \cite{sevilla2022compute}, in contrast to the human ability to effectively learn from "small" data.
 It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,
@@
 
-Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}.
+On the other hand, with recent technological advancements, quantum computation has moved beyond its academic roots.
+Although, its present scale may not yet rival classical computation, there is considerable excitement within this field and high expectations for rapid developments for commercial opportunities in the near future.
+In 2022, the Nobel Prize in Physics honored groundbreaking contributions to quantum information science, the Breakthrough Prize in Fundamental Physics recognized achievements in quantum communication and quantum computing, and IBM unveiled a quantum computer of 433 qubits.
+According to their roadmaps, companies such as IBM and Origin Quantum will empower their NISQ processors with more than 1000 qubits by 2024 and 2025, respectively.
+Recently, a lot of new algorithms were developed for such devices (see \cite{Cerezo2021,Bharti2022} for comprehensive review).
+In particular, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019, Benedetti2019, Cerezo2022, Sajjan2022}.
 The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
@@
 The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits,
-where the parameters behave much like the weights of a neural network \cite{lewenstein1994}.
+where the parameters behave much like the weights of a neural network \cite{lewenstein1994,Benedetti2019}.
 In particular, quantum algorithms for training and evaluating feedforward neural networks were developed,
@@
 For the problem of classification which is closely connected to the problem of clusterization, a protocol of quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
-Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
+A quantum capsule network together with an efficient quantum dynamic routing algorithm was proposed \cite{Liu2022}.
+A quantum version of vision transformer architecture was developed for quantum many-body systems \cite{Viteritti2023} that provides possibilities for quantum natural language processing.
+Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, Heese2022, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
 
@@
 \cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
-Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on Noisy Intermediate Scale Quantum (NISQ) devices \cite{preskill2018}.
+Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on Noisy Intermediate Scale Quantum (NISQ) devices \cite{preskill2018,Bharti2022,Cerezo2021}.
 
@@
 \begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{vectorized_sample.png}
+    \centering
+	\includegraphics[width=0.95\columnwidth]{vectorized_sample.pdf}
 	\caption{

### 2024-11-11 manuscript/main.tex
Remove trailing whitespace
Replace distance matrix figures with one that used 4 data qubits and extra averaging

@@
 
-%%%%%
 \begin{figure}[t]
 	\includegraphics[width=\columnwidth]{qcircuit.eps}
-    \caption{
+  \caption{%
 		Quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of binary vectors from two sets ${X}$ and ${Y}$.
@@
 \end{figure}
-%%%%%
 
-%%%%%
 \begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
-	\caption{
-		Hamming distance matrix between two data sets of binary vectors.
-		Values of distance that less than median distance value was marked black.
-		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
-		Result obtained on  IBMQ ``ibmq\_16\_melbourne'' backend is shown on the right figure and the accuracy in comparison with the classical simulation is more than 92 $\%$.
+  \includegraphics[width=\columnwidth]{distance_matrix.eps}
+  \caption{%
+    Result of tomography of all vector pairs that differ by up to one bit --- neighbors.
+    Distance values less than 1.7 are marked black.
+    Distance calculated by $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
+    where $d_{ij}$ is extracted from amplitudes by~(\ref{eq:amplitudes2distance}).
+    Amplitudes obtained on the IBM quantum backend named ``ibm\_sherbrook'' with only 8192 shots.
+    Question marks and crosses indicate missing and erroneous results compared to theoretical calculations.
+    The accuracy compared to classical simulation is better than 97 $\%$.
   }
@@
 \end{figure}
-%%%%%
 
@@
 %
-\begin{align}
-    d_{i,j}^H & \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))) .
-\end{align}
+\begin{equation}\label{eq:amplitudes2distance}
+    d_{i,j}^H \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))).
+\end{equation}
 %

### 2024-11-11 manuscript/main.tex
Update qcircuit figure with one for 4 qubits

@@
   \caption{%
-		Quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of binary vectors from two sets ${X}$ and ${Y}$.
+    Example of quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of 4-bit binary vectors from two sets ${X}$ and ${Y}$.
     First, we encode information about pairwise different qubits in a quantum state of the $X$-register by applying the CNOT gates.

### 2024-11-11 manuscript/main.tex
Expand tabs
Fix "No Society specified, using generic APS style"
See https://tex.stackexchange.com/a/19585
No type size specified, using default 10
Add draft of scalling discussion

@@
-\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
+\documentclass[pra,showkeys,twocolumn,showpacs,aps,10pt]{revtex4-1}
 \usepackage{amsmath}
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{tomography-errors-by-shots-n4.eps}
+  \caption{%
+  }
+  \label{fig:tomography-errors-by-shots-n4}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{tomography-errors-by-shots-n5.eps}
+  \caption{%
+  }
+  \label{fig:tomography-errors-by-shots-n5}
+\end{figure}
+
+\section{Scalability and robustness}
+
+Most quantum algorithms face three main scaling problems.
+The first is the exponential increase in the number of iterations of the quantum circuit required.
+This is not necessary in QASOFM\@.
+SOFM assumes that the cluster points do not have to move monotonically towards the cluster centroids,
+and that the vectors closest to the true BMU have the highest probability of being measured.
+The number of runs should only be chosen based on a given accuracy of BMU selection.
+It will only affect the speed of reaching the centres of the clusters.
+Fig.~\ref{fig:tomography-errors-by-shots-n4} and Fig.~\ref{fig:tomography-errors-by-shots-n5} show the dependence of the number of tomography errors of vectors
+which differ by no more than one bit (see Fig.~\ref{fig:distance_matrix}) on the number of shots.
+Here, tomography implies the determination of all pairs of nearest vectors in the whole state space.
+In Fig.~\ref{fig:tomography-errors-by-shots-n4} shows the state space of $2^4$ vectors, and Fig.~\ref{fig:tomography-errors-by-shots-n5} --- $2^5$ vectors.
+The points in Fig.~\ref{fig:tomography-errors-by-shots-n4} and Fig.~\ref{fig:tomography-errors-by-shots-n5} are the number of misidentified neighbors
+by the quantum circuit run on a noise simulator taking into account parameters of the real quantum platform.
+The lines are the result of a regression analysis of these data.
+The coefficient of determination of the model is $R^2 = 0.82$ for the state space of $2^4$ and $R^2 = 0.9$ for $2^5$.
+The triangles are the number of neighbors misidentified by the quantum scheme run on the IBM quantum backend named ``ibm\_sherbrook''.
+The number of errors decreases exponentially as the number of shots increases linearly.
+In contrast to QASOFM,
+tomography with increasing state space requires an exponential increase in the number of shots
+to distinguish neighbors of different orders.
+
+The second problem is a linear increase in a number of connections required for a single qubit.
+In QASOFM, this is the ancilla that accumulates distances in phase.
+Currently, this is a significant limitation for scaling on platforms with a rigid topology,
+such as IBM or Google Sycamore processors.
+For this reason, we cannot implement an algorithm for a larger problem.
+The only solution is to create a large auxiliary network,
+but this requires numerous two-qubit operations,
+which are currently performed with insufficient accuracy.
+Other approaches to implementing the quantum platform,
+for example for ions in traps or Bose-Einstein condensates,
+are likely to overcome this limitation.
+
+The third problem is the difficulty of converting digital data into a quantum state.
+The dimension of the data is usually much larger than the number of cluster vectors.
+Consequently, the size of the subspace stretched over the cluster vectors will be negligible compared to the whole space.
+This makes it possible to efficiently generate quantum circuits to initialise such superposition quantum states.
+
 \section{Discussions}

### 2024-12-11 manuscript/main.tex
Improve text

--- 2024-11-11 manuscript/main.tex
@@
 \section{Introduction}
-The combination of big data and artificial intelligence --  dubbed the fourth industrial revolution -- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
+The combination of big data and artificial intelligence (AI) --  dubbed the fourth industrial revolution -- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-In 2022-2023, AI and neural networks experienced remarkable advancements, sparking substantial excitement in the field, particularly with the emergence of ChatGPT-4.
-Nonetheless, the quantity of parameters in such neural networks rapidly increase, leading to a significant impact on their interpretability and, consequently, their trustworthiness.
+From 2022, AI and neural networks experienced remarkable advancements, sparking substantial excitement in the field, particularly with the emergence of Large Language Models (LLMs) such as ChatGPT-4\cite{GPT4}.
+In 2024, Nobel prizes in physics\cite{NobelPrizePhysics} and chemistry\cite{NobelPrizeChemistry} were awarded for foundational discoveries for developement of machine learning and applications of AI in chemistry.
+Despite such groundbreaking achievements, interpretability and trustworthiness of such NNs is very low due to the quantity of parameters in such neural networks rapidly increase.
 Paradoxically, the intrinsic mechanism of how such neural networks work and why they are so powerful remains unknown (in many cases it is regarded as universal black box oracle).
 Additionally, these black-box oracles demand substantial resources for their training and exhibit a propensity to overfit when data is scarce \cite{sevilla2022compute}, in contrast to the human ability to effectively learn from "small" data.
-It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,
-and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
 
-In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN)
-that is trained in an unsupervised manner.
-SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
-and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
-The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
-Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
+%It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
 
-On the other hand, with recent technological advancements, quantum computation has moved beyond its academic roots.
+On the other hand, quantum computation is another fast growing area of research that is moving beyond its academic roots.
 Although, its present scale may not yet rival classical computation, there is considerable excitement within this field and high expectations for rapid developments for commercial opportunities in the near future.
-In 2022, the Nobel Prize in Physics honored groundbreaking contributions to quantum information science, the Breakthrough Prize in Fundamental Physics recognized achievements in quantum communication and quantum computing, and IBM unveiled a quantum computer of 433 qubits.
-According to their roadmaps, companies such as IBM and Origin Quantum will empower their NISQ processors with more than 1000 qubits by 2024 and 2025, respectively.
-Recently, a lot of new algorithms were developed for such devices (see \cite{Cerezo2021,Bharti2022} for comprehensive review).
+In 2022, the Nobel Prize in Physics honored groundbreaking contributions to quantum information science\cite{NobelPrizePhysicsa}, the Breakthrough Prize in Fundamental Physics in 2023 recognized achievements in quantum communication and quantum computing\cite{BreakthroughPrizeWinners}.
+From practical point of view, IBM  empowered their NISQ processors with more than 1000 qubits\cite{IBMRoadmapQuantumcentric}, Google demonstrated quantum error correction below the surface code threshold\cite{acharyaQuantumErrorCorrection2024}, and QuEra ambitiously announced 100 logical qubit processor in 2026 in their roadmap\cite{QuantumErrorCorrection}.
+A lot of new algorithms were developed for such devices (see \cite{Cerezo2021,Bharti2022} for comprehensive review).
 In particular, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019, Benedetti2019, Cerezo2022, Sajjan2022}.
+In comparison with classical ML, it was recently shown that Quantum Machine Learning (QML) models can be traine with fewer parameters in comparison with classical ANNs\cite{abbasPowerQuantumNeural2021,caroGeneralizationQuantumMachine2022}.
 The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
@@
 where the parameters behave much like the weights of a neural network \cite{lewenstein1994,Benedetti2019}.
-In particular, quantum algorithms for training and evaluating feedforward neural networks were developed,
+In particular, quantum algorithms for training and evaluating feed forward neural networks were developed,
 which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
 Quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}.
-Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}.
+Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019,daiQuantumBayesianOptimization2023}.
 For the problem of classification which is closely connected to the problem of clusterization, a protocol of quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
 A quantum capsule network together with an efficient quantum dynamic routing algorithm was proposed \cite{Liu2022}.
-A quantum version of vision transformer architecture was developed for quantum many-body systems \cite{Viteritti2023} that provides possibilities for quantum natural language processing.
-Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, Heese2022, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
+The quantum versions of transformer architecture and transformer of quantum states were developed\cite{Viteritti2023,cherratQuantumVisionTransformers2024,khatriQuixerQuantumTransformer2024,zhangTransformerQuantumState2023} that provides possibilities for quantum natural language processing.
+Inspired by the classical counterpart, the quantum denoising diffusion probabilistic model (QuDDPM) was proposed to enable efficiently trainable generative learning of quantum data\cite{zhangGenerativeQuantumMachine2024}.
+Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{kolleQuantumDenoisingDiffusion2024,li2019, Heese2022, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
 
@@
 In this paper, we develop a hybrid quantum-assisted SOFM (QASOFM)
-and apply it to the data clustering problem in an unsupervised manner.
-The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
+and apply it to the data clustering problem in an unsupervised manner. The self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN)
+that is trained in an unsupervised manner.
+SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
+and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
+The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
+Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
+The idea of QASOFM is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
 In order to make our protocol more appropriate for the currently available generation of NISQ quantum devices, we optimized a circuit for realizing the Hamming distance by reducing number of one-qubit operations.
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{distance_matrix.eps}
-  \caption{%
-    Result of tomography of all vector pairs that differ by up to one bit --- neighbors.
-    Distance values less than 1.7 are marked black.
-    Distance calculated by $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
-    where $d_{ij}$ is extracted from amplitudes by~(\ref{eq:amplitudes2distance}).
-    Amplitudes obtained on the IBM quantum backend named ``ibm\_sherbrook'' with only 8192 shots.
-    Question marks and crosses indicate missing and erroneous results compared to theoretical calculations.
-    The accuracy compared to classical simulation is better than 97 $\%$.
-  }
-  \label{fig:distance_matrix}
-\end{figure}
+In the work \cite{trugenberger2001}, the concept of probabilistic quantum memory was introduced.
+In this framework, quantum entanglement is utilized to significantly enhance the storage capacity of associative memories and to retrieve noisy or incomplete information.
+Specifically, the number of binary patterns that can be stored in such a quantum memory is exponential in relation to the number $ n $ of qubits, with a maximum capacity of $ p_{\text{max}} = 2^n $.
+This means it can optimally store all possible binary patterns that can be formed with $ n $ bits.
+
+The retrieval process is probabilistic and involves postselection of the measurement results.
+This implies that the retrieval algorithm must be repeated until a certain threshold $T$ is met, or until the measurement of a control qubit yields a specific outcome.
+In the first scenario, the input is not recognized, while in the second, the output is determined by a probability distribution over the memory that is concentrated around the stored patterns that are closest to the input, as measured by Hamming distance.
+Here, we improve the approach for Hamming distance calculation proposed for realization of probabilistic quantum memory\cite{trugenberger2001} and apply it for realization of QASOFM.
 
@@
 %
+
+%%%%%
+\begin{figure*}[t]
+    \centering
+  \includegraphics[width=1.85\columnwidth]{vectorized_sample.pdf}
+  \caption{
+    (a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
+    Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
+    The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
+    (b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
+  }
+  \label{fig:vectorized_sample}
+\end{figure*}
+%%%%%
+
 This completes the step of projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit.
@@
 %
+
 Currently available quantum platforms are still subject to a substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
@@
 
-%%%%%
-\begin{figure}[t]
-    \centering
-	\includegraphics[width=0.95\columnwidth]{vectorized_sample.pdf}
-	\caption{
-		(a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
-		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
-		The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
-		(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
-	}
-	\label{fig:vectorized_sample}
-\end{figure}
-%%%%%
-
 \section{Experimental demonstration of QASOFM}
 
-We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.
+We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section with the use of IBM Q Experience.
+
 We perform unsupervised data clustering for three sets of paper abstracts from three different fields:
@@
 
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{distance_matrix.eps}
+  \caption{%
+    Comparison of classical and quantum calculations of Humming distance matrix.
+Result of tomography of all vector pairs that differ not more than one bit --- nearest neighbors.
+The distances are calculated as $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
+    where $d_{ij}$ was extracted from quantum state amplitudes with the use formula of ~(\ref{eq:amplitudes2distance}).
+    The calculations of amplitudes were implemented on the IBM quantum backend named ``ibm\_sherbrook'' with statistics of 8192 shots.
+Distance values less than 1.7 are marked black and coincide to the nearest neighbors.
+    Question marks and crosses indicate missing and incorrect results compared to theoretical calculations.
+    The accuracy compared to classical simulation is better than 97 $\%$.
+  }
+  \label{fig:distance_matrix}
+\end{figure}
+
 In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.~\ref{fig:distance_matrix}.
-We see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
-The match is close, but not perfect, and we attribute it to noise in the currently available non fault-tolerant quantum processors.
+We see very good agreement between the distance matrices calculated classically and on the IBM Q Experience.
+
+%%%%%
+\begin{figure*}[t]
+  \includegraphics[width=1.85\columnwidth]{convergence.png}
+  \caption{
+    (a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2),
+    for the selected 9 words from the bag of words model.
+    (b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend.
+    Vectors mean cluster elements for BIO, MED, QML groups,
+    for the C0, C1, C2 cluster vectors, respectively.
+    (c) The evolution of label distribution on each learning epoch.
+  }
+  \label{convergence}
+\end{figure*}
+%%%%%
 
@@
 Initially, the cluster vectors were randomly chosen and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
-In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT\cite{qiskit} library.
+For the cluster  vectors (binary representation of rows of Fig.\ref{convergence}a) we use the simplest basis encoding into quantum  states with the standard initialization of QISKIT\cite{rathQuantumDataEncoding2024,qiskit} library.
 Each epoch of the algorithm consists of distance calculation between all data and cluster vectors
@@
 
-%%%%%
-\begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{convergence.png}
-	\caption{
-		(a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2),
-		for the selected 9 words from the bag of words model.
-		(b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend.
-		Vectors mean cluster elements for BIO, MED, QML groups,
-		for the C0, C1, C2 cluster vectors, respectively.
-		(c) The evolution of label distribution on each learning epoch.
+\begin{figure*}[t]
+  \includegraphics[width=1.85\columnwidth]{tomography-errors-by-shots-n4-n5.eps}
+  \caption{Error in the distance matrix of QASOFM versus number of runs of the quantum backend for (a) $4$ and (b) $5$ qubits.
+We can see exponential decrease in error with linear increase in number of runs.
   }
-	\label{convergence}
-\end{figure}
-%%%%%
+  \label{fig:tomography-errors-by-shots-n4}
+\end{figure*}
+
+\section{Scalability Of QASOFM}
+
+For scalability of NISQ algorithms both software and hardware restrictions are important.
+
+From software point of view the main challenge is effective encoding of classical data into quantum states.
+In the proposed QASOFM we used the simplest basis encoding model.
+This model is non-efficient in the meaning of required quantum resources.
+Furthermore, the bag-of-word model is not optimal itself (we used it because of very small classical data needed to be embedded).
+The use of better word embedding models will allow to improve QASOFM performance.
+Also, the dimension of the input data is usually much larger than the number of cluster vectors.
+Consequently, the size of the subspace spanned on the cluster vectors will be negligible compared to the whole space.
+This in theory provides opportunities to efficiently generate quantum circuits for initialization of such superposition of sparse quantum states with smaller number of operations.
+
+The second challenge is in currently available hardware.
+In QASOFM, the ancillary qubit  accumulates the distances in phase and because of that requires to be as much connected as possible to other qubits.
+The topology of superconducting quantum processors,such as IBM or Google Sycamore, presents the significant limitations for that and consequently for the scaling of QASOFM algorithm to a larger problem.
+The possible solution is to create a large auxiliary network,
+but it requires a lot of two-qubit operations,
+which are currently performed with insufficient accuracy.
+Another solution is to use platforms with star architecture (all to all connected) which are under development with trapped ion technology\cite{QuantinuumsHSeriesHits}.
+It will higly improve performance and scalability.
+
+Furthermore, our QASOFM allows to avoid the use exponential resource requirement at read-out operation (as for quantum tomography). 
+SOFM assumes that the cluster points do not have to move monotonically towards the cluster centroids,
+and that the vectors closest to the true BMU have the highest probability of being measured.
+Thus, the number of runs should only be chosen to measure distances to nearest neighbors with high accuracy, not all possible distance calculations.
+The protocol implies finding of all pairs of nearest vectors in the whole state space.
+Fig.~\ref{fig:tomography-errors-by-shots-n4} shows the error versus number of runs for determination of distance matrix for $4$  and $5$ qubits respectively.
+The points in Fig.~\ref{fig:tomography-errors-by-shots-n4} define the number of errors in neighbors
+by the quantum circuit runs on a noise simulator taking into account parameters of the real quantum platform.
+The lines are the result of a regression analysis of these data.
+The coefficient of determination of the model is $R^2 = 0.86$ for the state space of $4$ qubits and $R^2 = 0.92$ for $5$ qubits.
+The triangles are the number of errors by the quantum scheme run on the IBM quantum backend named ``ibm\_sherbrook''.
+The number of errors decreases exponentially as the number of shots increases linearly. This provide advantage for our protocol in read out.
 
@@
 In addition, as our algorithm performs the Hamming distance calculation,
-it has the potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
+it has the potential to enhance any classical algorithm that relies on calculating distances between binary data.
+In particular, nowadays a lot of algorithms are under development for quantized datasets with binary embeddings\cite{zhuangFastTrainingTripletBased2016,yiBinaryEmbeddingFundamental2015,AsymmetricDistancesBinary}.
+Furthermore, recently quantum protocols for euclidean distance calculations were introduced\cite{yuQuantumAlgorithmsSimilarity2020,zardiniQuantumKnearestNeighbors2024} that allows to generalize QASOFM to non-binary data.
 In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure
-and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required,
+and our circuit could potentially enhance other distance based algorithms for binary data  as long as exact distance is not required,
 but when knowledge of nearest vectors is sufficient.
@@
 \section*{Acknowledgments}
-The work was performed as a part of a state task, State Registration No. AAAA-A19-119071190017-7.
+The work was partly supported by the
+Russian Science Foundation (grant no 23-21-00507) and partly performed as a part of a state task, State Registration No. 124013000760-0.
 T.B. is supported by the National Natural Science Foundation of China (62071301);
@@
 
+\section*{Data Availability Statement}
+
+The source code is available on https://github.com/kephircheek/qasofm
+
 \section*{Conflict of interest}
@@
 \bibliographystyle{naturemag}
-\bibliography{bibliography}
+\bibliography{bibliography,bibl}
 

### 2024-12-12 manuscript/main.tex
Improve text

--- 2024-11-11 manuscript/main.tex
@@
-\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
+\documentclass[pra,showkeys,twocolumn,showpacs,aps,10pt]{revtex4-2}
 \usepackage{amsmath}
@@
 \author{Tim Byrnes}
-\affiliation{New York University Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
-\affiliation{NYU-ECNU Institute of Physics at NYU Shanghai, 567 West Yangsi Road, Shanghai, 200126, China.}
-\affiliation{Shanghai Frontiers Science Center of Artificial Intelligence and Deep Learning, 567 West Yangsi Road, Shanghai, 200126, China.}
+\affiliation{New York University Shanghai; NYU-ECNU Institute of Physics at NYU Shanghai; Shanghai Frontiers Science Center of Artificial Intelligence and Deep Learning, 567 West Yangsi Road, Shanghai, 200126, China.}
 \affiliation{State Key Laboratory of Precision Spectroscopy, School of Physical and Material Sciences, East China Normal University, Shanghai 200062, China}
@@
   Unsupervised machine learning is one of the main techniques employed in artificial intelligence.
-	Quantum computers offer opportunities to speed up such machine learning techniques.
-	Here, we introduce an algorithm for  quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural network.
-	We make a proof-of-concept realization of one of the central components on the IBM Q Experience
-	and show that it allows us to reduce the number of calculations in a number of clusters.
-	We compare the results with the classical algorithm on a toy example of unsupervised text clustering.
+We introduce an algorithm for quantum assisted unsupervised data clustering using the self-organizing feature map, a type of artificial neural network.
+The complexity of  our algorithm scales as $O(LN)$, in comparison to
+the classical case which scales as $O(LMN)$,
+where $N$ is number of samples, $M$ is number of randomly sampled cluster vectors, and $L$ is number of the shifts of cluster vectors.
+  We perform a proof-of-concept experimental demonstration of one of the central components on the IBM quantum computer and show that it allows us to reduce the number of calculations in the number of clusters.
+Our algorithm exhibits exponential decrease in the errors of the distance matrix with the number of runs of the algorithm.
 \end{abstract}
@@
 \section{Introduction}
-The combination of big data and artificial intelligence --  dubbed the fourth industrial revolution -- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
-Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-In 2022-2023, AI and neural networks experienced remarkable advancements, sparking substantial excitement in the field, particularly with the emergence of ChatGPT-4.
-Nonetheless, the quantity of parameters in such neural networks rapidly increase, leading to a significant impact on their interpretability and, consequently, their trustworthiness.
+The combination of big data and artificial intelligence (AI) ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
+Contemporary artificial intelligence methods based on neural networks (NNs) also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018,NobelPrizeChemistry,NobelPrizePhysics}.
+Despite such groundbreaking achievements, interpretability of such NNs is very low due to the quantity of parameters in such neural networks rapidly increase.
 Paradoxically, the intrinsic mechanism of how such neural networks work and why they are so powerful remains unknown (in many cases it is regarded as universal black box oracle).
-Additionally, these black-box oracles demand substantial resources for their training and exhibit a propensity to overfit when data is scarce \cite{sevilla2022compute}, in contrast to the human ability to effectively learn from "small" data.
-It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,
-and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
+Additionally, these black-box oracles demand substantial resources for their training and exhibit a propensity to overfit when data is scarce \cite{sevilla2022compute}, in contrast to the human ability to effectively learn from ``small'' data.
 
-In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN)
-that is trained in an unsupervised manner.
-SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
-and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
-The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
-Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
+%It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
 
-On the other hand, with recent technological advancements, quantum computation has moved beyond its academic roots.
-Although, its present scale may not yet rival classical computation, there is considerable excitement within this field and high expectations for rapid developments for commercial opportunities in the near future.
-In 2022, the Nobel Prize in Physics honored groundbreaking contributions to quantum information science, the Breakthrough Prize in Fundamental Physics recognized achievements in quantum communication and quantum computing, and IBM unveiled a quantum computer of 433 qubits.
-According to their roadmaps, companies such as IBM and Origin Quantum will empower their NISQ processors with more than 1000 qubits by 2024 and 2025, respectively.
-Recently, a lot of new algorithms were developed for such devices (see \cite{Cerezo2021,Bharti2022} for comprehensive review).
-In particular, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019, Benedetti2019, Cerezo2022, Sajjan2022}.
-The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
+Meanwhile, quantum computers are a rapidly developing technology that promises improvement in algorithmic scaling for particular types of problems \cite{nielsen2010quantum,Cerezo2021,Bharti2022,byrnes2021quantum}.
+In particular, there has been much interest recently in applying quantum computing techniques to machine learning (ML) problems \cite{dunjko2018, biamonte2017, schuld2014, carleo2019, Benedetti2019, Cerezo2022, Sajjan2022}.
+In comparison with classical ML, it was recently shown that Quantum Machine Learning (QML) models can be trained with fewer parameters in comparison with classical NNs \cite{abbasPowerQuantumNeural2021,caroGeneralizationQuantumMachine2022}.
+The main focus of early works in QML was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
 Quantum algorithms were developed for  linear regression, principal component analysis, support vector machine, K-means algorithm and others \cite{lloyd2013,lloyd2014,dunjko2016,paparo2014,rebentrost2014}.
@@
 where the parameters behave much like the weights of a neural network \cite{lewenstein1994,Benedetti2019}.
-In particular, quantum algorithms for training and evaluating feedforward neural networks were developed,
+In particular, quantum algorithms for training and evaluating feed forward neural networks were developed,
 which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
 Quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}.
-Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}.
-For the problem of classification which is closely connected to the problem of clusterization, a protocol of quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
-A quantum capsule network together with an efficient quantum dynamic routing algorithm was proposed \cite{Liu2022}.
-A quantum version of vision transformer architecture was developed for quantum many-body systems \cite{Viteritti2023} that provides possibilities for quantum natural language processing.
-Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, Heese2022, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
+Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019,daiQuantumBayesianOptimization2023}.
+For the problem of classification which is closely connected to the problem of clusterization, a protocol for quantum classification via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
+In Ref.  \cite{Liu2022}, a quantum capsule network together with an efficient quantum dynamic routing algorithm was proposed.
+The quantum versions of transformer architecture and transformer of quantum states were developed \cite{Viteritti2023,cherratQuantumVisionTransformers2024,khatriQuixerQuantumTransformer2024,zhangTransformerQuantumState2023}, that provides possibilities for quantum natural language processing.
+Inspired by the classical counterpart, the quantum denoising diffusion probabilistic model (QuDDPM) was proposed to enable efficiently trainable generative learning of quantum data \cite{zhangGenerativeQuantumMachine2024}.
+Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{kolleQuantumDenoisingDiffusion2024,li2019, Heese2022,killoran2019,bondarenko2019,dunjko2017,nautrup2019,foesel2018, rebentrost2018,purushothaman1997,verdon2019,cherny2019,byrnes2013,mishra2019, pyrkov2019, vinci2019, lu2019,mohseni2023deep,mohseni2024deep}.
 
-At the same time, non-neural-network based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}.
-Such hybrid algorithms involve quantum circuits used for part of the execution and is typically trained in a classical learning loop.
+At the same time, non-NN based hybrid quantum classical algorithms have become a new direction of significant interest \cite{mcclean2016,arute2020,akshay2020}.
+Such hybrid algorithms involve quantum circuits used for part of the execution and are typically trained in a classical learning loop.
 In particular, the Quantum Approximate Optimization Algorithm (QAOA) was developed to find approximate solutions to combinatorial optimization problems \cite{farhi2014,farhi2016}
-and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019,byrnes2018}.
+and designed for problems such as MAX-CUT and Grover's algorithm \cite{arute2020,akshay2020,wang2018,jiang2017,huang2019,wecker2016,pagano2019,byrnes2018,liao2021quadratic}.
 Another example of a well-known hybrid quantum classical algorithm is the Variational Quantum Eigensolver (VQE) for applications in quantum simulations
@@
 
-In this paper, we develop a hybrid quantum-assisted SOFM (QASOFM)
+In this paper, we develop a hybrid quantum-assisted self-organizing feature map (QASOFM)
 and apply it to the data clustering problem in an unsupervised manner.
-The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
-In order to make our protocol more appropriate for the currently available generation of NISQ quantum devices, we optimized a circuit for realizing the Hamming distance by reducing number of one-qubit operations.
-We then apply it to a toy example of clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM Q experience quantum computer \cite{ibmq}
+The self-organizing feature map (SOFM) was first proposed by Kohonen \cite{kohonen1990,kohonen1996,kohonen1997} as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}.
+SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
+and in comparison with many other NNs, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
+The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
+Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
+The idea of the QASOFM is based on the use of the Hamming distance as a distance metric for training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
+In order to make our protocol more appropriate for the currently available generation of NISQ quantum devices, we optimized the quantum circuit for realizing the Hamming distance by reducing the number of one-qubit operations.
+We then apply it to a toy example of clustering paper abstracts and give a proof-of-concept realization of the quantum assisted SOFM on the IBM quantum computer \cite{ibmq}
 and compare it to the classical case.
@@
     Schematic illustration of the clustering problem considered in this paper.
-		Blue triangles represent clusters and red circles are data points.
+    Triangles represent clusters and circles are data points.
     The training process moves clusters to fit the data points.
@@
 
-\section{The quantum assisted self-organizing feature map (QASOFM)}
+\section{The quantum assisted self-organizing feature map}
 \label{sec:qasofm}
@@
 
-The SOFM is one of the most widely-used unsupervised learning methods used in various areas of modern science.
-It was first proposed by Kohonen as a self-organizing unsupervised learning algorithm which produces feature maps similar to those occurring in the brain \cite{solan2001}.
 The SOFM algorithm operates with a set of input objects, each represented by a $N$-dimensional vector,
 and describes a mapping from a higher-dimensional input space to a lower-dimensional map space that preserves the topological properties of the input space, commonly referred to as a bi-dimensional map.
-
 The input dimensions are associated with the features,
@@
 selecting a winning cluster vector, also called the best matching unit (BMU), and updating its weights (Fig.~\ref{fig:sofm_fitting}).
-More specifically, they consist of the four step process:
-\begin{enumerate*}
-\item select an input vector randomly from the set of all input vectors;
-\item find a cluster vector which is closest to the input vector;
-\item adjust the weights of the best matching unit and neurons close to it on the feature map in such a way
-	that these vectors  becomes closer to the input vector;
-\item repeat this process for many iterations until it converges.
-\end{enumerate*}
+The basic idea of the SOFM can be summarized by the four step process:
+%
+\begin{enumerate}
+\item Select an input vector randomly from the set of all input vectors.
+\item Find a cluster vector which is closest to the input vector.
+\item Adjust the weights of the best matching unit and neurons close to it on the feature map in such a way   that these vectors  becomes closer to the input vector.
+\item Repeat this process for many iterations until it converges.
+\end{enumerate}
 
-On a step $t$ when the best matching unit $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected,
-the weights $\vec{w}_{i}$ of the best matching unit and its neighbors on the feature map are adjusted according to
+Specifically, when the BMU $\vec{w}_{c}$ for a input $\vec{x}(t)$ is selected,
+the weights $\vec{w}_{i}$ of the BMU and its neighbors on the feature map are adjusted according to
 %
@@
 %
-Here,  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually taken as a Gaussian or delta function) which defines the vicinity of the BMU labeled by an index $c$, the weights of the neighbors in the vicinity should also be adjusted in the same manner as for the BMU.
+Here, $ t $ is the iteration step,  $\alpha(t)$ is the monotonically decreasing learning rate and $\theta(c, i, t)$ is the neighborhood function (usually taken as a Gaussian or delta function), which defines the vicinity of the BMU labeled by an index $c$.
+The weights of the neighbors in the vicinity should also be adjusted in the same manner as for the BMU.
 This expression can be understood in the following way: if a component of the input vector $\vec{x}(t)$ is greater than the corresponding weight $ \vec{w}_{i}(t) $, increase the weight of the BMU and the weights indexed by $i$ and defined with the neighborhood function by a small amount defined by the learning rate $\alpha(t)$; if the input component is smaller than the weight, decrease the weight by a small amount.
@@
 
-Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (blue triangles on Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (red circles on Fig.~\ref{fig:sofm_fitting}).
+Intuitively, this procedure can be geometrically interpreted as iteratively moving the cluster vectors defined by the corresponding weight $ \vec{w}_{i}(t) $ (triangles in Fig.~\ref{fig:sofm_fitting}) in space one at a time in a way that ensures each move is following the current trends inferred from their distances to the input objects defined by $\vec{x}(t)$ (circles in Fig.~\ref{fig:sofm_fitting}).
 
 In the original version of SOFM, the winning cluster vector is selected based on the Euclidean distance between an input vector and the cluster vectors.
-In this paper, we deal with a binary vector clustering problem, where the Hamming distance is more suitable \cite{appiah2009, santana2017}.
+In this paper, we deal with a binary vector clustering problem, where we use the Hamming distance \cite{appiah2009, santana2017}.
 Using a simple technique of encoding classical binary information into a quantum register \cite{trugenberger2001},
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{qcircuit.eps}
+\begin{figure*}[t]
+  \includegraphics[width=1.85\columnwidth]{qcircuit.eps}
   \caption{%
-    Example of quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of 4-bit binary vectors from two sets ${X}$ and ${Y}$.
-    First, we encode information about pairwise different qubits in a quantum state of the $X$-register by applying the CNOT gates.
+    The quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of 4-bit binary vectors from two sets ${X}$ and ${Y}$.
+    First, we encode the information about pairwise different qubits in a quantum state of the $X$-register by applying the CNOT gates.
     Second, the Hamming distance values are extracted to the amplitudes of superposition with the control phase rotation (see Eq.~(\ref{eq:controled_phase_rotation})) and Hadamard gates.
-    Finally, a quantum space of the $X$-register is returned to the initial basis for information retrieval.
-  }
+    Finally, the $X$-register is returned to the initial basis for information retrieval.   }
   \label{fig:qcircuit}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{distance_matrix.eps}
-  \caption{%
-    Result of tomography of all vector pairs that differ by up to one bit --- neighbors.
-    Distance values less than 1.7 are marked black.
-    Distance calculated by $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
-    where $d_{ij}$ is extracted from amplitudes by~(\ref{eq:amplitudes2distance}).
-    Amplitudes obtained on the IBM quantum backend named ``ibm\_sherbrook'' with only 8192 shots.
-    Question marks and crosses indicate missing and erroneous results compared to theoretical calculations.
-    The accuracy compared to classical simulation is better than 97 $\%$.
-  }
-  \label{fig:distance_matrix}
-\end{figure}
+In Ref.
+\cite{trugenberger2001}, the concept of a probabilistic quantum memory was introduced.
+In this framework, quantum entanglement is utilized to significantly enhance the storage capacity of associative memories and to retrieve noisy or incomplete information.
+Specifically, the number of binary patterns that can be stored in such a quantum memory is exponential in relation to the number $ n $ of qubits, with a maximum capacity of $ p_{\text{max}} = 2^n $.
+This means it can optimally store all possible binary patterns that can be formed with $ n $ bits.
 
-The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single auxiliary qubit $\left| a \right\rangle$.
-During the whole process, the $\left| Y \right\rangle$ register is used to store the cluster states.
+The retrieval process is probabilistic and involves postselection of the measurement results.
+This implies that the retrieval algorithm must be repeated until a certain threshold is met, or until the measurement of a control qubit yields a specific outcome.
+In the first scenario, the input is not recognized, while in the second, the output is determined by a probability distribution over the memory that is concentrated around the stored patterns that are closest to the input, as measured by Hamming distance.
+Here, we improve the approach for Hamming distance calculation proposed for realization of probabilistic quantum memory in Ref.
+\cite{trugenberger2001} and apply it to the realization of the QASOFM.
+
+The overall procedure involves two registers of $n$ qubits each, denoted $\left| X \right\rangle$ and $\left| Y \right\rangle$, along with a single ancilla qubit $\left| a \right\rangle$.
+Throughout the whole process, the $\left| Y \right\rangle$ register is used to store the cluster states.
 At the beginning and end of the procedure, the $\left| X \right\rangle$ register stores the input vectors.
@@
 %
-The two registers along with the auxiliary qubit comprise the initial state of the quantum computer according to
+The two registers along with the ancilla qubit comprise the initial state of the quantum computer according to
 %
@@
 Given this initial state we may begin the processing of the problem.
-The quantum circuit that we follow is shown in Fig.\ref{fig:qcircuit}.
+The quantum circuit that we follow is shown in Fig.
+\ref{fig:qcircuit}.
 We start by elementwise applying CNOT gates between all the qubits of the $X$ and $Y$ registers (the ``Encoding'' part of Fig.
-\ref{fig:qcircuit})
+\ref{fig:qcircuit}), such that the state becomes
 %
@@
 %
-where $d^{(\alpha)}_{ij} = \mathrm{CNOT}(y^{(\alpha)}_i, x^{(\alpha)}_j)$, and $\alpha = 1 \dots n$  is the qubit index in the register.
+where $d^{(\alpha)}_{ij} = \mathrm{CNOT}(y^{(\alpha)}_i, x^{(\alpha)}_j)$, and $\alpha \in  [1,n]$  is the qubit index in the register.
 At this stage of the computation the $\left| X \right\rangle$ no longer stores the input vectors,
-instead it stores the information about pairwise different qubits between the input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$.
+instead it stores the information about pairwise difference of the qubits between the input vectors $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$.
 
-The next step, depicted on the Fig.\ref{fig:qcircuit} as the ``Extraction'' stage, is to extract the accumulated information about the differences between each pair $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ by projection onto the amplitude of the superposed state.
-This is achieved by applying the Hadamard gate on the auxiliary qubit, followed by a controlled phase gate on $\left|a\right\rangle \left|d_{ij}^{(\alpha)}\right\rangle$, where the phase gate is defined by
+The next step, depicted in Fig.
+\ref{fig:qcircuit} as the ``Extraction'' stage, is to extract the accumulated information about the differences between each pair $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ by projection onto the amplitude of the superposed state.
+This is achieved by applying the Hadamard gate on the ancilla qubit, followed by a controlled phase gate on $ | a \rangle  |d_{ij}^{(\alpha)} \rangle$, where the phase gate is defined by
 %~(\ref{eq:control_phase_rotation})
@@
 %
-where $d_{ij}^{(\alpha)}$ is the control qubit and the ancilla qubit $\left| a \right\rangle$ is the target.
+where $ | d_{ij}^{(\alpha)} \rangle $ is the control qubit and the ancilla qubit $\left| a \right\rangle$ is the target.
 Finally, another Hadamard gate is applied on the ancilla qubit (see Fig.~\ref{fig:qcircuit}).
 
-After the first Hadamard on the ancilla qubit the state is
+After the first Hadamard on the ancilla qubit, the state is
 %
@@
 %
-where $\left| d_{ij} \right\rangle = \left| d_{ij}^{(1)},\ldots,d_{ij}^{(n)} \right\rangle$.
-Applying the controlled phase gate on $\left| x^{(\alpha)} a \right\rangle$, where $\alpha = 1\dots n$  the state then becomes
+where $| d_{ij} \rangle = | d_{ij}^{(1)},\ldots,d_{ij}^{(n)} \rangle$.
+Applying the controlled phase gate on $| x^{(\alpha)} a \rangle$, where $\alpha \in  [1,n]$  the state then becomes
 %
@@
 %
-This completes the step of projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit.
+
+%%%%%
+\begin{figure*}[t]
+    \centering
+  \includegraphics[width=1.85\columnwidth]{vectorized_sample.pdf}
+  \caption{
+    (a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
+    Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
+    The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
+    (b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
+  }
+  \label{fig:vectorized_sample}
+\end{figure*}
+%%%%%
+
+This completes the step of projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the ancilla qubit.
 The process is done in the $x$-basis, achieved by the surrounding Hadamard gates.
-There are two possible measurement outcomes of the auxiliary qubit.
+There are two possible measurement outcomes of the ancilla qubit.
 Each pair of $\left|x_i \right\rangle$ and $\left| y_j \right\rangle$ forms a subspace of the Hilbert space,
@@
 %
-Thus we return into the initial basis and the amplitudes of the auxiliary qubit are proportional to how different each pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
+Thus we return into the initial basis and the amplitudes of the ancilla qubit are proportional to how different each pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ are.
 
@@
 %
+
 Currently available quantum platforms are still subject to a substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
-In order to reduce noise we average the measurement results over different states of the ancillary qubit, thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
+In order to reduce noise we average the measurement results over different states of the ancilla qubit, thus the measured Hamming distance between the input vector $\left| x_i \right\rangle$ and cluster vector $\left| y_j \right\rangle$ is
 %
@@
 %
-The Hamming distance measured in this way is bounded $0 \leq d_{i,j}^H \leq 1$,
+The Hamming distance measured in this way is bounded as $0 \leq d_{i,j}^H \leq 1$,
 where  $0$ indicates that $x_i$ and $y_j$ are identical and $1$ means they are the completely opposite in terms of their pairwise binary coordinates.
@@
 
-%%%%%
-\begin{figure}[t]
-    \centering
-	\includegraphics[width=0.95\columnwidth]{vectorized_sample.pdf}
-	\caption{
-		(a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
-		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
-		The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
-		(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
-	}
-	\label{fig:vectorized_sample}
-\end{figure}
-%%%%%
+\section{Experimental demonstration}
 
-\section{Experimental demonstration of QASOFM}
-
-We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.
+We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section with the use of the IBM quantum computer.
 We perform unsupervised data clustering for three sets of paper abstracts from three different fields:
@@
 We restricted our bag-of-word size to 9 of the most frequent words from the full bags-of-word  due to limitations on the number of qubits.
-After vectorizing and pre-processing  the data, the clusters are well-separated with the Hamming distance.
+After vectorizing and pre-processing  the data, the clusters are well-separated in terms of Hamming distance.
 We observe that distances between the abstracts inside clusters are smaller than distances between the abstracts from different clusters,
@@
 
-In classical SOFM \cite{kohonen1990}, first a distance calculation between a sample vector and all cluster vectors is made,
-then the closest cluster vector is shifted towards the sample vector.
-The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$,
-where $N$ is number of samples,
-$M$ is number of randomly sampled cluster vectors,
-and $L$ is number of the shifts of cluster vectors.
-In the QASOFM described in the previous section, distance calculations are realized on a quantum device (i.e.
-the IBM Q Experience) with the use of circuit presented in Fig.~\ref{fig:qcircuit}.
-This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates
-that are possible to realize on  quantum computing devices that are currently available.
-The calculation of Hamming distance implemented step-by-step between each sample vector and all cluster vectors is realized in one operation.
-In this case, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as the measurement no longer needs to indicate for which input vector the distance has been measured.
-The complexity of the quantum assisted SOFM then scales as $O(LN)$.
+\begin{figure}[t]
+  \includegraphics[width=\columnwidth]{distance_matrix.eps}
+  \caption{%
+    Comparison of classical and quantum calculations of Humming distance matrix.
+Result of tomography of all vector pairs that differ not more than one bit --- nearest neighbors.
+The distances are calculated as $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
+    where $d_{ij}$ was extracted from quantum state amplitudes with the use formula of ~(\ref{eq:amplitudes2distance}).
+    The calculations of amplitudes were implemented on the IBM quantum backend ``ibm\_sherbrook'' with statistics of 8192 shots.
+Distance values less than 1.7 are marked black and coincide to the nearest neighbors.
+    Question marks and crosses indicate missing and incorrect results compared to theoretical calculations.
+    The accuracy compared to classical simulation is better than 97 $\%$.
+  }
+  \label{fig:distance_matrix}
+\end{figure}
 
 In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.~\ref{fig:distance_matrix}.
-We see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
-The match is close, but not perfect, and we attribute it to noise in the currently available non fault-tolerant quantum processors.
-
-An example of the QASOFM learning process is shown in Fig.~\ref{convergence}.
-Initially, the cluster vectors were randomly chosen and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
-In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT\cite{qiskit} library.
-Each epoch of the algorithm consists of distance calculation between all data and cluster vectors
-and requires 9 distance calculations in the  quantum implementation ($N$ in general case)
-or 27 distance calculations for classical realization
-($MN$ in general case).
-After the distance calculation from each sample to all cluster vectors at each epoch we label each sample with the index of the closest cluster vector
-and shift the closest cluster vectors to the sample vectors.
-The shift is made by the change of the first binary element in the cluster vectors which does not match the binary element in the sample vector.
-The evolution of the labels presented on Fig.~\ref{convergence}(c).
-Good convergence is already observed in the fourth epoch.
-
-This proof of concept example shows that the development of noisy intermediate scale quantum hardware will allow to solve some practical problems in unsupervised manner with a very simple encoding of categorical data to the quantum register.
-Furthermore, it shows that it is possible to develop distance based hybrid quantum classical algorithms
-that speed up classical counterparts and can be implemented on near--term quantum devices
-developing distance based hybrid algorithms. Such algorithms,
-that can learn in an unsupervised manner, for categorical data such as genomic data has significant interest.
+We see very good agreement between the distance matrices calculated classically and on the quantum computer.
 
 %%%%%
-\begin{figure}[t]
-	\includegraphics[width=0.95\columnwidth]{convergence.png}
+\begin{figure*}[t]
+  \includegraphics[width=1.85\columnwidth]{convergence.png}
   \caption{
@@
   \label{convergence}
-\end{figure}
+\end{figure*}
 %%%%%
+
+An example of the QASOFM learning process is shown in Fig.~\ref{convergence}.
+Initially, the cluster vectors were randomly chosen and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
+For the cluster  vectors (binary representation of rows of Fig.
+\ref{convergence}(a)) we use the simplest basis encoding into quantum  states with the standard initialization of QISKIT \cite{rathQuantumDataEncoding2024,qiskit} library.
+Each epoch of the algorithm consists of a distance calculation between all data and cluster vectors
+and requires 9 distance calculations in the  quantum implementation ($N$ in the general case)
+or 27 distance calculations for the classical realization
+($MN$ in the general case).
+After the distance calculation from each sample to all cluster vectors at each epoch, we label each sample with the index of the closest cluster vector
+and shift the closest cluster vectors to the sample vectors.
+The shift is made by the change of the first binary element in the cluster vectors which does not match the binary element in the sample vector.
+The evolution of the labels is presented in Fig.~\ref{convergence}(c).
+Good convergence is already observed at the fourth epoch.
+
+This proof of concept example shows that the development of NISQ hardware will allow to solve some practical problems in an unsupervised manner with a very simple encoding of categorical data to the quantum register.
+Furthermore, it shows that it is possible to develop a distance based hybrid quantum classical algorithms
+that speeds up classical counterparts and can be implemented on near-term quantum devices
+developing distance based hybrid algorithms. Such algorithms,
+that can learn in an unsupervised manner, for categorical data such as genomic data has significant interest.
+
+\section{Complexity and scaling challenges}
+
+In this section, we discuss the complexity and the challenges that are present for scaling QASOFM to large problems sizes.
+
+\subsection{Complexity}
+
+In classical SOFM \cite{kohonen1990}, first a distance calculation between a sample vector and all cluster vectors is made,
+then the closest cluster vector is shifted towards the sample vector.
+The complexity of algorithm, in the sense of the number of distance calculations, scales as $O(LMN)$,
+where $N$ is number of samples,
+$M$ is number of randomly sampled cluster vectors,
+and $L$ is number of the shifts of cluster vectors.
+In the QASOFM described in the previous section, distance calculations are realized on the quantum computer with the use of quantum circuit presented in Fig.~\ref{fig:qcircuit}.
+This approach allows one to reduce the number of operations in a number of cluster states with an optimized number of gates
+that is possible to realize on quantum computers that are currently available.
+The calculation of Hamming distance implemented step-by-step between each sample vector and all cluster vectors is realized in one operation.
+In this case, as there is only one input vector considered, the ``Decoding'' stage (Fig.~\ref{fig:qcircuit}) can be removed as the measurement no longer needs to indicate for which input vector the distance has been measured.
+The complexity of the quantum assisted SOFM is then $O(LN)$.  Thus QASOFM algorithm has a quantum speedup by a factor of $ M $ over the classical algorithm.
+
+\begin{figure*}[t]
+  \includegraphics[width=1.85\columnwidth]{tomography-errors-by-shots-n4-n5.eps}
+  \caption{Error in the distance matrix of QASOFM versus number of runs of the quantum backend for (a) $4$ and (b) $5$ qubits.
+The triangles are the number of errors by the quantum scheme run on the IBM quantum backend named ``ibm\_sherbrook''.
+The lines are the result of a regression analysis of the noise simulator data.
+The coefficient of determination of the model is $R^2 = 0.86$ for the state space of $4$ qubits and $R^2 = 0.92$ for $5$ qubits.
+  }   \label{fig:tomography-errors-by-shots-n4}
+\end{figure*}
+
+\subsection{Scaling challenges}
+
+The above complexity estimate neglects the steps required to prepare the initial states (\ref{eq:encodnig}).
+This is a standard problem that affects many quantum machine learning techniques and is called the ``input problem'' \cite{biamonte2017}.
+It is typically assumed that there is an external resource (e.g.
+a qRAM) that can be used to prepare such states.
+There are, however, potential ways to alleviate this by using a more effective encoding scheme into quantum states.
+In the proposed QASOFM as given in Sec.
+\ref{sec:qasofm} we used the simplest encoding model.
+This model is inefficient with respect to quantum resources.
+Furthermore, the bag-of-word model is not optimal itself (we used it because of very small classical data needed to be embedded).
+The use of better word embedding models will allow to improve QASOFM performance.
+Also, the dimension of the input data is usually much larger than the number of cluster vectors.
+Consequently, the size of the subspace spanned on the cluster vectors is negligible compared to the whole space.
+This in theory provides opportunities to efficiently generate quantum circuits for initialization of such superposition of sparse quantum states with a smaller number of operations.
+
+On the other hand, our QASOFM avoids the use exponential resources for the read-out  (i.e. no ``output problem'' due to the need of quantum tomography \cite{biamonte2017}).
+The SOFM assumes that the cluster points do not have to move monotonically towards the cluster centroids,
+and that the vectors closest to the true BMU have the highest probability of being measured.
+Thus, the number of runs should only be chosen to measure distances to nearest neighbors with high accuracy, not all possible distances.
+The protocol implies finding all pairs of nearest vectors in the whole state space.
+Fig.~\ref{fig:tomography-errors-by-shots-n4} shows the error versus number of runs for determination of the distance matrix for $4$  and $5$ qubits.
+The points show the number of errors in the neighbors versus the the number of quantum circuit runs using a noise simulator using real parameters.
+The number of errors decrease exponentially with the number of shots, which shows the efficient performance of our algorithm with respect to read-out.
+
+Another challenge relates to the connectivity limitations of currently available quantum computers.
+In QASOFM, the ancilla qubit accumulates the distances in the phase information and as such it ideally requires a high connectivity to the $X $ register.
+The topology of currently available superconducting quantum processors  presents a significant limitation and consequently to the scaling of the QASOFM algorithm.
+A possible solution is to create a large auxiliary network,
+but it requires many two-qubit operations, which reduces the accuracy.
+Another solution is to use platforms with a star architecture (or all to all connected) which are under development with trapped ion technology \cite{QuantinuumsHSeriesHits}.
 
@@
 We introduced an optimized circuit for Hamming distance calculations that can be implemented on currently available quantum computing devices with high fidelity.
-Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm
+Our quantum circuit performs the distance-computing component of a classical SOFMs algorithm in a parallelized way,
 and in this way improves its performance.
@@
 and $L$ is number of the shifts of cluster vectors.
+We showed that the number of errors decreases exponentially with the number of runs of the algorithm, highlighting the efficiency of our algorithm.
 Due to wide use of classical SOFMs in different areas of modern research and technology,
-this can give opportunities for the use of QASOFM in practical applications in the near term, outperforming classical algorithms.
+this can give opportunities for the use of QASOFM in practical applications in the near term.
+
 In addition, as our algorithm performs the Hamming distance calculation,
-it has the potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
-In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure
-and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required,
-but when knowledge of nearest vectors is sufficient.
+it has the potential to enhance any classical algorithm that relies on calculating distances between binary data.
+In particular, recently many algorithms are under development for quantized datasets with binary embeddings \cite{zhuangFastTrainingTripletBased2016,yiBinaryEmbeddingFundamental2015,AsymmetricDistancesBinary}.
+Furthermore, quantum protocols for Euclidean distance calculations have been developed \cite{yuQuantumAlgorithmsSimilarity2020,zardiniQuantumKnearestNeighbors2024}, which allows the generalization of QASOFM to non-binary data.
+In machine learning, data science, statistics and optimization, distance is a common way of representing similarity.
+Calculating it between large data sets is common procedure
+and our circuit could potentially enhance other distance based algorithms for binary data, particularly when the exact distance is not required,
+and the knowledge of nearest vectors is sufficient.
 
 \section*{Acknowledgments}
-The work was performed as a part of a state task, State Registration No. AAAA-A19-119071190017-7.
+The work was partly supported by the
+Russian Science Foundation (grant no 23-21-00507) and partly performed as a part of a state task, State Registration No. 124013000760-0.
 T.B. is supported by the National Natural Science Foundation of China (62071301);
@@
 
-\section*{Conflict of interest}
+\appendix
 
-The authors declare no conflict of interest.
+\section{Source code}
 
-\section*{Keywords}
-quantum self-organizing feature map; quantum machine learning; neural network; quantum unsupervised data clustering; Hamming distance.
+The source code is available on https://github.com/kephircheek/qasofm.
 
-\bibliographystyle{naturemag}
-\bibliography{bibliography}
+% \section*{Conflict of interest}
+
+% The authors declare no conflict of interest.
+
+% \section*{Keywords}
+% quantum self-organizing feature map; quantum machine learning; neural network; quantum unsupervised data clustering; Hamming distance.
+
+\bibliographystyle{apsrev4-2}
+\bibliography{bibliography,bibl}
 

### 2024-12-12 manuscript/main.tex
Clafiry 'limitation to connectivity'

@@
 In QASOFM, the ancilla qubit accumulates the distances in the phase information and as such it ideally requires a high connectivity to the $X $ register.
-The topology of currently available superconducting quantum processors  presents a significant limitation and consequently to the scaling of the QASOFM algorithm.
+The topology of currently available superconducting quantum processors  presents a significant limitation to connectivity and consequently to the scaling of the QASOFM algorithm.
 A possible solution is to create a large auxiliary network,

### 2024-12-14 manuscript/main.tex
Use more common way 'superscriptaddress'
Merge bibl into bibliography
Use default revtex bibliography style with names
Replace vectorized_sample.png with *.pdf

--- 2024-12-11 manuscript/main.tex
@@
-\documentclass[pra,showkeys,twocolumn,showpacs,aps,10pt]{revtex4-1}
+\documentclass[pra,showkeys,twocolumn,showpacs]{revtex4-1}
 \usepackage{amsmath}
@@
 \section{Introduction}
-The combination of big data and artificial intelligence (AI) --  dubbed the fourth industrial revolution -- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
+The combination of big data and artificial intelligence ---  dubbed the fourth industrial revolution --- has profoundly affected the modern economy in a plethora of different ways from robotics to agriculture \cite{Lecun2015, ghahramani2015,schwab2017,esteva2019, tyrsa2017}.
 Contemporary artificial intelligence methods based on neural networks also have the potential to enhance the role of novel analytical methods in science and engineering \cite{kaggle2014, radovic2018, butler2018, radovic2018}.
-From 2022, AI and neural networks experienced remarkable advancements, sparking substantial excitement in the field, particularly with the emergence of Large Language Models (LLMs) such as ChatGPT-4\cite{GPT4}.
-In 2024, Nobel prizes in physics\cite{NobelPrizePhysics} and chemistry\cite{NobelPrizeChemistry} were awarded for foundational discoveries for developement of machine learning and applications of AI in chemistry.
-Despite such groundbreaking achievements, interpretability and trustworthiness of such NNs is very low due to the quantity of parameters in such neural networks rapidly increase.
-Paradoxically, the intrinsic mechanism of how such neural networks work and why they are so powerful remains unknown (in many cases it is regarded as universal black box oracle).
-Additionally, these black-box oracles demand substantial resources for their training and exhibit a propensity to overfit when data is scarce \cite{sevilla2022compute}, in contrast to the human ability to effectively learn from "small" data.
+Paradoxically, the intrinsic mechanism of how neural networks work and why they are so powerful remains unknown (in many cases it is regarded as a black box).
+It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,
+and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
 
-%It has been speculated that limits to the neural network paradigm based on the computational power of von Neumann architecture are being approached,and improvements appear only due to heuristic escalation of complexity \cite{marcus2018,sze2017,kourtis2020}.
+In particular, the self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN)
+that is trained in an unsupervised manner.
+SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
+and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
+The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
+Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
 
-On the other hand, quantum computation is another fast growing area of research that is moving beyond its academic roots.
-Although, its present scale may not yet rival classical computation, there is considerable excitement within this field and high expectations for rapid developments for commercial opportunities in the near future.
-In 2022, the Nobel Prize in Physics honored groundbreaking contributions to quantum information science\cite{NobelPrizePhysicsa}, the Breakthrough Prize in Fundamental Physics in 2023 recognized achievements in quantum communication and quantum computing\cite{BreakthroughPrizeWinners}.
-From practical point of view, IBM  empowered their NISQ processors with more than 1000 qubits\cite{IBMRoadmapQuantumcentric}, Google demonstrated quantum error correction below the surface code threshold\cite{acharyaQuantumErrorCorrection2024}, and QuEra ambitiously announced 100 logical qubit processor in 2026 in their roadmap\cite{QuantumErrorCorrection}.
-A lot of new algorithms were developed for such devices (see \cite{Cerezo2021,Bharti2022} for comprehensive review).
-In particular, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019, Benedetti2019, Cerezo2022, Sajjan2022}.
-In comparison with classical ML, it was recently shown that Quantum Machine Learning (QML) models can be traine with fewer parameters in comparison with classical ANNs\cite{abbasPowerQuantumNeural2021,caroGeneralizationQuantumMachine2022}.
+Meanwhile, there has been much interest recently in applying quantum computing techniques to machine learning \cite{dunjko2018, biamonte2017, schuld2014, carleo2019}.
 The main focus of early works in quantum machine learning (QML) was in obtaining a quantum speedup \cite{biamonte2017, schuld2014} by applying quantum approaches to solve linear algebra problems, such as the Fourier transform or solving systems of linear equations \cite{wiebe2012,harrow2009,childs2017}.
@@
 The interest in quantum neural networks was inspired by progress in experimental quantum computing when it became possible to use parametrized quantum circuits,
-where the parameters behave much like the weights of a neural network \cite{lewenstein1994,Benedetti2019}.
-In particular, quantum algorithms for training and evaluating feed forward neural networks were developed,
+where the parameters behave much like the weights of a neural network \cite{lewenstein1994}.
+In particular, quantum algorithms for training and evaluating feedforward neural networks were developed,
 which are one of the most usable neural network models \cite{allcock2018, tacchino2019}.
 Quantum models for convolutional neural networks, which may be suitable for the problems of learning of quantum states were also proposed \cite{cong2019, liu2019}.
-Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019,daiQuantumBayesianOptimization2023}.
+Recent results connecting the classical Bayesian approach to deep learning allowed for the development of a new algorithm for Bayesian deep learning on quantum computers \cite{zhao2019}.
 For the problem of classification which is closely connected to the problem of clusterization, a protocol of quantum classification, tested on the MNIST dataset, via slow feature analysis based on the use of quantum Frobenius distance was proposed \cite{kerenidis2018}.
-A quantum capsule network together with an efficient quantum dynamic routing algorithm was proposed \cite{Liu2022}.
-The quantum versions of transformer architecture and transformer of quantum states were developed\cite{Viteritti2023,cherratQuantumVisionTransformers2024,khatriQuixerQuantumTransformer2024,zhangTransformerQuantumState2023} that provides possibilities for quantum natural language processing.
-Inspired by the classical counterpart, the quantum denoising diffusion probabilistic model (QuDDPM) was proposed to enable efficiently trainable generative learning of quantum data\cite{zhangGenerativeQuantumMachine2024}.
-Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{kolleQuantumDenoisingDiffusion2024,li2019, Heese2022, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
+Furthermore, many other hybrid and quantum protocols inspired classical neural networks were developed  \cite{li2019, killoran2019, bondarenko2019, dunjko2017, nautrup2019, foesel2018, rebentrost2018, purushothaman1997, verdon2019, cherny2019, byrnes2013, mishra2019, pyrkov2019, vinci2019, lu2019}.
 
@@
 \cite{kandala2017,aspuru-guzik2005,lanyon2010,peruzzo2014}.
-Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on Noisy Intermediate Scale Quantum (NISQ) devices \cite{preskill2018,Bharti2022,Cerezo2021}.
+Currently, it is believed that implementation of quantum neural networks and hybrid quantum classical algorithms can be the main test bed to achieve practical quantum supremacy on Noisy Intermediate Scale Quantum (NISQ) devices \cite{preskill2018}.
 
 In this paper, we develop a hybrid quantum-assisted SOFM (QASOFM)
-and apply it to the data clustering problem in an unsupervised manner. The self-organizing feature map (SOFM) \cite{kohonen1990,kohonen1996,kohonen1997}, is a type of artificial neural network (ANN)
-that is trained in an unsupervised manner.
-SOFMs are used in many areas \cite{vilibic2016, guido1998, doszkocs1990, jones2012,mori2019,corsello2017,zhu2018,chea2016}
-and in comparison with many other artificial neural networks, they apply competitive learning and preserve the topological properties of the input space \cite{kiviluotoa1996}.
-The SOFMs represent data in a fundamentally topological way that allows one to perform dimensionality reduction.
-Once it is trained, the map can classify a vector from the input space by finding the node with the smallest distance metric.
-The idea of QASOFM is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
+and apply it to the data clustering problem in an unsupervised manner.
+The idea is based on the use of the Hamming distance as a distance metric for the training the SOFM that allows, in the quantum case, to reduce the number of distance calculations in the number of clusters and thus to speed up the original classical protocol.
 In order to make our protocol more appropriate for the currently available generation of NISQ quantum devices, we optimized a circuit for realizing the Hamming distance by reducing number of one-qubit operations.
@@
 
+%%%%%
 \begin{figure}[t]
 	\includegraphics[width=\columnwidth]{qcircuit.eps}
-  \caption{%
-    Example of quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of 4-bit binary vectors from two sets ${X}$ and ${Y}$.
+    \caption{
+		Quantum circuit for the quantum parallelized Hamming distance calculation between all pairs of binary vectors from two sets ${X}$ and ${Y}$.
 		First, we encode information about pairwise different qubits in a quantum state of the $X$-register by applying the CNOT gates.
@@
 \end{figure}
+%%%%%
 
-In the work \cite{trugenberger2001}, the concept of probabilistic quantum memory was introduced.
-In this framework, quantum entanglement is utilized to significantly enhance the storage capacity of associative memories and to retrieve noisy or incomplete information.
-Specifically, the number of binary patterns that can be stored in such a quantum memory is exponential in relation to the number $ n $ of qubits, with a maximum capacity of $ p_{\text{max}} = 2^n $.
-This means it can optimally store all possible binary patterns that can be formed with $ n $ bits.
-
-The retrieval process is probabilistic and involves postselection of the measurement results.
-This implies that the retrieval algorithm must be repeated until a certain threshold $T$ is met, or until the measurement of a control qubit yields a specific outcome.
-In the first scenario, the input is not recognized, while in the second, the output is determined by a probability distribution over the memory that is concentrated around the stored patterns that are closest to the input, as measured by Hamming distance.
-Here, we improve the approach for Hamming distance calculation proposed for realization of probabilistic quantum memory\cite{trugenberger2001} and apply it for realization of QASOFM.
+%%%%%
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{distance_matrix.png}
+	\caption{
+		Hamming distance matrix between two data sets of binary vectors.
+		Values of distance that less than median distance value was marked black.
+		The classical simulation of the quantum circuit shows perfect agreement to theoretical calculations and presented on the left figure.
+		Result obtained on  IBMQ ``ibmq\_16\_melbourne'' backend is shown on the right figure and the accuracy in comparison with the classical simulation is more than 92 $\%$.
+	}
+	\label{fig:distance_matrix}
+\end{figure}
+%%%%%
 
@@
 %
-
-%%%%%
-\begin{figure*}[t]
-    \centering
-  \includegraphics[width=1.85\columnwidth]{vectorized_sample.pdf}
-  \caption{
-    (a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
-    Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
-    The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
-    (b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
-  }
-  \label{fig:vectorized_sample}
-\end{figure*}
-%%%%%
-
 This completes the step of projecting differences between pairs of $\left| x_i \right\rangle$ and $\left| y_j \right\rangle$ onto the amplitude of the auxiliary qubit.
@@
 %
-
 Currently available quantum platforms are still subject to a substantial level of noise and extracting the exact distance from amplitude is still a difficult task.
@@
 %
-\begin{equation}\label{eq:amplitudes2distance}
-    d_{i,j}^H \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))).
-\end{equation}
+\begin{align}
+    d_{i,j}^H & \propto 1 - \frac{1}{2}(a_0(x_i,y_j) + (1-a_1(x_i,y_j))) .
+\end{align}
 %
@@
 
+%%%%%
+\begin{figure}[t]
+    \centering
+	\includegraphics[width=0.95\columnwidth]{vectorized_sample.pdf}
+	\caption{
+		(a) Representation of the data set of abstracts with the bag-of-words \cite{weikang2016} model is shown.
+		Each abstract is represented by a binary vector with 9 elements, corresponding to the 9 words on the horizontal axis.
+		The samples are sorted into groups ``Quantum Machine Learning'' (QML), ``Cancer'' (MED) and ``Gene Expression'' (BIO) with 3 papers for each tag, for a total of 9 paper.
+		(b) The Hamming distance between each vectorized abstract is shown as a number in the matrix.
+	}
+	\label{fig:vectorized_sample}
+\end{figure}
+%%%%%
+
 \section{Experimental demonstration of QASOFM}
 
-We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section with the use of IBM Q Experience.
-
+We now show experimental results for a proof-of-concept demonstration of the algorithm introduced in the previous section, on a 16-qubit quantum computer provided by IBM Q Experience.
 We perform unsupervised data clustering for three sets of paper abstracts from three different fields:
@@
 
-\begin{figure}[t]
-  \includegraphics[width=\columnwidth]{distance_matrix.eps}
-  \caption{%
-    Comparison of classical and quantum calculations of Humming distance matrix.
-Result of tomography of all vector pairs that differ not more than one bit --- nearest neighbors.
-The distances are calculated as $\bar d_{ij} = (d_{ij} + d_{ji}) / 2$,
-    where $d_{ij}$ was extracted from quantum state amplitudes with the use formula of ~(\ref{eq:amplitudes2distance}).
-    The calculations of amplitudes were implemented on the IBM quantum backend named ``ibm\_sherbrook'' with statistics of 8192 shots.
-Distance values less than 1.7 are marked black and coincide to the nearest neighbors.
-    Question marks and crosses indicate missing and incorrect results compared to theoretical calculations.
-    The accuracy compared to classical simulation is better than 97 $\%$.
-  }
-  \label{fig:distance_matrix}
-\end{figure}
-
 In order to check that our algorithm gives the expected results we compare it to classical calculations of the distance matrix on two data sets of binary vectors, as shown in Fig.~\ref{fig:distance_matrix}.
-We see very good agreement between the distance matrices calculated classically and on the IBM Q Experience.
-
-%%%%%
-\begin{figure*}[t]
-  \includegraphics[width=1.85\columnwidth]{convergence.png}
-  \caption{
-    (a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2),
-    for the selected 9 words from the bag of words model.
-    (b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend.
-    Vectors mean cluster elements for BIO, MED, QML groups,
-    for the C0, C1, C2 cluster vectors, respectively.
-    (c) The evolution of label distribution on each learning epoch.
-  }
-  \label{convergence}
-\end{figure*}
-%%%%%
+We see good agreement between the distance matrices calculated classically and on the IBM Q Experience.
+The match is close, but not perfect, and we attribute it to noise in the currently available non fault-tolerant quantum processors.
 
@@
 Initially, the cluster vectors were randomly chosen and the label of sample distribution is shown for the zeroth epoch in Fig.~\ref{convergence}(a).
-For the cluster  vectors (binary representation of rows of Fig.\ref{convergence}a) we use the simplest basis encoding into quantum  states with the standard initialization of QISKIT\cite{rathQuantumDataEncoding2024,qiskit} library.
+In order to prepare a superposition of cluster vectors needed for the calculation of the distance matrix we use the standard initialization of QISKIT\cite{qiskit} library.
 Each epoch of the algorithm consists of distance calculation between all data and cluster vectors
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=1.85\columnwidth]{tomography-errors-by-shots-n4-n5.eps}
-  \caption{Error in the distance matrix of QASOFM versus number of runs of the quantum backend for (a) $4$ and (b) $5$ qubits.
-We can see exponential decrease in error with linear increase in number of runs.
+%%%%%
+\begin{figure}[t]
+	\includegraphics[width=0.95\columnwidth]{convergence.png}
+	\caption{
+		(a) Initial random binary vectors of cluster vectors (labeled as C0, C1, C2),
+		for the selected 9 words from the bag of words model.
+		(b) The result of applying our QASOFM implemented on the IBM Q Experience simulator backend.
+		Vectors mean cluster elements for BIO, MED, QML groups,
+		for the C0, C1, C2 cluster vectors, respectively.
+		(c) The evolution of label distribution on each learning epoch.
 	}
-  \label{fig:tomography-errors-by-shots-n4}
-\end{figure*}
-
-\section{Scalability Of QASOFM}
-
-For scalability of NISQ algorithms both software and hardware restrictions are important.
-
-From software point of view the main challenge is effective encoding of classical data into quantum states.
-In the proposed QASOFM we used the simplest basis encoding model.
-This model is non-efficient in the meaning of required quantum resources.
-Furthermore, the bag-of-word model is not optimal itself (we used it because of very small classical data needed to be embedded).
-The use of better word embedding models will allow to improve QASOFM performance.
-Also, the dimension of the input data is usually much larger than the number of cluster vectors.
-Consequently, the size of the subspace spanned on the cluster vectors will be negligible compared to the whole space.
-This in theory provides opportunities to efficiently generate quantum circuits for initialization of such superposition of sparse quantum states with smaller number of operations.
-
-The second challenge is in currently available hardware.
-In QASOFM, the ancillary qubit  accumulates the distances in phase and because of that requires to be as much connected as possible to other qubits.
-The topology of superconducting quantum processors,such as IBM or Google Sycamore, presents the significant limitations for that and consequently for the scaling of QASOFM algorithm to a larger problem.
-The possible solution is to create a large auxiliary network,
-but it requires a lot of two-qubit operations,
-which are currently performed with insufficient accuracy.
-Another solution is to use platforms with star architecture (all to all connected) which are under development with trapped ion technology\cite{QuantinuumsHSeriesHits}.
-It will higly improve performance and scalability.
-
-Furthermore, our QASOFM allows to avoid the use exponential resource requirement at read-out operation (as for quantum tomography). 
-SOFM assumes that the cluster points do not have to move monotonically towards the cluster centroids,
-and that the vectors closest to the true BMU have the highest probability of being measured.
-Thus, the number of runs should only be chosen to measure distances to nearest neighbors with high accuracy, not all possible distance calculations.
-The protocol implies finding of all pairs of nearest vectors in the whole state space.
-Fig.~\ref{fig:tomography-errors-by-shots-n4} shows the error versus number of runs for determination of distance matrix for $4$  and $5$ qubits respectively.
-The points in Fig.~\ref{fig:tomography-errors-by-shots-n4} define the number of errors in neighbors
-by the quantum circuit runs on a noise simulator taking into account parameters of the real quantum platform.
-The lines are the result of a regression analysis of these data.
-The coefficient of determination of the model is $R^2 = 0.86$ for the state space of $4$ qubits and $R^2 = 0.92$ for $5$ qubits.
-The triangles are the number of errors by the quantum scheme run on the IBM quantum backend named ``ibm\_sherbrook''.
-The number of errors decreases exponentially as the number of shots increases linearly. This provide advantage for our protocol in read out.
+	\label{convergence}
+\end{figure}
+%%%%%
 
@@
 In addition, as our algorithm performs the Hamming distance calculation,
-it has the potential to enhance any classical algorithm that relies on calculating distances between binary data.
-In particular, nowadays a lot of algorithms are under development for quantized datasets with binary embeddings\cite{zhuangFastTrainingTripletBased2016,yiBinaryEmbeddingFundamental2015,AsymmetricDistancesBinary}.
-Furthermore, recently quantum protocols for euclidean distance calculations were introduced\cite{yuQuantumAlgorithmsSimilarity2020,zardiniQuantumKnearestNeighbors2024} that allows to generalize QASOFM to non-binary data.
+it has the potential to enhance any classical algorithm that relies on calculating distances between data entries of vector form.
 In machine learning, data science, statistics and optimization, distance is a common way of representing similarity, calculating it between large data sets is common procedure
-and our circuit could potentially enhance other distance based algorithms for binary data  as long as exact distance is not required,
+and our circuit could potentially enhance other distance-based algorithms as long as exact distance is not required,
 but when knowledge of nearest vectors is sufficient.
@@
 \section*{Acknowledgments}
-The work was partly supported by the
-Russian Science Foundation (grant no 23-21-00507) and partly performed as a part of a state task, State Registration No. 124013000760-0.
+The work was performed as a part of a state task, State Registration No. AAAA-A19-119071190017-7.
 T.B. is supported by the National Natural Science Foundation of China (62071301); 
@@
 
-\section*{Data Availability Statement}
-
-The source code is available on https://github.com/kephircheek/qasofm
-
 \section*{Conflict of interest}
@@
 \bibliographystyle{naturemag}
-\bibliography{bibliography,bibl}
+\bibliography{bibliography}
 
